### 2024-08-26 research/notes.tex
created report & exported as latex & pdf

% Created 2024-08-25 Sun 14:46
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\author{Mark Krutzler}
\date{\today}
\title{Notes / Research on CPU-Scheduling}
\hypersetup{
 pdfauthor={Mark Krutzler},
 pdftitle={Notes / Research on CPU-Scheduling},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 29.4 (Org mode 9.8)}, 
 pdflang={English}}
\begin{document}

\maketitle
\tableofcontents

\part{Basics (OStep)}
\label{sec:orgce4bae2}
\chapter{Introduction (Chapter 7)}
\label{sec:org02640b6}
Scheduling is not a low level mechanism but a high level policy/disciplines.
We need to make simplifying assumptions of the workload:
\begin{enumerate}
\item Each job runs for the same amount of time
\item All jobs arrive at the same time.
\item Once started, each job runs to completion
\item All jobs only use the CPU (no I/O)
\item The run-time of each job is known.
\end{enumerate}
These rules will be eased over time until we get a fully functioning policy.
Of course the more you know the easier it is to schedule.
\section{Metrics}
\label{sec:orgb0f9911}
The fundamental question is: How to we measure the ``efficiency'' or the ``quality'' of the scheduler?
You can measure performance or fairness. Here are ways to measure performance:
\begin{itemize}
\item turnaround time
\begin{itemize}
\item Calculated as:
\(T_{turnaround} = T_{completion} - T_{arrival}\)
For us \(T_{arrival} = 0\), because of simplification 1. (can be neglected later)
\end{itemize}
\item response time: measures the frustration of the user, while looking at the spinning ball
\begin{itemize}
\item Calculated as:
\(T_{response} = T_{firstrun} - T_{arrival}\)
\begin{itemize}
\item For modern computers, it is essential that this is kept at a minimum
\end{itemize}
\end{itemize}
\item fairness: first job to finish divided by last job to finish (this is not a performance metric!!)
\end{itemize}
\section{First In, First Out (FIFO) / First Come, First Served (FCFS)}
\label{sec:org930f0a2}
\begin{itemize}
\item Most basic scheduling policy
\item Given our simplification it works really well and is easy to implement
\item However after relaxing assumption 1, it will perform poorly if a huge process gets infront of many small ones
\begin{itemize}
\item This is the so called \textbf{convoy effect}
\item It is like if you're at waiting in line to pay and before you have a family of five with two full carts: annoying
\end{itemize}
\end{itemize}
\section{Shortest Job First (SJF)}
\label{sec:org1f10514}
\begin{itemize}
\item The shortest job is run first
\begin{itemize}
\item non-preemptive: runs a process until finish
\item preemptive: can stop and perform a context switch
\end{itemize}
\item If the smaller tasks arrive later (by relaxing assumption 2), then we face the same problem as before. (due to this algorithm can't perform a context switch / is non-preemptive)
\end{itemize}
\section{Shortest Time-to-Completion (STCF) / Preemptive Shortest Job First (PSJF)}
\label{sec:orgb6c2621}
\begin{itemize}
\item This policy requires that rule 3 is ignored.
\item This is the preemptive version of SJF.
\item It updates, whenever a new job arrives or one is finished
\end{itemize}
\section{Round Robin (RR) / time-slicing}
\label{sec:orgc7555bf}
\begin{itemize}
\item this policy runs each job for a specified ``time slice'' / ``scheduling quantum'' (introducing a variable)
\item general technique is called ``amortization''.
\item The shorter the time slice, the more responsive the system, however context switching costs CPU time aswell, so you'll need to balance out
\item RR is one of the worst policies for turnaround time
\item It gives up performance for fairness
\end{itemize}
\section{Relaxing Assumptions 4 \& 5}
\label{sec:orgac13a29}
\begin{enumerate}
\item assumption 4
\label{sec:org61b9a39}
\begin{itemize}
\item If a job waits for I/O than it is in a state called ``blocked''
\item While a job is waiting for I/O, the CPU can be passed onto somebodye else: ``overlapping''
\end{itemize}
\item assumption 5
\label{sec:org64b9783}
\begin{itemize}
\item we usually have no idea how long a job will take
\item This actually breaks most of our policies, because they all rely on knowing the length of the job (except RR)
\item Solution: Multi-Level Feedback Queue (MLFQ) \(\Rightarrow\) See next Chapter
\end{itemize}
\end{enumerate}
\chapter{Multi-Level Feedback Queue (Chapter 8)}
\label{sec:org36243f5}
\begin{itemize}
\item One of the most known Policies (Turning Awarded)
\item It tries to:
\begin{itemize}
\item optimize turn around time (without knowing the length of the job)
\item minimize response time
\end{itemize}
\end{itemize}
\section{(Basic) Rules of MLFQ}
\label{sec:orga3d0fb4}
\begin{itemize}
\item There are multiple queues and each has their priority level. (higher priority is preferred when switching)
\item If multiple jobs are on the same priority than RR (Round Robin) is used
\item Priorities can change over time.
\item Assume that if a job is resource intensive than it will stay as such. (The history of the job determines the future)
\end{itemize}
\section{Changing Priority}
\label{sec:org4946c2d}
\begin{itemize}
\item Depending on the CPU time usage, the priority changes
\item ``allotment'': time that a job can spend at a given priority before demotion.
\end{itemize}
\section{Priority Boost}
\label{sec:orgfe8a82e}
\begin{itemize}
\item to counter starvation of longer jobs every now and then all of the jobs are put into the priority queue
\item also this counters the fact that some programs might start non interactively and than turn into interactive (you know what I mean)
\end{itemize}
\section{``Better Accounting'' (Anti gaming)}
\label{sec:orgd12a267}
\begin{itemize}
\item to prevent people from abusing the allotment method and game the CPU, we need to update rule 4:
\item previous:
\begin{enumerate}
\item If a job uses up its allotment while running, its priority is reduced
\item If a job gives up the CPU before the allotment is up, it stays at the same priority
\end{enumerate}
\item new:
Once a job uses up its time allotment at a given level, its priority is reduced
\end{itemize}
\section{Summary of Rules (Copied out of the book)}
\label{sec:orga6a11c6}
\begin{enumerate}
\item If Priority (A) > Priority (B) \(\Rightarrow\) A runs \& B doesn't
\item If Priority (A) = Priority (B) \(\Rightarrow\) A \& B run in RR
\item When a job enters the system, it is placed at the highest priority
\item Once a job uses up its time allotment at a given level, its priority is reduced
\item After some period S, move all the jobs in the system to the topmost queue
\end{enumerate}
\section{Voo-Doo Constants}
\label{sec:org578a8d8}
These constants heavily change how effective the MLFQ is:
\begin{itemize}
\item scheduling quantum (RR)
\item amount of queues
\item when to priority boost
\item allotment (could change in every priority queue)
\end{itemize}
\chapter{Proportional Share (Chapter 9)}
\label{sec:org0922768}
\begin{itemize}
\item This is a fair scheduler
\begin{itemize}
\item The more/longer jobs run the fairer it becomes
\end{itemize}
\item literally just hold a lottery to determine which programs runs next
\item ``tickets'' represent the share of a resource that a process should recieve = it is like a currency
\begin{itemize}
\item the more tickets you hold, the higher the chance that you have a winning one
\item every time slice a new ticket is picked out as the winning ticket
\item more generally tickets can represent the share of something.
\end{itemize}
\item the tickets are handed out to the user, who than can allocate among their jobs
\begin{itemize}
\item the user can use their ``own'' tickets which will be converted into the global currency
\end{itemize}
\item ticket transfer can be used to boost a process
\begin{itemize}
\item think server / client => client give server their tickets, so that the server has a higher global share
\end{itemize}
\item ?? in a trusted environment you could also inflate your own tickets to boost you own CPU time
\end{itemize}
\section{Advantages of using randomness}
\label{sec:org26fc12b}
\begin{itemize}
\item no strange corner-case behaviors
\item lightweight
\item if the randomizing algorithm is quick than the speed is quick
\begin{itemize}
\item faster algorithms tend to be more like pseudo-random
\end{itemize}
\end{itemize}
\section{Implementation}
\label{sec:org36bc201}
\begin{enumerate}
\item requirements:
\label{sec:orgf067f7d}
\begin{itemize}
\item random number generator
\item data structure (to track the processes of the system)
\item amount of total number of tickets
\end{itemize}
\item sample code (copied):
\label{sec:orgc4e4c39}
\begin{verbatim}
// counter: used to track if we’ve found the winner yet
int counter = 0;

// winner: call some random number generator to
//         get a value >= 0 and <= (totaltickets - 1)
int winner = getrandom(0, totaltickets);

// current: use this to walk through the list of jobs
node_t *current = head;
while (current) {
    counter = counter + current->tickets;
    if (counter > winner)
        break; // found the winner
    current = current->next;
}
// ’current’ is the winner: schedule it...
\end{verbatim}
\end{enumerate}
\section{Assigning tickets}
\label{sec:orga879a75}
\begin{itemize}
\item Remains open for now
\end{itemize}
\section{Stride Scheduling}
\label{sec:org3130976}
\begin{itemize}
\item it is a deterministic fair-share scheduler
\begin{itemize}
\item while lottery scheduling achieves the proportions with probability (can be off), stride scheduling gets it right each time.
\item PROBLEM: you can't have a new job entering, because it will monopolize the CPU (due to low pass value)
\end{itemize}
\item bit tricky to understand: there is another article about it that I'll later read (under heading Lottery Scheduling)
\item how it works:
\begin{itemize}
\item each process has a stride to begin with (the more tickets the smaller the stride)
\item each time the process runs, it's counter (called ``pass'') get incremented by the value of the stride
\begin{itemize}
\item this is tracking its global progress
\end{itemize}
\item scheduler schedules according to the pass and the stride
\begin{itemize}
\item pick the lowest pass
\end{itemize}
\end{itemize}
\end{itemize}
\begin{enumerate}
\item pseudo-implementation (code copied)
\label{sec:org4c549fe}
\begin{verbatim}
current = remove_min(queue); // pick client with min pass
schedule(current); // run for quantum
current->pass += current->stride; // update pass using stride
insert(queue, current); // return current to queue
\end{verbatim}
\end{enumerate}
\section{Sidequest: Linux Completely Fair Scheduler (CFS)}
\label{sec:orgf8d6ba7}
\begin{itemize}
\item will talk about it later as well
\item every process has a counter called ``vruntime''
\begin{itemize}
\item as they run it increases
\item the process with the lowest ``vruntime'' is next
\begin{itemize}
\item PROBLEM: while waiting / in I/O the process vruntime is not increased: after coming back alive, it'll monopolize the CPU
\item SOLUTION: Once a process wakes up, it will take the lowest amount of vruntime
\item PROBLEM: short sleep will make it less fair for you for you
\end{itemize}
\end{itemize}
\item the switching is controlled through parameters:
\begin{itemize}
\item sched\textsubscript{latency}: dynamic time slice (is calculated), typically 48ms divided by n number of processes
\item sched\textsubscript{latency} basicly determines the maximal time frame until each process has run atleast once (if not controlled for minimun time slice)
\end{itemize}
\item There is also a minimal time slice:
\begin{itemize}
\item min\textsubscript{granularity} (set to usually 6ms) ensures that each process runs atleast a certain amount of time switching
\begin{itemize}
\item else the context switch would be too expensive
\item with this the scheduler becomes less fair, when only looking at the sched\textsubscript{latency}, however it is a good tradeoff
\end{itemize}
\end{itemize}
\item CFS utilizes the periodic timer interrupt. This means every 1ms it can wake up and determine what to do next
\end{itemize}
\begin{enumerate}
\item Niceness (Priority setting)
\label{sec:orgb6d68eb}
\begin{itemize}
\item priority setting is done through the ``nice'' level
\begin{itemize}
\item default: 0 (min: +19, max: -20)
\item the level will be mapped to a ``weight'' according to a premade table
\begin{itemize}
\item this will keep the proportianility
meaning: if you have a difference of 5 levels between two jobs, than the ratio of sharing stays the same
\end{itemize}
\item The time slice is calculated as followed:
$$timeslice_k = \frac{weight_k}{\sum_{i=0}^{n-1} weight_i} * schedlatency$$
\begin{itemize}
\item here n is the amount of processes
\end{itemize}
\item new vruntime is also calculated according to the niceness:
$$vruntime_i = vruntime_i + \frac{weight_0}{weight_i}*runtime_i$$
\end{itemize}
\end{itemize}
\item Efficiency of CFS (Red-Black Trees)
\label{sec:orgfa1c167}
\begin{itemize}
\item a scheduler has to make decisions as quickly as possible (this should hopefully be scaleable)
\item only runnable processes are kept here
(removed while waiting for I/O)
\item efficiency should be logarithmic (what does that mean?)
\item how does it even work?
\end{itemize}
\end{enumerate}
\part{Computer Scheduling Methods and their Countermeasures}
\label{sec:org853bfc0}
\begin{itemize}
\item Started reading it and it didn't really say anything new. I than scanned over it and skipped the rest
\end{itemize}
\chapter{Classification of Policies}
\label{sec:org3ff5e09}
\section{Characteristics}
\label{sec:org0b5c812}
\begin{itemize}
\item preemptive vs non-preemptive (already mentionned above)
\begin{itemize}
\item preemptive: if a higher priority exists, than the task can and will be abrupted
\item non-preemptive: opposite of preemptive
\end{itemize}
\item resume vs restart
\begin{itemize}
\item if a preempted job ``comes into service again'', should we resume where we left off or should we restart the whole thing?
\end{itemize}
\item where does priority come from?
\begin{itemize}
\item job environment (e.g.: running time, I/O)
\item computer system enviroment (dynamic priorities: e.g.: amount of jobs)
\item users environment (assigned by user)
\end{itemize}
\item knowledge of estimated time until finished
\begin{itemize}
\item most of the computers processes don't have a preset time
\end{itemize}
\end{itemize}
\section{Priority Based on running time only}
\label{sec:org7d8be45}
\begin{itemize}
\item gives shorter jobs and advantage
\end{itemize}
\begin{enumerate}
\item Shortest Job First (SJF)
\label{sec:org1843cbb}
\begin{itemize}
\item it is assumed that we already know the running time at arrival
\item non-preemptive
\item rule only reapplied, when a job is finished (could be also giving back the CPU, while waiting for I/O)
\item better for shorter running jobs, worse for long ones
\end{itemize}
\item Preemptive Shortest Job First
\label{sec:orgfadfd55}
\begin{itemize}
\item it is assumed that we already know the running time at arrival
\item rule reapplied, when a job is finished (+ wait for I/O) or a new job arrives (+ I/O finish)
\item preemptive, resume principle
\item favors the short jobs even more
\item a bit more expensive, because of the context switch
\end{itemize}
\item Round Robin (RR)
\label{sec:org74c80d2}
\begin{itemize}
\item running times not known in advance
\item takes both running and arrival time in consideration
\item cannot make the time quantum too small, because context switch will get too expensive
\item ??? What happens if q = 0???
\item for further info read the heading in OStep/Chapter 7
\end{itemize}
\item Multiple-Level Feedback (FB)
\label{sec:orgb3bef9c}
\begin{itemize}
\item Do not confuse with the modern MLFQ
\item RR but if a task arrives later, it can catch up to the others first
\end{itemize}
\item Two-Level FB / Limited RR
\label{sec:org2737c32}
\begin{itemize}
\item work until a fixed amount of quanta, then put into the background (and only run if no one is in queue 1)
\end{itemize}
\item FB with finite number of levels
\label{sec:org387ecab}
\begin{itemize}
\item just Two-level FB but with a parameter that tells how many queues exist
\item => this gets pretty similar to modern MLFQ
\end{itemize}
\end{enumerate}
\part{Lottery/Stride Scheduling}
\label{sec:org71e0e31}
\part{Adjusting Parameters using Machine Learning}
\label{sec:orgc617b79}
\part{Examples}
\label{sec:orgb85a0df}
\chapter{Linux 2.6 Fair Scheduler}
\label{sec:org0cc4a49}
\chapter{Solaris Scheduling}
\label{sec:orgfa6e5fa}
\end{document}

### 2024-10-09 Handout/article.tex
some refactoring

\documentclass{report}

\usepackage{titlesec}
\usepackage{titling}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage[normalem]{ulem}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[framemethod=TikZ]{mdframed}
\usepackage{etoolbox}

% sets up mdframed
\newcounter{defi}[section]\setcounter{defi}{0}
\renewcommand{\thedefi}{\arabic{section}.\arabic{defi}}
\newenvironment{defi}[2][]{%
\refstepcounter{defi}%
\ifstrempty{#1}%
{\mdfsetup{%
frametitle={%
\tikz[baseline=(current bounding box.east),outer sep=0pt]
\node[anchor=east,rectangle,fill=blue!20]
{\strut Definition~\thedefi};}}
}%
{\mdfsetup{%
frametitle={%
\tikz[baseline=(current bounding box.east),outer sep=0pt]
\node[anchor=east,rectangle,fill=blue!20]
{\strut Definition~\thedefi:~#1};}}%
}%
\mdfsetup{innertopmargin=10pt,linecolor=blue!20,%
linewidth=2pt,topline=true,%
frametitleaboveskip=\dimexpr-\ht\strutbox\relax
}
\begin{mdframed}[]\relax % Added this line
}{%
\end{mdframed} % Added this line
}

% Example usage of mdframed
% \begin{defi}[Pythagoras' theorem]{defi:pythagoras}
% In a right triangle, the square of the hypotenuse is equal to the sum of the squares % of the catheti.
% $$a^2+b^2=c^2$$
% \end{defi}

%define "squiggly" to make red squiggly underline
\makeatletter
\def\squiggly{\bgroup \markoverwith{\textcolor{red}{\lower3.5\p@\hbox{\sixly \char58}}}\ULon}
\makeatother

% Stuff to edit.
\title{CPU-Scheduling}
\newcommand{\thesubtitle}{The Illusion of Multitasking}
\newcommand{\currentdate}{2024/25}
\author{Mark Krutzler}
\newcommand{\auinstitution}{Kantonschule Im Lee}

\titleformat{\section}
{\huge\bfseries}
{\thesection}
{0.5em}
{}[\titlerule]

\titleformat{\subsection}
{\bfseries\LARGE}
{\thesection}
{0.5em}
{}

\titleformat{\subsubsection}
{\large\bfseries}
{\thesection}
{0.5em}
{}

\renewcommand{\maketitle}{
\begin{center}

{\Huge\bfseries
\thetitle}
\vspace{0.5em}\\
{\LARGE\thesubtitle}

{\rule{0.4\textwidth}{.4pt}}

{\bfseries \theauthor}\\
\textit{\auinstitution \ --- \currentdate}
\end{center}
}

\begin{document}

\begin{titlepage}
\vspace*{\fill}
\centering
\maketitle
\vspace*{\fill}
\end{titlepage}

\tableofcontents

\part{Forewords?}

\chapter*{Introduction}
\addcontentsline{toc}{chapter}{Introduction} % Ensure Sources appears in the TOC

We all use computers.
The modern person couldn't do half of the things that is expected from him without an instance of Windows, MacOs or any other Operating System.We surf the web, write emails, documents and have a video call at the same time.
The question is: "How does CPU achieve all of this?" The things listed are only the parts of the OS that we come in contact with on a daily basis.
What about all the other things hidden under the hood of the graphics environment? If it comes down to doing one thing than it's fine.
The problem arises then we have multiple tasks: The CPU can only do one thing at a time and yet we all multitask on our machines.
This Black Box effect of taking multiple jobs and working on them in a proper order is what I will disect in this "Maturarbeit".

The art of scheduling tasks could seem easy for us humans.
We ?instinctively? estimate the many aspects of the task:
\begin{itemize}
\item Job: Convince our boss to give us a pay rise\\Priority: High\\Time Required: Medium (With a lot of Waiting Time)\\Effort: High (Said "No Way" last week)
\item Job: Walk the Dog\\Priority: High\\Time Required: High\\Effort: Low
\item Job: Go to the Hairdresser\\ Priority: Low (Went last week as well)\\Time Required: High\\Effort: Low
\end{itemize}
These prediction are often highly complex and are based on previous experiences.
We also include factors like: When was the last time I did it?
In addition to all that every human plans differently based on what's important to him/her.

A computer does nothing like that. How could it? What he sees is:
\begin{itemize}
\item Job: Process 1\\Priority: Needs to be Set\\Time Required: ??? (Waiting Time: ???)\\Effort: ???
\end{itemize}
What he knows is when the task arrived and how much time he worked on it.
With this in mind the let's get to a quick introduction of the metrics, so that we know how to rate the policies that we'll look at.

\chapter*{Metrics}
\addcontentsline{toc}{chapter}{Metrics} % Ensure Sources appears in the TOC

We can measure different aspects of a Scheduler Policy. There is however a differentiation between fairness and performance. It is often a tradeoff between the two. For example the concept of everybody getting a same sized piece of cake is fair. However, it would be faster to give more to the fast eaters than to those who like to talk during eating. In this scenario we can't have both fairness and performance. In reality it is really difficult to predict, who'll eat faster.

\begin{defi}[Turnaround Time]{dfn:turnaround}
The turnaround time is a performance metric. It measures how long a task took to be completed from the time it arrived.
It is calculated as follows:
$$T_{Turnaround} = T_{Completion} - T_{Arrival}$$
\vspace{1em}
\end{defi}

We'll often look at the Average Turnaround time for a predetermined set of jobs.
This helps us to quantify how "efficient" a single task is handeled.
In the best case scenario a single jobs turnaround time is equivalent to the runtime of that job.
Another really important thing for us is Response Time.
We want to move our mouse around.
If we type we want the letter to appear in an instant and the Microsoft suite should start without much delay.
Of course the startup time of a program depends on many other factors, however if we never even start to work on it, it'll take a good while.

\begin{defi}[Response Time]{dfn:response}
The response time is a performance metric. It measures the responsiveness of our computer. How much time does it take until the job is run?
It is calculated as follows:
$$T_{Response} = T_{First\_Run} - T_{Arrival}$$
\vspace{1em}
\end{defi}

\part{The ABCs of CPU Scheduling}

\chapter{Basic Algorithms}

\section{Optimized for Performance}

\subsection{First In, First Out (FIFO)}

\subsection{Shortest Job First (SJF)}

\subsection{Shortest Time-to-Completion (STCF)}

\section{Optimized for Fairness}

\subsection{Round Robin (RR)}

\part{The Industry Standard of CPU-Scheduling}

\chapter{Multi-Level Feedback Queue}

\chapter{Lottery and Stride Scheduling}

\part{Examples}

\chapter{My Implementation}

\chapter{Solaris Scheduling}

\chapter{Linux 2.6 Fair Scheduler}

\part{Conclusion}

\chapter*{Sources}
\addcontentsline{toc}{chapter}{Sources} % Ensure Sources appears in the TOC

\begin{itemize}
    \item \url{https://texblog.org/2015/09/30/fancy-boxes-for-theorem-lemma-and-proof-with-mdframed/}
\end{itemize}

\end{document}

### 2024-10-17 Handout/article.tex
Wrote FIFO Chapter

@@
 \usepackage[normalem]{ulem}
-\usepackage{xcolor}
+\usepackage[dvipsnames]{xcolor}
 \usepackage{graphicx}
@@
 \usepackage{etoolbox}
+\usepackage{listings}
 
@@
 % \end{defi}
+
+% setup listings for code blocks
+% -- colors
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolor}{rgb}{0.95,0.95,0.92}
+
+%custom style:
+\lsdefinestyle{codeblock}{
+  backgroundcolor=\color{backcolor},
+  commentstyle=\color{codepurple},
+  keywordstyle=\color{NavyBlue},
+  numberstyle=\tiny\color{codegray},
+  stringstyle=\color{codepurple},
+  basicstyle=\ttfamily\footnotesize\bfseries,
+  breakatwhitespace=false,
+  breaklines=true,
+  captionpos=t,
+  keepspaces=true,
+  numbers=left,
+  numbersep=5pt,
+  showspaces=false,
+  showstringspaces=false,
+  showtabs=false,
+  tabsize=2,
+}
+
+\lstset{
+  style=codeblock,
+  framexleftmargin=3.5mm,
+  frame=box
+  rulesepcolor=\color{black},
+  linewidth=0.6\linewidth,
+  xleftmargin=12pt,
+  aboveskip=12pt,
+  belowskip=12pt
+}
 
@@
 We all use computers.
-The modern person couldn't do half of the things that is expected from him without an instance of Windows, MacOs or any other Operating System.We surf the web, write emails, documents and have a video call at the same time.
+The modern person couldn't do half of the things that is expected from him without an instance of Windows, MacOs or any other Operating System.
+We surf the web, write emails, documents and have a video call at the same time.
 The question is: "How does CPU achieve all of this?" The things listed are only the parts of the OS that we come in contact with on a daily basis.
-What about all the other things hidden under the hood of the graphics environment? If it comes down to doing one thing than it's fine.
+What about all the other things hidden under the hood of a graphics environment? If it comes down to doing one thing than it's fine.
 The problem arises then we have multiple tasks: The CPU can only do one thing at a time and yet we all multitask on our machines.
@@
 \item Job: Convince our boss to give us a pay rise\\Priority: High\\Time Required: Medium (With a lot of Waiting Time)\\Effort: High (Said "No Way" last week)
-\item Job: Walk the Dog\\Priority: High\\Time Required: High\\Effort: Low
+\item Job: Walk the Dog\\Priority: High\\Time Required: Low\\Effort: Low
 \item Job: Go to the Hairdresser\\ Priority: Low (Went last week as well)\\Time Required: High\\Effort: Low
@@
 In addition to all that every human plans differently based on what's important to him/her.
+Still from just the priority, time required and effort, we would roughly know how to order these things.
+For example I would walk the dog first, because it is a "request" made by someone else and therefore the response time should be as low as possible.
+The Hairdresser can fade into the background, due to its low priority. Maybe making an appointment is a good idea though.
 
@@
 \addcontentsline{toc}{chapter}{Metrics} % Ensure Sources appears in the TOC
-
+MAYBE JUST LEAVING THIS OUT AND LOOKING AT THIS DURING A LATER TIME (WHEN THE METRICS ARE MENTIONNED OR LINK AS A FOOTNOTE?)
+ALSO THIS FOLLOWS THE OSTEP WAY TOO MUCH:
 We can measure different aspects of a Scheduler Policy.
@@
 
+MAYBE HERE THE METRICS PART?
+
 \chapter{Basic Algorithms}
 
-\section{Optimized for Performance}
+This chapter mainly focuses on simple policies that can be used to order processes.
+Usually there are multiple simplifications added to the scenario so that it makes sense.
+The policies are generally either an abstraction for though experiments or a part of a larger system.
+Therefore there will be an absense of answers to common questions like: "How do we know how long a process takes until completion?"
+(Actually we actually never really know how long something will take)
 
-\subsection{First In, First Out (FIFO)}
+\section{First In, First Out (FIFO)}
 
-\subsection{Shortest Job First (SJF)}
+Probably the most basic of them all.
+The "First In, First Our" is just as the name suggests it.
+Usually we humans know it as first come, first served.
+This policies runs into problems really quickly if the discard the possibility of all tasks requiring the same amount of run time.
+A so called "Convoy Effect" takes place.
+Just imagen how annoying it is, when you're at a store and a family of five cuts into the line in front of you.
+They have two full carts loaded.
+You have to wait until they finish.
+In addition to that they also send back their kid to get something.
+Wouldn't it be overall much faster if they let you in front, before the cashier starts scanning the items?
 
-\subsection{Shortest Time-to-Completion (STCF)}
+\begin{lstlisting}[language=Python, title={FIFO Implementation in Python}]
+# Simple implementation of FIFO in Python
+list = []
 
-\section{Optimized for Fairness}
+# Adding a new process
+def add_client(name):
+  list.append(name)
 
-\subsection{Round Robin (RR)}
+# Decide which process is next
+def allocate():
+  next = list.pop(0)
+  use_resource(next) # arbitrary function that will allow "next" to use the CPU
+\end{lstlisting}
 
-\part{The Industry Standard of CPU-Scheduling}
+As you can see the implementation is itself pretty easy.
+
+\section{Shortest Job First (SJF)}
+
+\section{Shortest Time-to-Completion (STCF)}
+
+\section{Round Robin (RR)}
+
+\part{Advanced Policies}
 
@@
 
-\part{Examples}
+\part{Implementations}
 
-\chapter{My Implementation}
+\chapter{My Implementation?}
 

### 2024-10-17 Handout/article.tex
fixing the formatting of the code

@@
 % -- colors
-\definecolor{codegreen}{rgb}{0,0.6,0}
-\definecolor{codegray}{rgb}{0.5,0.5,0.5}
-\definecolor{codepurple}{rgb}{0.58,0,0.82}
-\definecolor{backcolor}{rgb}{0.95,0.95,0.92}
+% Define a custom color
+\usepackage{color}
+\definecolor{darkred}{rgb}{0.6,0.0,0.0}
+\definecolor{darkgreen}{rgb}{0,0.50,0}
+\definecolor{lightblue}{rgb}{0.0,0.42,0.91}
+\definecolor{orange}{rgb}{0.99,0.48,0.13}
+\definecolor{grass}{rgb}{0.18,0.80,0.18}
+\definecolor{pink}{rgb}{0.97,0.15,0.45}
 
 %custom style:
-\lsdefinestyle{codeblock}{
-  backgroundcolor=\color{backcolor},
-  commentstyle=\color{codepurple},
-  keywordstyle=\color{NavyBlue},
-  numberstyle=\tiny\color{codegray},
-  stringstyle=\color{codepurple},
-  basicstyle=\ttfamily\footnotesize\bfseries,
-  breakatwhitespace=false,
+\lstset{
+  aboveskip=1em,
   breaklines=true,
-  captionpos=t,
-  keepspaces=true,
+  abovecaptionskip=-6pt,
+  captionpos=b,
+  escapeinside={\%*}{*)},
+  frame=single,
   numbers=left,
-  numbersep=5pt,
-  showspaces=false,
-  showstringspaces=false,
-  showtabs=false,
-  tabsize=2,
+  numbersep=15pt,
+  numberstyle=\tiny,
 }
-
-\lstset{
-  style=codeblock,
-  framexleftmargin=3.5mm,
-  frame=box
-  rulesepcolor=\color{black},
-  linewidth=0.6\linewidth,
-  xleftmargin=12pt,
-  aboveskip=12pt,
-  belowskip=12pt
+% 0. Basic Color Theme
+\lstdefinestyle{colored}{ %
+  basicstyle=\ttfamily,
+  backgroundcolor=\color{white},
+  commentstyle=\color{green}\itshape,
+  keywordstyle=\color{blue}\bfseries\itshape,
+  stringstyle=\color{red},
+}
+% 1. General Python Keywords List
+\lstdefinelanguage{PythonPlus}[]{Python}{
+  morekeywords=[1]{,as,assert,nonlocal,with,yield,self,True,False,None,} % Python builtin
+  morekeywords=[2]{,__init__,__add__,__mul__,__div__,__sub__,__call__,__getitem__,__setitem__,__eq__,__ne__,__nonzero__,__rmul__,__radd__,__repr__,__str__,__get__,__truediv__,__pow__,__name__,__future__,__all__,}, % magic methods
+  morekeywords=[3]{,object,type,isinstance,copy,deepcopy,zip,enumerate,reversed,list,set,len,dict,tuple,range,xrange,append,execfile,real,imag,reduce,str,repr,}, % common functions
+  morekeywords=[4]{,Exception,NameError,IndexError,SyntaxError,TypeError,ValueError,OverflowError,ZeroDivisionError,}, % errors
+  morekeywords=[5]{,ode,fsolve,sqrt,exp,sin,cos,arctan,arctan2,arccos,pi, array,norm,solve,dot,arange,isscalar,max,sum,flatten,shape,reshape,find,any,all,abs,plot,linspace,legend,quad,polyval,polyfit,hstack,concatenate,vstack,column_stack,empty,zeros,ones,rand,vander,grid,pcolor,eig,eigs,eigvals,svd,qr,tan,det,logspace,roll,min,mean,cumsum,cumprod,diff,vectorize,lstsq,cla,eye,xlabel,ylabel,squeeze,}, % numpy / math
+}
+% 2. New Language based on Python
+\lstdefinelanguage{PyBrIM}[]{PythonPlus}{
+  emph={d,E,a,Fc28,Fy,Fu,D,des,supplier,Material,Rectangle,PyElmt},
+}
+% 3. Extended theme
+\lstdefinestyle{colorEX}{
+  basicstyle=\ttfamily,
+  backgroundcolor=\color{white},
+  commentstyle=\color{darkgreen}\slshape,
+  keywordstyle=\color{blue}\bfseries\itshape,
+  keywordstyle=[2]\color{blue}\bfseries,
+  keywordstyle=[3]\color{grass},
+  keywordstyle=[4]\color{red},
+  keywordstyle=[5]\color{orange},
+  stringstyle=\color{darkred},
+  emphstyle=\color{pink}\underbar,
 }
@@
 
-\begin{lstlisting}[language=Python, title={FIFO Implementation in Python}]
+\begin{lstlisting}[language=Python, style=colorEX]
 # Simple implementation of FIFO in Python
@@
   next = list.pop(0)
-  use_resource(next) # arbitrary function that will allow "next" to use the CPU
+  use_resource(next) # function that will allow "next" to use the CPU
 \end{lstlisting}

### 2024-10-18 Handout/article.tex
added SJF to .tex

@@
 Just imagen how annoying it is, when you're at a store and a family of five cuts into the line in front of you. 
-They have two full carts loaded.
-You have to wait until they finish.
+They have two full carts loaded. You have to wait until they finish. 
 In addition to that they also send back their kid to get something. 
@@
 # Adding a new process
-def add_client(name):
-  list.append(name)
+def add_client(client):
+  list.append(client)
 
@@
 As you can see the implementation is itself pretty easy.
+\pagebreak
 
 \section{Shortest Job First (SJF)}
+
+Now the situation changes a bit. 
+The infamous family of five wants to go first, however the cashier picks you first after seeing, that you only have a few items.
+This is the "Shortest Job First" (SJF).
+
+\begin{lstlisting}[language=Python, style=colorEX]
+# Simple implementation of SJF in Python
+
+# Define client class and list
+class Client:
+  def __init__(self, ttc): # ttc as time to completion
+    self.ttc = ttc
+    # other stuff that is relevant for the client class
+
+list = []
+
+# Adding a new process
+client = Client(ttc) 
+def add_client(client): 
+  list.append(client)
+
+# Decide which process is next
+def allocate():
+  list.sort(key=lambda a: a.ttc) # sort according to ttc
+  next = list.pop(0)
+  use_resource(next) # function that will allow "next" to use the CPU
+\end{lstlisting}
+
+There are two notable changes in this code snippet, when compared to FIFO. One of them is relevant.
+First, the less exiting part is that I actually took my time to write out some of the client class definition.
+Second and the real difference to FIFO is that we sort the according to the time to completion. 
+This means that we will choose the task with the shortest time to completion.
+AT THIS PART THE KNOWLEDGE OF PROPER SORTING ALGORITHMS IS MISSING: MAYBE A REVERSE BUBBLE SORT? 
+THE LIST SHOULD BE ALREADY SOMEWHAT ORDERED AND THE LAST MEMBER HAS TO BE SORTED
+The only problem is when you arrive late. The scanning of products already started and you'll have to wait.
+This policy is a so called non-preemptive policy.
+"Non-preemptive" just means that once a task will run until it finishes.
 
@@
 
+TODOS:
+- PROPERLY LABEL IMAGES AND CODE BLOCKS
+
 \end{document}

### 2024-10-18 Handout/article.tex
added stcf/psjf

@@
 They have two full carts loaded. You have to wait until they finish. 
-In addition to that they also send back their kid to get something. 
+In addition to that they also send back their kid to get an extra bottle of ketchup. 
 Wouldn't it be overall much faster if they let you in front, before the cashier starts scanning the items?
@@
 
-\section{Shortest Time-to-Completion (STCF)}
+\section{Shortest Time-to-Completion First(STCF)}
+
+If SJF and FIFO are cousins that STCF is the brother.
+Here we will make SJF "preemptive", meaning that our cashier can now "save" the state of the previously scanned items. 
+Therefore STCF is also called "Preemptive Shortest Job First (PSJF)".
+Although the switching costs some time, you'd have to wait anyways until our little boy returns with the ketchup\footnote{If you don't know what I am talking about read Chapter 1.1 (FIFO)}
+The option of saving a state comes in handy, especially if a program has to wait for I/O.
+Preemptiveness also makes the computer much more response, because now you can really do certain things "at the same time".
+The best part is that we don't even have to change much of the previous code.
+\pagebreak
+\begin{lstlisting}[language=Python, style=colorEX]
+  # Simple implementation of STCF in Python
+
+  # definition of client class and list go here
+  current = Client(ttc)
+  current_runtime = 0
+
+  # Adding a new process
+  def add_client(client): 
+    list.append(client)
+    allocate()
+
+  # Decide which process is next
+  def allocate():
+    save_state() # saves of the current process
+
+    current.ttc -= current_runtime # first update ttc
+    current_runtime = 0
+
+    list.sort(key=lambda a: a.ttc) # sort according to ttc
+    current = list.pop(0)
+
+    use_resource(current) # function that will allow "next" to use the CPU (also tracks runtime)
+\end{lstlisting}
+IDK ABOUT THE CODE; IT IS JUST KINDA HACKED TOGETHER; IT IS DEFINITELY NOT SECURE WHATSOEVER! 
+Whenever a new client "joins", the "allocation" function gets run.
+In addition to that, whenever the process finishes, or it voluntarily gives up the CPU, the same "allocation" function can be run.
 

### 2024-10-18 Handout/article.tex
added rr and finished basics

@@
 The policies are generally either an abstraction for though experiments or a part of a larger system.
-Therefore there will be an absense of answers to common questions like: "How do we know how long a process takes until completion?"
-(Actually we actually never really know how long something will take)
+Therefore there will be an absense of answers to common questions like: "How do we know how long a process takes until completion?" \footnote{Actually we actually never really know how long something will take}
+There are also some example code snippets to each policy.
+These are just so you can understand the algorithm and they don't really work in a real life scenario.
+First of all they can't handle interrupts and therefore don't work in a real system.
+In addition to that Python is a high level language and is absolutely not made to handle low level stuff!
+Later on we'll switch to C++, because it fits better, however a "real" OS is written mostly in C.
 
@@
 class Client:
-  def __init__(self, ttc): # ttc as time to completion
-    self.ttc = ttc
+  def __init__(self, burst): # burst time is the time required for the process to finish
+    self.burst = burst
     # other stuff that is relevant for the client class
@@
 # Adding a new process
-client = Client(ttc) 
+client = Client(burst) 
 def add_client(client): 
@@
 def allocate():
-  list.sort(key=lambda a: a.ttc) # sort according to ttc
+  list.sort(key=lambda a: a.burst) # sort according to burst
   next = list.pop(0)
@@
 First, the less exiting part is that I actually took my time to write out some of the client class definition.
-Second and the real difference to FIFO is that we sort the according to the time to completion. 
+Second and the real difference to FIFO is that we sort the according to the burst time\footnote{Burst time is basically the runtime needed until the process finishes.
+This variable can be only predicted}.
 This means that we will choose the task with the shortest time to completion.
@@
   # definition of client class and list go here
-  current = Client(ttc)
+  current = Client(burst)
   current_runtime = 0
@@
 
-    current.ttc -= current_runtime # first update ttc
+    current.burst -= current_runtime # first update burst
     current_runtime = 0
 
-    list.sort(key=lambda a: a.ttc) # sort according to ttc
+    list.sort(key=lambda a: a.burst) # sort according to burst
     current = list.pop(0)
@@
 
+Now, why don't we just scan one item for everyone? 
+This would mean that, even though slow but progress is being constantly made.
+At this point the item scanning doesn't really work anymore\footnote{You can't scan a fraction of a product, or can you?} but let's just make the so called "quanta" (quantum in singular) even smaller.
+As already mentionned in the previous chapter, the cashier needs some time to "save" the state of the current purchases. 
+This effectively mean that after the quantum shrinks to a certain extent, the "context switching"\footnote{Context switch saves the state of a process so that it can be restored later} would get too expensive.
+The goal is to reach a balance, where we have enough responsibility, but the price for it is not too high.
+Notice how by treating the processes more "fairly" by don't caring about their overall burst time, we get rid of a variable, which can only be predicted and replace it with another one, which is more abstract and global.
+\pagebreak
+\begin{lstlisting}[language=Python, style=colorEX]
+  # Simple implementation of RR in Python
+
+  list = []
+  quantum = 1 # Adjust the time slice here
+
+  def add_client(client):
+    list.append(client)
+
+  def use_resource(client, quanta):
+    # definition of use_resource
+    # now it also takes an amount of quanta that the process can run for
+    # after that it saves the state and returns.
+
+  for client in list:
+    use_resource(client, quantum)
+
+\end{lstlisting}
+
+At this point it is time to advance to move onto more advanced policies.
+
 \part{Advanced Policies}
@@
 \part{Implementations}
-
-\chapter{My Implementation?}
 
@@
     \item \url{https://texblog.org/2015/09/30/fancy-boxes-for-theorem-lemma-and-proof-with-mdframed/}
+    \item \url{https://wiki.osdev.org/Context_Switching}
 \end{itemize}
@@
 - PROPERLY LABEL IMAGES AND CODE BLOCKS
+- MAYBE ADD TABLES AND CHARTS FOR THE BASICS, SO THAT THEY CAN SEE WHEN WHAT RUNS (SIMILAR TO OSTEP)
 

### 2024-10-18 Handout/article.tex
Basics of MLFQ jotted down
THis definitely needs further organization. it is just a draft for now....

@@
 
-\chapter*{Introduction}
-\addcontentsline{toc}{chapter}{Introduction} % Ensure Sources appears in the TOC
+\section{Introduction}
 
@@
 
-\chapter*{Metrics}
-\addcontentsline{toc}{chapter}{Metrics} % Ensure Sources appears in the TOC
+\section{Metrics}
 MAYBE JUST LEAVING THIS OUT AND LOOKING AT THIS DURING A LATER TIME (WHEN THE METRICS ARE MENTIONNED OR LINK AS A FOOTNOTE?)
@@
 The option of saving a state comes in handy, especially if a program has to wait for I/O.
+The state of waiting is usually called "blocked" and the technique of running something else during the task is blocked is called overlapping.
 Preemptiveness also makes the computer much more response, because now you can really do certain things "at the same time".
@@
 
+\section{Introduction and Abstract Rules}
+
+Here we'll look at policies doesn't "just invent".
+These are usually more sophisticated and can be hypothetically used in a real system (at least with some adjustemts). 
+First we'll look at the Multi-Level Feedback Queue, short MLFQ. 
+This is one of the most known ones an is used in Solaris.
+The creator Fernando J. Corbató even recieved a Turing Award for it in 1990.
+
+First I'll write down the abstract rules. These are usually complemented with an example. 
+Please feel free to jump around to get a better grasp of the theory
+THIS IS REALLY BAD!! MAYBE WRITE AN EVOLVING STUFF OR SOMETHING:
+THIS IS NO MANUAL; WHERE YOU JUST LIST OUT THE FEATURES LOL
+The most basic principle of MLFQ are the so called "Priority Queues". These are priorities that a program can belong into.
+They can change over time and these changes reflect the action/state of a program.
+Obviously processes at a higher priority are preferred over processes lower down.
+If two processes are at the same priority then the Round Robin policy will be used.
+The changing is priority is done as followed: every task has a certain amount of "allotment" for each priority.
+If this CPU time is used up, than the process will be moved down once and recieve a new set of allotment.
+Which each process starting in the highest queue, this means that the shorter tasks are generally preferred. 
+To solve the so called "starvation" issue, which essentially means that longer tasks get to the bottom priority and stay there being stuck, because there is no way to move, we'll introduce "priority boosts".
+As the name says this boosts the priority of all tasks into level 1 and therefore gives them a second chance. 
+So as "OSTEP REFERENCE HERE" would summarize it:
+\begin{enumerate}
+  \item If Priority (A) > Priority (B) $\Rightarrow$ A runs \& B doesn't
+  \item If Priority (A) = Priority (B) $\Rightarrow$ A \& B run in RR
+  \item When a job enters the system, it is placed at the highest priority
+  \item Once a job uses up its time allotment at a given level, its priority is reduced
+  \item After some period S, move all the jobs in the system to the topmost queue
+\end{enumerate}
+
+Even though MLFQ introduces multiple Voo-Doo Constant, like the scheduling quantum, amount of queues, period S, allotment size (constant or changing?), we still gain an advantage over just simply unknown constant like the burst time or a predetermined priority and effort.
+These are huge improvements, because this means that we could actually implement it properly and the use case is not just purely a though experiment.
+Now to further improve the understanding of the policy we'll look at some examples.
+
+\section{Examples}
+
+\section{Implementation}
+
 \chapter{Lottery and Stride Scheduling}
 
-\part{Implementations}
+\part{Real Life Usecases}
 
@@
     \item \url{https://wiki.osdev.org/Context_Switching}
+    \item \url{https://en.wikipedia.org/wiki/Fernando_J._Corbató}
+    \item \url{https://github.com/jkoritzinsky/xv6-MLFQ}
 \end{itemize}
 
-TODOS:
-- PROPERLY LABEL IMAGES AND CODE BLOCKS
-- MAYBE ADD TABLES AND CHARTS FOR THE BASICS, SO THAT THEY CAN SEE WHEN WHAT RUNS (SIMILAR TO OSTEP)
-
 \end{document}

### 2024-10-18 Research/MLFQ-Example/kernel/proc.c
added some exmaple implementaion
this is obviously a huge commit. shoutout to the original creators. i "unrepod" them so that they can be uploaded to github

@@ -1,0 +1,557 @@
+#include "types.h"
+#include "defs.h"
+#include "param.h"
+#include "mmu.h"
+#include "x86.h"
+#include "proc.h"
+#include "spinlock.h"
+#include "../include/pstat.h"
+
+struct {
+  struct spinlock lock;
+  struct proc proc[NPROC];
+} ptable;
+
+static struct proc *initproc;
+
+int nextpid = 1;
+extern void forkret(void);
+extern void trapret(void);
+
+extern uint ticks;
+
+static void wakeup1(void *chan);
+
+void
+pinit(void)
+{
+  initlock(&ptable.lock, "ptable");
+}
+
+// Look in the process table for an UNUSED proc.
+// If found, change state to EMBRYO and initialize
+// state required to run in the kernel.
+// Otherwise return 0.
+static struct proc*
+allocproc(void)
+{
+  struct proc *p;
+  char *sp;
+
+  acquire(&ptable.lock);
+  for(p = ptable.proc; p < &ptable.proc[NPROC]; p++)
+    if(p->state == UNUSED)
+      goto found;
+  release(&ptable.lock);
+  return 0;
+
+found:
+  p->state = EMBRYO;
+  p->pid = nextpid++;
+  release(&ptable.lock);
+
+  // Allocate kernel stack if possible.
+  if((p->kstack = kalloc()) == 0){
+    p->state = UNUSED;
+    return 0;
+  }
+  sp = p->kstack + KSTACKSIZE;
+  
+  // Leave room for trap frame.
+  sp -= sizeof *p->tf;
+  p->tf = (struct trapframe*)sp;
+  
+  // Set up new context to start executing at forkret,
+  // which returns to trapret.
+  sp -= 4;
+  *(uint*)sp = (uint)trapret;
+
+  sp -= sizeof *p->context;
+  p->context = (struct context*)sp;
+  memset(p->context, 0, sizeof *p->context);
+  p->context->eip = (uint)forkret;
+  p->priority = 0;
+  p->ticks[0] = p->ticks[1] = p->ticks[2] = p->ticks[3] = 0;
+  p->lastScheduledOnTick = 0;
+  return p;
+}
+
+// Set up first user process.
+void
+userinit(void)
+{
+  struct proc *p;
+  extern char _binary_initcode_start[], _binary_initcode_size[];
+  
+  p = allocproc();
+  acquire(&ptable.lock);
+  initproc = p;
+  if((p->pgdir = setupkvm()) == 0)
+    panic("userinit: out of memory?");
[…]

### 2024-10-18 Research/MLFQ-Example/kernel/proc.h

@@ -1,0 +1,92 @@
+#ifndef _PROC_H_
+#define _PROC_H_
+// Segments in proc->gdt.
+// Also known to bootasm.S and trapasm.S
+#define SEG_KCODE 1  // kernel code
+#define SEG_KDATA 2  // kernel data+stack
+#define SEG_KCPU  3  // kernel per-cpu data
+#define SEG_UCODE 4  // user code
+#define SEG_UDATA 5  // user data+stack
+#define SEG_TSS   6  // this process's task state
+#define NSEGS     7
+
+#include "../include/pstat.h"
+
+// Per-CPU state
+struct cpu {
+  uchar id;                    // Local APIC ID; index into cpus[] below
+  struct context *scheduler;   // swtch() here to enter scheduler
+  struct taskstate ts;         // Used by x86 to find stack for interrupt
+  struct segdesc gdt[NSEGS];   // x86 global descriptor table
+  volatile uint booted;        // Has the CPU started?
+  int ncli;                    // Depth of pushcli nesting.
+  int intena;                  // Were interrupts enabled before pushcli?
+
+  // Cpu-local storage variables; see below
+  struct cpu *cpu;
+  struct proc *proc;           // The currently-running process.
+};
+
+extern struct cpu cpus[NCPU];
+extern int ncpu;
+
+// Per-CPU variables, holding pointers to the
+// current cpu and to the current process.
+// The asm suffix tells gcc to use "%gs:0" to refer to cpu
+// and "%gs:4" to refer to proc.  seginit sets up the
+// %gs segment register so that %gs refers to the memory
+// holding those two variables in the local cpu's struct cpu.
+// This is similar to how thread-local variables are implemented
+// in thread libraries such as Linux pthreads.
+extern struct cpu *cpu asm("%gs:0");       // &cpus[cpunum()]
+extern struct proc *proc asm("%gs:4");     // cpus[cpunum()].proc
+
+// Saved registers for kernel context switches.
+// Don't need to save all the segment registers (%cs, etc),
+// because they are constant across kernel contexts.
+// Don't need to save %eax, %ecx, %edx, because the
+// x86 convention is that the caller has saved them.
+// Contexts are stored at the bottom of the stack they
+// describe; the stack pointer is the address of the context.
+// The layout of the context matches the layout of the stack in swtch.S
+// at the "Switch stacks" comment. Switch doesn't save eip explicitly,
+// but it is on the stack and allocproc() manipulates it.
+struct context {
+  uint edi;
+  uint esi;
[…]

### 2024-10-19 Handout/article.tex
rewrote the mlfq theory chapter & added some stuff

@@
 \usepackage{listings}
+\usepackage{indentfirst}
 
@@
 
-\section{Introduction and Abstract Rules}
+\section{Abstract Rules}
 
+\subsubsection{Introduction}
 Here we'll look at policies doesn't "just invent".
@@
 The creator Fernando J. Corbató even recieved a Turing Award for it in 1990.
+First I'll write down some abstract rules. These are usually complemented with an example. 
+After that there will be a quick implementation of the policy in C++.
+Please feel free to jump around to get a better grasp of the (theory).
 
-First I'll write down the abstract rules. These are usually complemented with an example. 
-Please feel free to jump around to get a better grasp of the theory
-THIS IS REALLY BAD!! MAYBE WRITE AN EVOLVING STUFF OR SOMETHING:
-THIS IS NO MANUAL; WHERE YOU JUST LIST OUT THE FEATURES LOL
+\subsubsection{Basics}
+
 The most basic principle of MLFQ are the so called "Priority Queues". These are priorities that a program can belong into.
 They can change over time and these changes reflect the action/state of a program.
-Obviously processes at a higher priority are preferred over processes lower down.
-If two processes are at the same priority then the Round Robin policy will be used.
-The changing is priority is done as followed: every task has a certain amount of "allotment" for each priority.
+The changing of the priority is done as followed: every task has a certain amount of "allotment" for each priority.
 If this CPU time is used up, than the process will be moved down once and recieve a new set of allotment.
-Which each process starting in the highest queue, this means that the shorter tasks are generally preferred. 
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=0.7\textwidth]{assets/mlfq-diag-1.png}
+    \caption{Basic Overview}
+    \label{mlfq-diag-1}
+  \end{center}
+\end{figure}
+
+As you can see in the figure \ref{mlfq-diag-1}, we have 5 queues. They are represented on the y axis. 
+Although there are usually more than just five, too many queues would just make things more complicated. 
+The x axis represents the time.
+After a while process 1, gets demoted from the current queue. This happens, because it used up it's allotment.
+None the less, process one keeps running, because there is no other competing. 
+However if the add another process 2, than the situation changes.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=0.7\textwidth]{assets/mlfq-diag-2.png}
+    \caption{Introducing Process 2}
+    \label{mlfq-diag-2}
+  \end{center}
+\end{figure}
+
+As you can see processes at a higher priority are preferred over processes lower down. 
+This effectively means that once process 2 joins the comptetition it can monopolize the CPU until they end up on the same level.
+Once they are there Round Robin is used to give them both a fair share of allotment.
+With each process starting in the highest queue, this means that the shorter tasks are generally preferred, because they can finish quickly, before they end up on a lower queue.
+Therefore the system becomes more responsive.
+Keep in mind that if a process gives up it's CPU before the allotment is used up (for example if they wait for I/O), than they will "stay" in the same priority and can use up the rest of the CPU time, before getting demoted.
+Therefore not the actual time spent there matters, but the time spent using the resource.
+As you might have to already noticed poor process 1 didn't get any runtime while process 2 was getting their priority prioperly adjusted. Note: for simplicity there are only 4 queues.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=0.7\textwidth]{assets/mlfq-diag-3.png}
+    \caption{Starvation of Process 1}
+    \label{mlfq-diag-3}
+  \end{center}
+\end{figure}
+
 To solve the so called "starvation" issue, which essentially means that longer tasks get to the bottom priority and stay there being stuck, because there is no way to move, we'll introduce "priority boosts".
-As the name says this boosts the priority of all tasks into level 1 and therefore gives them a second chance. 
-So as "OSTEP REFERENCE HERE" would summarize it:
+This effectively means that after a certain period S, all processes are boosted to queue 1 and therefore it gives everybody a "second chance".
+If the processes finish still in priority 1 or 2, than they deserved that queue.
+If however the processes do not finish than they get promoted together to the bottom queue and everybody recieves a fair share.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=0.7\textwidth]{assets/mlfq-diag-4.png}
+    \caption{Priority Boost}
+    \label{mlfq-diag-4}
+  \end{center}
+\end{figure}
+
+\noindent So as "OSTEP REFERENCE HERE" would summarize it:
 \begin{enumerate}
-  \item If Priority (A) > Priority (B) $\Rightarrow$ A runs \& B doesn't
+  \item If Priority (A) $>$ Priority (B) $\Rightarrow$ A runs \& B doesn't
   \item If Priority (A) = Priority (B) $\Rightarrow$ A \& B run in RR
@@
 
+\subsubsection{Voo-Doo Constants}
+
 Even though MLFQ introduces multiple Voo-Doo Constant, like the scheduling quantum, amount of queues, period S, allotment size (constant or changing?), we still gain an advantage over just simply unknown constant like the burst time or a predetermined priority and effort.
 These are huge improvements, because this means that we could actually implement it properly and the use case is not just purely a though experiment.
-Now to further improve the understanding of the policy we'll look at some examples.
-
-\section{Examples}
+It makes sense to find some sensible defaults for these constants and let the system admin adjust it, if they feel like it.
+One could find these magic numbers through machine learning, however that is out of scope of the topic?.
+LINK THE ONE LINK HERE ABOUT ADJUSTING WITH MACHINE LEARNING
 

### 2024-10-20 Handout/article.tex
created main.py and stuff

@@
 First I'll write down some abstract rules. These are usually complemented with an example. 
-After that there will be a quick implementation of the policy in C++.
+After that there will be a quick implementation of the policy in Python.
 Please feel free to jump around to get a better grasp of the (theory).
@@
 
-\noindent So as "OSTEP REFERENCE HERE" would summarize it:
+\noindent So as "OSTEP REFERENCE HERE" would summarize it: DO I EVEN NEED THIS?
 \begin{enumerate}
@@
 \section{Implementation}
+
+For the whole script, please checkout my Github. There you can find
+
+\begin{lstlisting}[language=Python, style=colorEX]
+    # Implementation of MLFQ in Python
+
+  \end{lstlisting}
 

### 2024-10-20 Implementation/main.py

@@ -1,0 +1,21 @@
+# question: Research/MLFQ-Example/kernel/proc.c: 314
+#           why no else: or case: for everything else why just nothing = break
+
+
+# IMPLEMENT THE BASICS AS WELL!!!!!!!
+
+
+#########################################
+##                                     ##
+##                MLFQ                 ##
+##                                     ##
+#########################################
+
+class Process:
+    def __init__(self, pid, priority, burst_time):
+        self.pid = pid
+        self.priority = priority
+        self.burst_time = burst_time
+
+def print_state():
+    print("testing")

### 2024-10-20 Research/MLFQ-Example/kernel/proc.c

@@ -276,15 +276,15 @@ scheduler(void)
     acquire(&ptable.lock);
 
 
-      // Priority boost
-      if(xticks % 100 == 0){
+    // Priority boost
+    if(xticks % 100 == 0){
         for(j = 0; j < NPROC; ++j){
-          if(j == lastScheduled || ptable.proc[j].state != RUNNABLE) continue;
-          if(ptable.proc[j].priority > 0 && xticks - ptable.proc[j].lastScheduledOnTick >= 100){
-            ptable.proc[j].priority--;
-          }
+            if(j == lastScheduled || ptable.proc[j].state != RUNNABLE) continue;
+            if(ptable.proc[j].priority > 0 && xticks - ptable.proc[j].lastScheduledOnTick >= 100){
+                ptable.proc[j].priority--;
+            }
         }
-      }
+    }
 
     struct proc *procToSchedForPriority[4] = {0};
     for(i = (lastScheduled + 1) % NPROC, j = 0; i != lastScheduled && j < NPROC; i = (i + 1) % NPROC, ++j){

### 2024-10-20 Handout/article.tex
Started cleaning up draft & added basic python structure

@@
 \section{Metrics}
-MAYBE JUST LEAVING THIS OUT AND LOOKING AT THIS DURING A LATER TIME (WHEN THE METRICS ARE MENTIONNED OR LINK AS A FOOTNOTE?)
-ALSO THIS FOLLOWS THE OSTEP WAY TOO MUCH:
+% COMMENT: MAYBE JUST LEAVING THIS OUT AND LOOKING AT THIS DURING A LATER TIME (WHEN THE METRICS ARE MENTIONNED OR LINK AS A FOOTNOTE?)
+% COMMENT: ALSO THIS FOLLOWS THE OSTEP WAY TOO MUCH:
 We can measure different aspects of a Scheduler Policy.
@@
 
-MAYBE HERE THE METRICS PART?
+% COMMENT: MAYBE HERE THE METRICS PART?
 
@@
 
-\begin{lstlisting}[language=Python, style=colorEX]
+\begin{lstlisting}[language=Python, style=colorEX, label=cd:fifo-python]
 # Simple implementation of FIFO in Python
@@
 # Adding a new process
-def add_client(client):
-  list.append(client)
+def add_process(process):
+  list.append(process)
 
@@
 
-\begin{lstlisting}[language=Python, style=colorEX]
+\begin{lstlisting}[language=Python, style=colorEX, label=cd:sjf-python]
 # Simple implementation of SJF in Python
 
-# Define client class and list
-class Client:
+# Define process class and list
+class Process:
   def __init__(self, burst): # burst time is the time required for the process to finish
     self.burst = burst
-    # other stuff that is relevant for the client class
+    # other stuff that is relevant for the process class
 
@@
 # Adding a new process
-client = Client(burst) 
-def add_client(client): 
-  list.append(client)
+process = Process(burst) 
+def add_process(process): 
+  list.append(process)
 
@@
 There are two notable changes in this code snippet, when compared to FIFO. One of them is relevant.
-First, the less exiting part is that I actually took my time to write out some of the client class definition.
+First, the less exiting part is that I actually took my time to write out some of the process class definition.
 Second and the real difference to FIFO is that we sort the according to the burst time\footnote{Burst time is basically the runtime needed until the process finishes.
@@
 This means that we will choose the task with the shortest time to completion.
-AT THIS PART THE KNOWLEDGE OF PROPER SORTING ALGORITHMS IS MISSING: MAYBE A REVERSE BUBBLE SORT? 
-THE LIST SHOULD BE ALREADY SOMEWHAT ORDERED AND THE LAST MEMBER HAS TO BE SORTED
+% COMMENT: AT THIS PART THE KNOWLEDGE OF PROPER SORTING ALGORITHMS IS MISSING: MAYBE A REVERSE BUBBLE SORT? 
+% COMMENT: => Sorting algorithm input?
+% COMMENT: THE LIST SHOULD BE ALREADY SOMEWHAT ORDERED AND THE LAST MEMBER HAS TO BE SORTED
 The only problem is when you arrive late. The scanning of products already started and you'll have to wait.
@@
 \pagebreak
-\begin{lstlisting}[language=Python, style=colorEX]
+\begin{lstlisting}[language=Python, style=colorEX, label=cd:stcf-python]
   # Simple implementation of STCF in Python
 
-  # definition of client class and list go here
-  current = Client(burst)
+  # definition of process class and list go here
+  current = Process(burst)
   current_runtime = 0
@@
   # Adding a new process
-  def add_client(client): 
-    list.append(client)
+  def add_process(process): 
+    list.append(process)
     allocate()
@@
 \end{lstlisting}
-IDK ABOUT THE CODE; IT IS JUST KINDA HACKED TOGETHER; IT IS DEFINITELY NOT SECURE WHATSOEVER! 
-Whenever a new client "joins", the "allocation" function gets run.
+% COMMENT: IDK ABOUT THE CODE; IT IS JUST KINDA HACKED TOGETHER; IT IS DEFINITELY NOT SECURE WHATSOEVER! 
+Whenever a new process "joins", the "allocation" function gets run.
 In addition to that, whenever the process finishes, or it voluntarily gives up the CPU, the same "allocation" function can be run.
@@
 \pagebreak
-\begin{lstlisting}[language=Python, style=colorEX]
+\begin{lstlisting}[language=Python, style=colorEX, label=cd:rr-python]
   # Simple implementation of RR in Python
@@
 
-  def add_client(client):
-    list.append(client)
+  def add_process(process):
+    list.append(process)
 
-  def use_resource(client, quanta):
+  def use_resource(process, quanta):
     # definition of use_resource
@@
 
-  for client in list:
-    use_resource(client, quantum)
+  for process in list:
+    use_resource(process, quantum)
 
@@
 
-\noindent So as "OSTEP REFERENCE HERE" would summarize it: DO I EVEN NEED THIS?
+\noindent So as "OSTEP REFERENCE HERE" would summarize it: % COMMENT: DO I EVEN NEED THIS?
 \begin{enumerate}
@@
 One could find these magic numbers through machine learning, however that is out of scope of the topic?.
-LINK THE ONE LINK HERE ABOUT ADJUSTING WITH MACHINE LEARNING
+% COMMENT: LINK THE ONE LINK HERE ABOUT ADJUSTING WITH MACHINE LEARNING
 
@@
 
-For the whole script, please checkout my Github. There you can find
+For the whole script, please checkout my Github.
 
-\begin{lstlisting}[language=Python, style=colorEX]
+\begin{lstlisting}[language=Python, style=colorEX, label=cd:mlfq-python]
   # Implementation of MLFQ in Python
@@
     \item \url{https://github.com/jkoritzinsky/xv6-MLFQ}
+    \item Obviously this is not complete and the main sources are not even here yet
 \end{itemize}

### 2024-10-20 Implementation/main.py

@@ -1,21 +1,0 @@
-# question: Research/MLFQ-Example/kernel/proc.c: 314
-#           why no else: or case: for everything else why just nothing = break
-
-
-# IMPLEMENT THE BASICS AS WELL!!!!!!!
-
-
-#########################################
-##                                     ##
-##                MLFQ                 ##
-##                                     ##
-#########################################
-
-class Process:
-    def __init__(self, pid, priority, burst_time):
-        self.pid = pid
-        self.priority = priority
-        self.burst_time = burst_time
-
-def print_state():
-    print("testing")

### 2024-10-20 Implementation/mlfq.py

@@ -1,0 +1,71 @@
+# question: Research/MLFQ-Example/kernel/proc.c: 314
+#           why no else: or case: for everything else why just nothing = break
+
+
+#########################################
+#                                       #
+#                 MLFQ                  #
+#                                       #
+#########################################
+
+import sys
+
+
+class CustomError(Exception):
+    def __init__(self, msg):
+        self.msg = msg
+
+
+class Process:
+    def __init__(self, pid, burst_time):
+        # TODO: Limit each of these to max 99
+        self.pid = pid
+        self.priority = 0  # Processes start at Priority 0
+        self.burst_time = burst_time
+
+    def __str__(self):
+        str_pid = makeTwoDigit(self.pid)
+        str_priority = makeTwoDigit(self.priority)
+        str_burst_time = makeTwoDigit(self.burst_time)
+
+        print("|¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯|")
+        print("| Process:    ", str_pid, " |")
+        print("| Priority:   ", str_priority, " |")
+        print("| Burst Time: ", str_burst_time, " |")
+        print("|__________________|")
+
+
+class Queue:
+    def __init__(self, priority):
+        self.priority = priority  # Should be an int
+        self.processes = []  # Keeps track of all processes
+
+    def __str__(self):
+        print("Queue")
+
+
+# Make a number two digit long for better printing
+def makeTwoDigit(number):
+    if number < 10:
+        return "0" + str(number)
+    else:
+        return str(number)
+
+
+def print_state():
+    print("Current State:")
+
+
+def main():
+    process1 = Process(1, 10)
+    print(process1.__str__())
+    print("main")
+
+
+try:
+    main()
+except CustomError as e:
+    print("An Exception occured: ", e.msg)
+except Exception:
+    print("Unexpected error: ", sys.exc_info()[0])
+    raise

### 2024-10-20 Handout/article.tex
2. batch cleaning up

@@
 
+% TODO: Explain why a CPU can only do one operation at the time
 We all use computers.
@@
 The question is: "How does CPU achieve all of this?" The things listed are only the parts of the OS that we come in contact with on a daily basis.
-What about all the other things hidden under the hood of a graphics environment? If it comes down to doing one thing than it's fine.
+What about all the other things hidden under the hood? If it comes down to doing one thing than it's fine.
 The problem arises then we have multiple tasks: The CPU can only do one thing at a time and yet we all multitask on our machines.
@@
 \begin{itemize}
-\item Job: Convince our boss to give us a pay rise\\Priority: High\\Time Required: Medium (With a lot of Waiting Time)\\Effort: High (Said "No Way" last week)
+\item Job: Convince our boss to give us a pay rise\\Priority: High\\Time Required: Medium (With a lot of waiting time)\\Effort: High (Said "No way" last week)
 \item Job: Walk the Dog\\Priority: High\\Time Required: Low\\Effort: Low
@@
 What he knows is when the task arrived and how much time he worked on it.
+The goal of scheduling is somehow not to rely on predictions and still decide wisely.
 With this in mind the let's get to a quick introduction of the metrics, so that we know how to rate the policies that we'll look at.
 
+\pagebreak
 \section{Metrics}
-% COMMENT: MAYBE JUST LEAVING THIS OUT AND LOOKING AT THIS DURING A LATER TIME (WHEN THE METRICS ARE MENTIONNED OR LINK AS A FOOTNOTE?)
-% COMMENT: ALSO THIS FOLLOWS THE OSTEP WAY TOO MUCH:
-We can measure different aspects of a Scheduler Policy.
+% TODO: MAYBE JUST LEAVING THIS OUT AND LOOKING AT THIS DURING A LATER TIME (WHEN THE METRICS ARE MENTIONNED OR LINK AS A FOOTNOTE?)
+% TODO: ALSO THIS FOLLOWS THE OSTEP WAY TOO MUCH:
+We can measure different aspects of a "scheduler policy".
 There is however a differentiation between fairness and performance.
 It is often a tradeoff between the two.
+% TODO: Lateron mention, where it is obvious
 For example the concept of everybody getting a same sized piece of cake is fair.
@@
 If we type we want the letter to appear in an instant and the Microsoft suite should start without much delay.
-Of course the startup time of a program depends on many other factors, however if we never even start to work on it, it'll take a good while.
+Of course the startup time of a program depends on many other factors, however if we never even start working on it, it'll take a good while.
 
@@
 
-% COMMENT: MAYBE HERE THE METRICS PART?
+% TODO: MAYBE HERE THE METRICS PART?
 
@@
 In addition to that Python is a high level language and is absolutely not made to handle low level stuff!
-Later on we'll switch to C++, because it fits better, however a "real" OS is written mostly in C.
 
@@
 Usually we humans know it as first come, first served. 
-This policies runs into problems really quickly if the discard the possibility of all tasks requiring the same amount of run time. 
+This policies runs into problems really quickly if we discard the possibility of all tasks requiring the same amount of run time. 
 A so called "Convoy Effect" takes place. 
@@
 In addition to that they also send back their kid to get an extra bottle of ketchup. 
-Wouldn't it be overall much faster if they let you in front, before the cashier starts scanning the items?
+Wouldn't it be overall much faster if they let you in front before the cashier starts scanning the items?
 
@@
 # Simple implementation of FIFO in Python
-list = []
+queue = []
 
@@
 def add_process(process):
-  list.append(process)
+  queue.append(process)
 
@@
 def allocate():
-  next = list.pop(0)
+  next = queue.pop(0)
   use_resource(next) # function that will allow "next" to use the CPU
@@
 Now the situation changes a bit. 
-The infamous family of five wants to go first, however the cashier picks you first after seeing, that you only have a few items.
+The infamous family of five wants to go first, however the cashier picks you after seeing that you only have a few items.
 This is the "Shortest Job First" (SJF).
@@
 
-# Define process class and list
+# Define process class and queue
 class Process:
@@
 
-list = []
+queue = []
 
@@
 def add_process(process): 
-  list.append(process)
+  queue.append(process)
 
@@
 def allocate():
-  list.sort(key=lambda a: a.burst) # sort according to burst
-  next = list.pop(0)
+  queue.sort(key=lambda a: a.burst) # sort according to burst
+  next = queue.pop(0)
   use_resource(next) # function that will allow "next" to use the CPU
@@
 
-There are two notable changes in this code snippet, when compared to FIFO. One of them is relevant.
+There are two notable changes in this code snippet when compared to FIFO. One of them is relevant.
 First, the less exiting part is that I actually took my time to write out some of the process class definition.
-Second and the real difference to FIFO is that we sort the according to the burst time\footnote{Burst time is basically the runtime needed until the process finishes.
+Second and the real difference to FIFO is that we sort the queue according to the burst time\footnote{Burst time is basically the runtime needed until the process finishes.
 This variable can be only predicted}.
 This means that we will choose the task with the shortest time to completion.
-% COMMENT: AT THIS PART THE KNOWLEDGE OF PROPER SORTING ALGORITHMS IS MISSING: MAYBE A REVERSE BUBBLE SORT? 
-% COMMENT: => Sorting algorithm input?
-% COMMENT: THE LIST SHOULD BE ALREADY SOMEWHAT ORDERED AND THE LAST MEMBER HAS TO BE SORTED
+% TODO: AT THIS PART THE KNOWLEDGE OF PROPER SORTING ALGORITHMS IS MISSING: MAYBE A REVERSE BUBBLE SORT? 
+% TODO: => Sorting algorithm input?
+% TODO: THE LIST SHOULD BE ALREADY SOMEWHAT ORDERED AND THE LAST MEMBER HAS TO BE SORTED
 The only problem is when you arrive late. The scanning of products already started and you'll have to wait.
@@
 Therefore STCF is also called "Preemptive Shortest Job First (PSJF)".
-Although the switching costs some time, you'd have to wait anyways until our little boy returns with the ketchup\footnote{If you don't know what I am talking about read Chapter 1.1 (FIFO)}
+Although the switching costs some time it does not matter here, because one would have to wait until the little boy returns with the ketchup\footnote{If you don't know what I am talking about read Chapter 1.1 (FIFO)}
+% TODO: Pretty change perspective switch, should revisit it later
 The option of saving a state comes in handy, especially if a program has to wait for I/O.
@@
 
-  # definition of process class and list go here
+  # definition of process class and queue go here
   current = Process(burst)
@@
   def add_process(process): 
-    list.append(process)
+    queue.append(process)
     allocate()
@@
 
-    list.sort(key=lambda a: a.burst) # sort according to burst
-    current = list.pop(0)
+    queue.sort(key=lambda a: a.burst) # sort according to burst
+    current = queue.pop(0)
 
@@
 \end{lstlisting}
-% COMMENT: IDK ABOUT THE CODE; IT IS JUST KINDA HACKED TOGETHER; IT IS DEFINITELY NOT SECURE WHATSOEVER! 
+% TODO: IDK ABOUT THE CODE; IT IS JUST KINDA HACKED TOGETHER; IT IS DEFINITELY NOT SECURE WHATSOEVER! Also should explain it more in words!
+% TODO: GENErally code is trash. should do a more RR style, with after every allotment new scheduling
 Whenever a new process "joins", the "allocation" function gets run.
@@
 Now, why don't we just scan one item for everyone? 
+% TODO: Exlain how that would make it better...
 This would mean that, even though slow but progress is being constantly made.
-At this point the item scanning doesn't really work anymore\footnote{You can't scan a fraction of a product, or can you?} but let's just make the so called "quanta" (quantum in singular) even smaller.
+Even if at this point the item scanning doesn't really work anymore\footnote{You can't scan a fraction of a product, or can you?}, let's just make the so called "quanta" (quantum in singular) even smaller.
+% TODO: Explain term quantum better
 As already mentionned in the previous chapter, the cashier needs some time to "save" the state of the current purchases. 
@@
 
-  list = []
+  queue = []
   quantum = 1 # Adjust the time slice here
@@
   def add_process(process):
-    list.append(process)
+    queue.append(process)
 
@@
 
-  for process in list:
+while true:
+  for process in queue:
     use_resource(process, quantum)
@@
 \subsubsection{Introduction}
-Here we'll look at policies doesn't "just invent".
+Here we'll look at policies that one doesn't "just invent".
 These are usually more sophisticated and can be hypothetically used in a real system (at least with some adjustemts). 
 First we'll look at the Multi-Level Feedback Queue, short MLFQ. 
-This is one of the most known ones an is used in Solaris.
+This is one of the most known ones and is used in Solaris.
 The creator Fernando J. Corbató even recieved a Turing Award for it in 1990.
@@
 The most basic principle of MLFQ are the so called "Priority Queues". These are priorities that a program can belong into.
+% TODO: SIMILAR TO PREVIOUS QUEUES??? MAKE A CONNECTINjjjjkk
 They can change over time and these changes reflect the action/state of a program.
@@
 After a while process 1, gets demoted from the current queue. This happens, because it used up it's allotment.
-None the less, process one keeps running, because there is no other competing. 
+None the less, process one keeps running, because there is no one other competing. 
 However if the add another process 2, than the situation changes.
@@
 Once they are there Round Robin is used to give them both a fair share of allotment.
-With each process starting in the highest queue, this means that the shorter tasks are generally preferred, because they can finish quickly, before they end up on a lower queue.
+With each process starting in the highest queue, this means that the shorter tasks are generally preferred, because they can finish quickly before they end up on a lower queue.
 Therefore the system becomes more responsive.
 Keep in mind that if a process gives up it's CPU before the allotment is used up (for example if they wait for I/O), than they will "stay" in the same priority and can use up the rest of the CPU time, before getting demoted.
+% TODO: INSERT WATING TIEM EXAMPLE HERE
 Therefore not the actual time spent there matters, but the time spent using the resource.
-As you might have to already noticed poor process 1 didn't get any runtime while process 2 was getting their priority prioperly adjusted. Note: for simplicity there are only 4 queues.
+As you might have to already noticed poor process 1 didn't get any runtime while process 2 was getting their priority prioperly adjusted. Note: for simplicity there are now only 4 queues.
 
@@
 
-\noindent So as "OSTEP REFERENCE HERE" would summarize it: % COMMENT: DO I EVEN NEED THIS?
+\noindent So as "OSTEP REFERENCE HERE" would summarize it: % TODO: DO I EVEN NEED THIS?
 \begin{enumerate}
@@
 
-Even though MLFQ introduces multiple Voo-Doo Constant, like the scheduling quantum, amount of queues, period S, allotment size (constant or changing?), we still gain an advantage over just simply unknown constant like the burst time or a predetermined priority and effort.
+Even though MLFQ introduces multiple Voo-Doo Constant, like the scheduling quantum, amount of queues, period S, allotment size (constant or changing?), % TODO: EXPALIN CONSTAND OR CHANGIN
+we still gain an advantage over just simply unknown constant like the burst time or a predetermined priority and effort.
 These are huge improvements, because this means that we could actually implement it properly and the use case is not just purely a though experiment.
@@
 One could find these magic numbers through machine learning, however that is out of scope of the topic?.
-% COMMENT: LINK THE ONE LINK HERE ABOUT ADJUSTING WITH MACHINE LEARNING
-
+% TODO: LINK THE ONE LINK HERE ABOUT ADJUSTING WITH MACHINE LEARNING
+% TODO: ADD AN EXAMPLE WITH I/O WAITNG TIME AND EARLY FINISH
 \section{Implementation}

### 2024-10-20 Implementation/mlfq.py

@@ -9,30 +9,12 @@
 #########################################
 
 import sys
+import process
 
 
 class CustomError(Exception):
     def __init__(self, msg):
         self.msg = msg
-
-
-class Process:
-    def __init__(self, pid, burst_time):
-        # TODO: Limit each of these to max 99
-        self.pid = pid
-        self.priority = 0  # Processes start at Priority 0
-        self.burst_time = burst_time
-
-    def __str__(self):
-        str_pid = makeTwoDigit(self.pid)
-        str_priority = makeTwoDigit(self.priority)
-        str_burst_time = makeTwoDigit(self.burst_time)
-
-        print("|¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯|")
-        print("| Process:    ", str_pid, " |")
-        print("| Priority:   ", str_priority, " |")
-        print("| Burst Time: ", str_burst_time, " |")
-        print("|__________________|")
 
 
 class Queue:
@@ -57,7 +39,7 @@ def print_state():
 
 
 def main():
-    process1 = Process(1, 10)
+    process1 = process.Process(1, 10)
     print(process1.__str__())
     print("main")
 

### 2024-10-20 Implementation/process.py

@@ -1,0 +1,25 @@
+class Process:
+    def __init__(self, pid, burst_time):
+        # TODO: Limit each of these to max 99
+        self.pid = pid
+        self.priority = 0  # Processes start at Priority 0
+        self.burst_time = burst_time
+
+    def __str__(self):
+        str_pid = makeTwoDigit(self.pid)
+        str_priority = makeTwoDigit(self.priority)
+        str_burst_time = makeTwoDigit(self.burst_time)
+
+        print("|¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯¯|")
+        print("| Process:    ", str_pid, " |")
+        print("| Priority:   ", str_priority, " |")
+        print("| Burst Time: ", str_burst_time, " |")
+        print("|__________________|")
+
+
+# Make a number two digit long for better printing
+def makeTwoDigit(number):
+    if number < 10:
+        return "0" + str(number)
+    else:
+        return str(number)

### 2024-11-09 Implementation/mlfq.py
modified python implementation

@@ -22,8 +22,12 @@ def mlfq(queues):
     for queue_index in range(len(queues)-1):
         while queues[queue_index].processes != []:
             queues[queue_index].__str__()
+            inp = input("press enter to continue")
             for process in queues[queue_index].processes:
+                process.allotment += quantum
+                process.runtime += quantum
                 if process.is_finished() == True:
+                    print("process", process.pid, "finished")
                     queues[queue_index].remove_process(process)
                     continue
                 else:
@@ -32,8 +36,7 @@ def mlfq(queues):
                         queues[queue_index].remove_process(process)
                         queues[queue_index + 1].add_process(process)
                     else:
-                        process.allotment += quantum
-                        process.runtime += quantum
+                        continue
 
 
                     
@@ -49,7 +52,7 @@ if __name__ == "__main__":
         i = 0
         for queue in queues:
             i += 1
-            queue.add_process(process.Process(i,i*2+3))
+            queue.add_process(process.Process(i,i*4+3))
         mlfq(queues)
     except CustomError as e:
         print("An Exception occured: ", e.msg)

### 2024-11-09 Implementation/queue_list.py

@@ -12,6 +12,7 @@ class Queue:
         print("|                  |")
         for process in self.processes:
             process.__str__()
+        print("\n")
     
     def add_process(self, process):
         process.priority = self.priority

### 2024-11-11 Archive/Handout/article.tex
read README 2024-11-11

--- 2024-10-20 Handout/article.tex
@@
 
-% TODO: Explain why a CPU can only do one operation at the time
 We all use computers.
@@
 The question is: "How does CPU achieve all of this?" The things listed are only the parts of the OS that we come in contact with on a daily basis.
-What about all the other things hidden under the hood? If it comes down to doing one thing than it's fine.
-The problem arises then we have multiple tasks: The CPU can only do one thing at a time and yet we all multitask on our machines.
+What about all the other things hidden under the hood?
+If simplified, a simple CPU is like a simple production line. It does one things at a time.
+Therefore a problem arises then we have multiple tasks: The CPU can only do one thing at a time\footnote{When talking about a single-core CPU for simplification} and yet we all multitask on our machines.
 This Black Box effect of taking multiple jobs and working on them in a proper order is what I will disect in this "Maturarbeit".
@@
 
+% TODO: WAY TOO DEEP, BECAUSE WILL NEVER REALLY USE IT AGAIN THAT DEEPLY (EXCEPT IF A PROPER EXPLANATION IS DONE FOR EACH POLICY)
 \begin{defi}[Turnaround Time]{dfn:turnaround}
@@
 \part{The ABCs of CPU Scheduling}
-
-% TODO: MAYBE HERE THE METRICS PART?
 
@@
 This means that we will choose the task with the shortest time to completion.
-% TODO: AT THIS PART THE KNOWLEDGE OF PROPER SORTING ALGORITHMS IS MISSING: MAYBE A REVERSE BUBBLE SORT? 
-% TODO: => Sorting algorithm input?
-% TODO: THE LIST SHOULD BE ALREADY SOMEWHAT ORDERED AND THE LAST MEMBER HAS TO BE SORTED
-The only problem is when you arrive late. The scanning of products already started and you'll have to wait.
+At this point one could go into which sort algorithm to use and which one is the most efficient, however that is not something I will touch upon here.
+The reason is that I don't think that it is necessary to understand those in order to grasp the concept of SJF.
+
+Back to our story, now you only have a problem when you arrive late. The scanning of products has already started and you'll have to wait.
 This policy is a so called non-preemptive policy.
@@
 
-If SJF and FIFO are cousins that STCF is the brother.
-Here we will make SJF "preemptive", meaning that our cashier can now "save" the state of the previously scanned items. 
+If SJF and FIFO are cousins than STCF is the brother.
+To fix the issue from before, we'll make SJF "preemptive", meaning that our cashier can now "save" the state of the previously scanned items. 
 Therefore STCF is also called "Preemptive Shortest Job First (PSJF)".
-Although the switching costs some time it does not matter here, because one would have to wait until the little boy returns with the ketchup\footnote{If you don't know what I am talking about read Chapter 1.1 (FIFO)}
-% TODO: Pretty change perspective switch, should revisit it later
 The option of saving a state comes in handy, especially if a program has to wait for I/O.
+Although the switching costs some time, this can be disregarded, because one would have to wait until the little boy returns with the ketchup\footnote{If you don't know what I am talking about read Chapter 1.1 (FIFO)} anyways.
 The state of waiting is usually called "blocked" and the technique of running something else during the task is blocked is called overlapping.
-Preemptiveness also makes the computer much more response, because now you can really do certain things "at the same time".
+% TODO: explain blocked and overlapping with the cashier in a better way
+Preemptiveness also makes the computer much more responsive, because now you can really do certain things "at the same time".
 The best part is that we don't even have to change much of the previous code.
@@
 \end{lstlisting}
-% TODO: IDK ABOUT THE CODE; IT IS JUST KINDA HACKED TOGETHER; IT IS DEFINITELY NOT SECURE WHATSOEVER! Also should explain it more in words!
-% TODO: GENErally code is trash. should do a more RR style, with after every allotment new scheduling
 Whenever a new process "joins", the "allocation" function gets run.
@@
 Now, why don't we just scan one item for everyone? 
-% TODO: Exlain how that would make it better...
-This would mean that, even though slow but progress is being constantly made.
-Even if at this point the item scanning doesn't really work anymore\footnote{You can't scan a fraction of a product, or can you?}, let's just make the so called "quanta" (quantum in singular) even smaller.
+This could mean that every one sees progress and that same sized jobs finish at almost the same time
+Even if at this point the item scanning explanation doesn't really work anymore\footnote{You can't scan a fraction of a product, or can you?}, let's just make the so called "quanta" (quantum in singular) even smaller.
 % TODO: Explain term quantum better
@@
 The most basic principle of MLFQ are the so called "Priority Queues". These are priorities that a program can belong into.
-% TODO: SIMILAR TO PREVIOUS QUEUES??? MAKE A CONNECTINjjjjkk
 They can change over time and these changes reflect the action/state of a program.
@@
 If this CPU time is used up, than the process will be moved down once and recieve a new set of allotment.
+This way you could have a fast lane checkout, where you can only scan a few items.
+A separation of small and bulk purchases would occur making the experience overall faster.
+There is still only one cashier though, so there are some pitfall that one must avoid.
 
@@
 Keep in mind that if a process gives up it's CPU before the allotment is used up (for example if they wait for I/O), than they will "stay" in the same priority and can use up the rest of the CPU time, before getting demoted.
-% TODO: INSERT WATING TIEM EXAMPLE HERE
+% TODO: INSERT WATING TIEM EXAMPLE HERE meaning a sketch
 Therefore not the actual time spent there matters, but the time spent using the resource.
@@
 
-\noindent So as "OSTEP REFERENCE HERE" would summarize it: % TODO: DO I EVEN NEED THIS?
+\noindent So as "OSTEP REFERENCE HERE" would summarize it: % TODO: find and insert reference
 \begin{enumerate}

### 2024-11-11 Research/MLFQ-Example/include/pstat.h

@@ -1,17 +1,0 @@
-#ifndef _PSTAT_H_
-#define _PSTAT_H_
-
-#include "param.h"
-
-enum procstate { UNUSED, EMBRYO, SLEEPING, RUNNABLE, RUNNING, ZOMBIE };
-
-struct pstat {
-  int inuse[NPROC]; // whether this slot of the process table is in use (1 or 0)
-  int pid[NPROC];   // PID of each process
-  int priority[NPROC];  // current priority level of each process (0-3)
-  enum procstate state[NPROC];  // current state (e.g., SLEEPING or RUNNABLE) of each process
-  int ticks[NPROC][4];  // number of ticks each process has accumulated at each of 4 priorities
-};
-
-
-#endif // _PSTAT_H_

### 2024-11-11 Research/MLFQ-Example/kernel/proc.c

@@ -1,557 +1,0 @@
-#include "types.h"
-#include "defs.h"
-#include "param.h"
-#include "mmu.h"
-#include "x86.h"
-#include "proc.h"
-#include "spinlock.h"
-#include "../include/pstat.h"
-
-struct {
-  struct spinlock lock;
-  struct proc proc[NPROC];
-} ptable;
-
-static struct proc *initproc;
-
-int nextpid = 1;
-extern void forkret(void);
-extern void trapret(void);
-
-extern uint ticks;
-
-static void wakeup1(void *chan);
-
-void
-pinit(void)
-{
-  initlock(&ptable.lock, "ptable");
-}
-
-// Look in the process table for an UNUSED proc.
-// If found, change state to EMBRYO and initialize
-// state required to run in the kernel.
-// Otherwise return 0.
-static struct proc*
-allocproc(void)
-{
-  struct proc *p;
-  char *sp;
-
-  acquire(&ptable.lock);
-  for(p = ptable.proc; p < &ptable.proc[NPROC]; p++)
-    if(p->state == UNUSED)
-      goto found;
-  release(&ptable.lock);
-  return 0;
-
-found:
-  p->state = EMBRYO;
-  p->pid = nextpid++;
-  release(&ptable.lock);
-
-  // Allocate kernel stack if possible.
-  if((p->kstack = kalloc()) == 0){
-    p->state = UNUSED;
-    return 0;
-  }
-  sp = p->kstack + KSTACKSIZE;
-  
-  // Leave room for trap frame.
-  sp -= sizeof *p->tf;
-  p->tf = (struct trapframe*)sp;
-  
-  // Set up new context to start executing at forkret,
-  // which returns to trapret.
-  sp -= 4;
-  *(uint*)sp = (uint)trapret;
-
-  sp -= sizeof *p->context;
-  p->context = (struct context*)sp;
-  memset(p->context, 0, sizeof *p->context);
-  p->context->eip = (uint)forkret;
-  p->priority = 0;
-  p->ticks[0] = p->ticks[1] = p->ticks[2] = p->ticks[3] = 0;
-  p->lastScheduledOnTick = 0;
-  return p;
-}
-
-// Set up first user process.
-void
-userinit(void)
-{
-  struct proc *p;
-  extern char _binary_initcode_start[], _binary_initcode_size[];
-  
-  p = allocproc();
-  acquire(&ptable.lock);
-  initproc = p;
-  if((p->pgdir = setupkvm()) == 0)
-    panic("userinit: out of memory?");
[…]

### 2024-11-11 Research/MLFQ-Example/kernel/proc.h

@@ -1,92 +1,0 @@
-#ifndef _PROC_H_
-#define _PROC_H_
-// Segments in proc->gdt.
-// Also known to bootasm.S and trapasm.S
-#define SEG_KCODE 1  // kernel code
-#define SEG_KDATA 2  // kernel data+stack
-#define SEG_KCPU  3  // kernel per-cpu data
-#define SEG_UCODE 4  // user code
-#define SEG_UDATA 5  // user data+stack
-#define SEG_TSS   6  // this process's task state
-#define NSEGS     7
-
-#include "../include/pstat.h"
-
-// Per-CPU state
-struct cpu {
-  uchar id;                    // Local APIC ID; index into cpus[] below
-  struct context *scheduler;   // swtch() here to enter scheduler
-  struct taskstate ts;         // Used by x86 to find stack for interrupt
-  struct segdesc gdt[NSEGS];   // x86 global descriptor table
-  volatile uint booted;        // Has the CPU started?
-  int ncli;                    // Depth of pushcli nesting.
-  int intena;                  // Were interrupts enabled before pushcli?
-
-  // Cpu-local storage variables; see below
-  struct cpu *cpu;
-  struct proc *proc;           // The currently-running process.
-};
-
-extern struct cpu cpus[NCPU];
-extern int ncpu;
-
-// Per-CPU variables, holding pointers to the
-// current cpu and to the current process.
-// The asm suffix tells gcc to use "%gs:0" to refer to cpu
-// and "%gs:4" to refer to proc.  seginit sets up the
-// %gs segment register so that %gs refers to the memory
-// holding those two variables in the local cpu's struct cpu.
-// This is similar to how thread-local variables are implemented
-// in thread libraries such as Linux pthreads.
-extern struct cpu *cpu asm("%gs:0");       // &cpus[cpunum()]
-extern struct proc *proc asm("%gs:4");     // cpus[cpunum()].proc
-
-// Saved registers for kernel context switches.
-// Don't need to save all the segment registers (%cs, etc),
-// because they are constant across kernel contexts.
-// Don't need to save %eax, %ecx, %edx, because the
-// x86 convention is that the caller has saved them.
-// Contexts are stored at the bottom of the stack they
-// describe; the stack pointer is the address of the context.
-// The layout of the context matches the layout of the stack in swtch.S
-// at the "Switch stacks" comment. Switch doesn't save eip explicitly,
-// but it is on the stack and allocproc() manipulates it.
-struct context {
-  uint edi;
-  uint esi;
[…]

### 2024-11-11 Research/MLFQ-Example/kernel/sysproc.c

@@ -1,100 +1,0 @@
-#include "types.h"
-#include "x86.h"
-#include "defs.h"
-#include "param.h"
-#include "mmu.h"
-#include "proc.h"
-#include "sysfunc.h"
-#include "../include/pstat.h"
-
-int
-sys_fork(void)
-{
-  return fork();
-}
-
-int
-sys_exit(void)
-{
-  exit();
-  return 0;  // not reached
-}
-
-int
-sys_wait(void)
-{
-  return wait();
-}
-
-int
-sys_kill(void)
-{
-  int pid;
-
-  if(argint(0, &pid) < 0)
-    return -1;
-  return kill(pid);
-}
-
-int
-sys_getpid(void)
-{
-  return proc->pid;
-}
-
-int
-sys_sbrk(void)
-{
-  int addr;
-  int n;
-
-  if(argint(0, &n) < 0)
-    return -1;
-  addr = proc->sz;
-  if(growproc(n) < 0)
-    return -1;
-  return addr;
-}
-
-int
-sys_sleep(void)
-{
-  int n;
-  uint ticks0;
-  
-  if(argint(0, &n) < 0)
-    return -1;
-  acquire(&tickslock);
-  ticks0 = ticks;
-  while(ticks - ticks0 < n){
-    if(proc->killed){
-      release(&tickslock);
-      return -1;
-    }
-    sleep(&ticks, &tickslock);
-  }
-  release(&tickslock);
-  return 0;
-}
-
-// return how many clock tick interrupts have occurred
-// since boot.
-int
-sys_uptime(void)
-{
-  uint xticks;
-  
-  acquire(&tickslock);
-  xticks = ticks;
-  release(&tickslock);
-  return xticks;
-}
-
-int
-sys_getpinfo(void)
-{
-  char* pstat;
-  if(argptr(0, &pstat, sizeof(struct pstat)) < 0)
-    return -1;
-  return getpinfo((struct pstat*)pstat);
-}

### 2024-11-11 Research/MLFQ-Example/kernel/trap.c

@@ -1,112 +1,0 @@
-#include "types.h"
-#include "defs.h"
-#include "param.h"
-#include "mmu.h"
-#include "proc.h"
-#include "x86.h"
-#include "traps.h"
-#include "spinlock.h"
-
-// Interrupt descriptor table (shared by all CPUs).
-struct gatedesc idt[256];
-extern uint vectors[];  // in vectors.S: array of 256 entry pointers
-struct spinlock tickslock;
-uint ticks;
-
-void
-tvinit(void)
-{
-  int i;
-
-  for(i = 0; i < 256; i++)
-    SETGATE(idt[i], 0, SEG_KCODE<<3, vectors[i], 0);
-  SETGATE(idt[T_SYSCALL], 1, SEG_KCODE<<3, vectors[T_SYSCALL], DPL_USER);
-  
-  initlock(&tickslock, "time");
-}
-
-void
-idtinit(void)
-{
-  lidt(idt, sizeof(idt));
-}
-
-void
-trap(struct trapframe *tf)
-{
-  if(tf->trapno == T_SYSCALL){
-    if(proc->killed)
-      exit();
-    proc->tf = tf;
-    syscall();
-    if(proc->killed)
-      exit();
-    return;
-  }
-
-  switch(tf->trapno){
-  case T_IRQ0 + IRQ_TIMER:
-    if(cpu->id == 0){
-      acquire(&tickslock);
-      ticks++;
-      if(proc){
-        proc->ticks[proc->priority]++;
-      }
-      wakeup(&ticks);
-      release(&tickslock);
-    }
-    lapiceoi();
-    break;
-  case T_IRQ0 + IRQ_IDE:
-    ideintr();
-    lapiceoi();
-    break;
-  case T_IRQ0 + IRQ_IDE+1:
-    // Bochs generates spurious IDE1 interrupts.
-    break;
-  case T_IRQ0 + IRQ_KBD:
-    kbdintr();
-    lapiceoi();
-    break;
-  case T_IRQ0 + IRQ_COM1:
-    uartintr();
-    lapiceoi();
-    break;
-  case T_IRQ0 + 7:
-  case T_IRQ0 + IRQ_SPURIOUS:
-    cprintf("cpu%d: spurious interrupt at %x:%x\n",
-            cpu->id, tf->cs, tf->eip);
-    lapiceoi();
-    break;
-   
-  default:
-    if(proc == 0 || (tf->cs&3) == 0){
[…]

### 2024-11-11 Research/xv6/fmt/console.c

@@ -1,400 +1,0 @@
-7900 // Console input and output.
-7901 // Input is from the keyboard or serial port.
-7902 // Output is written to the screen and serial port.
-7903 
-7904 #include "types.h"
-7905 #include "defs.h"
-7906 #include "param.h"
-7907 #include "traps.h"
-7908 #include "spinlock.h"
-7909 #include "sleeplock.h"
-7910 #include "fs.h"
-7911 #include "file.h"
-7912 #include "memlayout.h"
-7913 #include "mmu.h"
-7914 #include "proc.h"
-7915 #include "x86.h"
-7916 
-7917 static void consputc(int);
-7918 
-7919 static int panicked = 0;
-7920 
-7921 static struct {
-7922   struct spinlock lock;
-7923   int locking;
-7924 } cons;
-7925 
-7926 static void
-7927 printint(int xx, int base, int sign)
-7928 {
-7929   static char digits[] = "0123456789abcdef";
-7930   char buf[16];
-7931   int i;
-7932   uint x;
-7933 
-7934   if(sign && (sign = xx < 0))
-7935     x = -xx;
-7936   else
-7937     x = xx;
-7938 
-7939   i = 0;
-7940   do{
-7941     buf[i++] = digits[x % base];
-7942   }while((x /= base) != 0);
-7943 
-7944   if(sign)
-7945     buf[i++] = '-';
-7946 
-7947   while(--i >= 0)
-7948     consputc(buf[i]);
-7949 }
-7950 
-7951 
-7952 
-7953 
-7954 
-7955 
-7956 
-7957 
-7958 
-7959 
-7960 
-7961 
-7962 
-7963 
-7964 
-7965 
-7966 
-7967 
-7968 
-7969 
-7970 
-7971 
-7972 
-7973 
-7974 
-7975 
-7976 
-7977 
-7978 
-7979 
-7980 
-7981 
-7982 
-7983 
-7984 
-7985 
-7986 
-7987 
-7988 
-7989 
-7990 
-7991 
-7992 
-7993 
-7994 
-7995 
-7996 
-7997 
-7998 
-7999 
-8000 // Print to the console. only understands %d, %x, %p, %s.
-8001 void
-8002 cprintf(char *fmt, ...)
-8003 {
-8004   int i, c, locking;
-8005   uint *argp;
-8006   char *s;
-8007 
-8008   locking = cons.locking;
-8009   if(locking)
-8010     acquire(&cons.lock);
-8011 
-8012   if (fmt == 0)
[…]

### 2024-11-11 Research/xv6/fmt/fs.c

@@ -1,900 +1,0 @@
-4950 // File system implementation.  Five layers:
-4951 //   + Blocks: allocator for raw disk blocks.
-4952 //   + Log: crash recovery for multi-step updates.
-4953 //   + Files: inode allocator, reading, writing, metadata.
-4954 //   + Directories: inode with special contents (list of other inodes!)
-4955 //   + Names: paths like /usr/rtm/xv6/fs.c for convenient naming.
-4956 //
-4957 // This file contains the low-level file system manipulation
-4958 // routines.  The (higher-level) system call implementations
-4959 // are in sysfile.c.
-4960 
-4961 #include "types.h"
-4962 #include "defs.h"
-4963 #include "param.h"
-4964 #include "stat.h"
-4965 #include "mmu.h"
-4966 #include "proc.h"
-4967 #include "spinlock.h"
-4968 #include "sleeplock.h"
-4969 #include "fs.h"
-4970 #include "buf.h"
-4971 #include "file.h"
-4972 
-4973 #define min(a, b) ((a) < (b) ? (a) : (b))
-4974 static void itrunc(struct inode*);
-4975 // there should be one superblock per disk device, but we run with
-4976 // only one device
-4977 struct superblock sb;
-4978 
-4979 // Read the super block.
-4980 void
-4981 readsb(int dev, struct superblock *sb)
-4982 {
-4983   struct buf *bp;
-4984 
-4985   bp = bread(dev, 1);
-4986   memmove(sb, bp->data, sizeof(*sb));
-4987   brelse(bp);
-4988 }
-4989 
-4990 
-4991 
-4992 
-4993 
-4994 
-4995 
-4996 
-4997 
-4998 
-4999 
-5000 // Zero a block.
-5001 static void
-5002 bzero(int dev, int bno)
-5003 {
-5004   struct buf *bp;
-5005 
-5006   bp = bread(dev, bno);
-5007   memset(bp->data, 0, BSIZE);
-5008   log_write(bp);
-5009   brelse(bp);
-5010 }
-5011 
-5012 // Blocks.
-5013 
-5014 // Allocate a zeroed disk block.
-5015 static uint
-5016 balloc(uint dev)
-5017 {
-5018   int b, bi, m;
-5019   struct buf *bp;
-5020 
-5021   bp = 0;
-5022   for(b = 0; b < sb.size; b += BPB){
-5023     bp = bread(dev, BBLOCK(b, sb));
[…]

### 2024-11-11 Research/xv6/fmt/lapic.c

@@ -1,250 +1,0 @@
-7350 // The local APIC manages internal (non-I/O) interrupts.
-7351 // See Chapter 8 & Appendix C of Intel processor manual volume 3.
-7352 
-7353 #include "param.h"
-7354 #include "types.h"
-7355 #include "defs.h"
-7356 #include "date.h"
-7357 #include "memlayout.h"
-7358 #include "traps.h"
-7359 #include "mmu.h"
-7360 #include "x86.h"
-7361 
-7362 // Local APIC registers, divided by 4 for use as uint[] indices.
-7363 #define ID      (0x0020/4)   // ID
-7364 #define VER     (0x0030/4)   // Version
-7365 #define TPR     (0x0080/4)   // Task Priority
-7366 #define EOI     (0x00B0/4)   // EOI
-7367 #define SVR     (0x00F0/4)   // Spurious Interrupt Vector
-7368   #define ENABLE     0x00000100   // Unit Enable
-7369 #define ESR     (0x0280/4)   // Error Status
-7370 #define ICRLO   (0x0300/4)   // Interrupt Command
-7371   #define INIT       0x00000500   // INIT/RESET
-7372   #define STARTUP    0x00000600   // Startup IPI
-7373   #define DELIVS     0x00001000   // Delivery status
-7374   #define ASSERT     0x00004000   // Assert interrupt (vs deassert)
-7375   #define DEASSERT   0x00000000
-7376   #define LEVEL      0x00008000   // Level triggered
-7377   #define BCAST      0x00080000   // Send to all APICs, including self.
-7378   #define BUSY       0x00001000
-7379   #define FIXED      0x00000000
-7380 #define ICRHI   (0x0310/4)   // Interrupt Command [63:32]
-7381 #define TIMER   (0x0320/4)   // Local Vector Table 0 (TIMER)
-7382   #define X1         0x0000000B   // divide counts by 1
-7383   #define PERIODIC   0x00020000   // Periodic
-7384 #define PCINT   (0x0340/4)   // Performance Counter LVT
-7385 #define LINT0   (0x0350/4)   // Local Vector Table 1 (LINT0)
-7386 #define LINT1   (0x0360/4)   // Local Vector Table 2 (LINT1)
[…]

### 2024-11-11 Research/xv6/fmt/sleeplock.c

@@ -1,100 +1,0 @@
-4600 // Sleeping locks
-4601 
-4602 #include "types.h"
-4603 #include "defs.h"
-4604 #include "param.h"
-4605 #include "x86.h"
-4606 #include "memlayout.h"
-4607 #include "mmu.h"
-4608 #include "proc.h"
-4609 #include "spinlock.h"
-4610 #include "sleeplock.h"
-4611 
-4612 void
-4613 initsleeplock(struct sleeplock *lk, char *name)
-4614 {
-4615   initlock(&lk->lk, "sleep lock");
-4616   lk->name = name;
-4617   lk->locked = 0;
-4618   lk->pid = 0;
-4619 }
-4620 
-4621 void
-4622 acquiresleep(struct sleeplock *lk)
-4623 {
-4624   acquire(&lk->lk);
-4625   while (lk->locked) {
-4626     sleep(lk, &lk->lk);
-4627   }
-4628   lk->locked = 1;
-4629   lk->pid = myproc()->pid;
-4630   release(&lk->lk);
-4631 }
-4632 
-4633 void
-4634 releasesleep(struct sleeplock *lk)
-4635 {
-4636   acquire(&lk->lk);
-4637   lk->locked = 0;
-4638   lk->pid = 0;
-4639   wakeup(lk);
-4640   release(&lk->lk);
-4641 }
-4642 
-4643 
-4644 
-4645 
-4646 
-4647 
-4648 
-4649 
-4650 int
-4651 holdingsleep(struct sleeplock *lk)
-4652 {
-4653   int r;
-4654 
-4655   acquire(&lk->lk);
-4656   r = lk->locked && (lk->pid == myproc()->pid);
-4657   release(&lk->lk);
-4658   return r;
-4659 }
-4660 
-4661 
-4662 
-4663 
-4664 
-4665 
-4666 
-4667 
-4668 
-4669 
-4670 
-4671 
-4672 
-4673 
-4674 
-4675 
-4676 
-4677 
-4678 
-4679 
-4680 
-4681 
-4682 
-4683 
-4684 
-4685 
-4686 
-4687 
-4688 
-4689 
-4690 
-4691 
-4692 
-4693 
-4694 
-4695 
-4696 
-4697 
-4698 
-4699 

### 2024-11-11 Research/xv6/fmt/sysfile.c

@@ -1,550 +1,0 @@
-6050 //
-6051 // File-system system calls.
-6052 // Mostly argument checking, since we don't trust
-6053 // user code, and calls into file.c and fs.c.
-6054 //
-6055 
-6056 #include "types.h"
-6057 #include "defs.h"
-6058 #include "param.h"
-6059 #include "stat.h"
-6060 #include "mmu.h"
-6061 #include "proc.h"
-6062 #include "fs.h"
-6063 #include "spinlock.h"
-6064 #include "sleeplock.h"
-6065 #include "file.h"
-6066 #include "fcntl.h"
-6067 
-6068 // Fetch the nth word-sized system call argument as a file descriptor
-6069 // and return both the descriptor and the corresponding struct file.
-6070 static int
-6071 argfd(int n, int *pfd, struct file **pf)
-6072 {
-6073   int fd;
-6074   struct file *f;
-6075 
-6076   if(argint(n, &fd) < 0)
-6077     return -1;
-6078   if(fd < 0 || fd >= NOFILE || (f=myproc()->ofile[fd]) == 0)
-6079     return -1;
-6080   if(pfd)
-6081     *pfd = fd;
-6082   if(pf)
-6083     *pf = f;
-6084   return 0;
-6085 }
-6086 
-6087 
-6088 
-6089 
-6090 
-6091 
-6092 
-6093 
-6094 
-6095 
-6096 
-6097 
-6098 
-6099 
-6100 // Allocate a file descriptor for the given file.
-6101 // Takes over file reference from caller on success.
-6102 static int
-6103 fdalloc(struct file *f)
-6104 {
-6105   int fd;
-6106   struct proc *curproc = myproc();
-6107 
-6108   for(fd = 0; fd < NOFILE; fd++){
-6109     if(curproc->ofile[fd] == 0){
-6110       curproc->ofile[fd] = f;
-6111       return fd;
-6112     }
-6113   }
-6114   return -1;
-6115 }
-6116 
-6117 int
-6118 sys_dup(void)
-6119 {
-6120   struct file *f;
-6121   int fd;
-6122 
-6123   if(argfd(0, 0, &f) < 0)
-6124     return -1;
-6125   if((fd=fdalloc(f)) < 0)
-6126     return -1;
-6127   filedup(f);
-6128   return fd;
[…]

### 2024-11-12 LaTeX/Thesis.tex
added a tex template and adjusted title page

\documentclass{mimosis}

\usepackage{metalogo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Some of my favourite personal adjustments
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% These are the adjustments that I consider necessary for typesetting
% a nice thesis. However, they are *not* included in the template, as
% I do not want to force you to use them.

% This ensures that I am able to typeset bold font in table while still aligning the numbers
% correctly.
\usepackage{etoolbox}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Hyperlinks & bookmarks
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[%
  colorlinks = true,
  citecolor  = RoyalBlue,
  linkcolor  = RoyalBlue,
  urlcolor   = RoyalBlue,
  unicode,
  ]{hyperref}

\usepackage{bookmark}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% I like the bibliography to be extremely plain, showing only a numeric
%% identifier and citing everything in simple brackets. The first names,
%% if present, will be initialized. DOIs and URLs will be preserved.
%
%\usepackage[%
%  autocite     = plain,
%  backend      = biber,
%  doi          = true,
%  url          = true,
%  giveninits   = true,
%  hyperref     = true,
%  maxbibnames  = 99,
%  maxcitenames = 99,
%  sortcites    = true,
%  style        = numeric,
%  ]{biblatex}
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Fonts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\ifxetexorluatex
  \usepackage{unicode-math}
  \setmainfont{EB Garamond}
  \setmathfont{Garamond Math}

  % Load some missing symbols from another font.
  \setmathfont{STIX Two Math}[%
    range = {
      \sharp,
      \natural,
      \flat,
      \clubsuit,
      \spadesuit,
      \checkmark
    }
  ]
  \setmonofont[Scale=MatchLowercase]{Source Code Pro}
\else
  \usepackage[lf]{ebgaramond}
  \usepackage[oldstyle,scale=0.7]{sourcecodepro}
  \singlespacing
\fi

\newglossaryentry{LaTeX}{%
  name        = {\LaTeX},
  description = {A document preparation system},
  sort        = {LaTeX},
}

\newglossaryentry{Real numbers}{%
  name        = {$\real$},
  description = {The set of real numbers},
  sort        = {Real numbers},
}

\makeindex
\makeglossaries

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Ordinals
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\makeatletter
\@ifundefined{st}{%
  \newcommand{\st}{\textsuperscript{\textup{st}}\xspace}
}{}
\@ifundefined{rd}{%
  \newcommand{\rd}{\textsuperscript{\textup{rd}}\xspace}
}{}
\@ifundefined{nd}{%
  \newcommand{\nd}{\textsuperscript{\textup{nd}}\xspace}
}{}
\makeatother

\renewcommand{\th}{\textsuperscript{\textup{th}}\xspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Incipit
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{{CPU-Scheduling}}
\subtitle{What a waiting in line at the supermarket and your computer has in common}
\author{Mark Krutzler}

\begin{document}

\frontmatter
  \include{Sources/Title}
  \include{Sources/Abstract}

  \tableofcontents

\mainmatter

  \part[A good part]{%
    A good part\\
    %
    \vspace{1cm}
    %
    \begin{minipage}[l]{\textwidth}
    %
    \textnormal{%
      \normalsize
      %
      \begin{singlespace*}
        \onehalfspacing
        %
        You can also use parts in order to partition your great work
        into larger `chunks'. This involves some manual adjustments in
        terms of the layout, though.
      \end{singlespace*}
    }
    \end{minipage}
  }

  \include{Sources/Introduction}

% This ensures that the subsequent sections are being included as root
% items in the bookmark structure of your PDF reader.
\bookmarksetup{startatroot}
\backmatter

  \begingroup
    \let\clearpage\relax
    \glsaddall
   \printglossary
  \endgroup

  \printindex
  % \printbibliography

\end{document}

### 2024-11-13 LaTeX/Sources/Basics.tex
further tweaking of the template

--- 2024-11-12 LaTeX/Thesis.tex
@@
-\documentclass{mimosis}
-
-\usepackage{metalogo}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Some of my favourite personal adjustments
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% These are the adjustments that I consider necessary for typesetting
-% a nice thesis. However, they are *not* included in the template, as
-% I do not want to force you to use them.
-
-% This ensures that I am able to typeset bold font in table while still aligning the numbers
-% correctly.
-\usepackage{etoolbox}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Hyperlinks & bookmarks
+\chapter{Example}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\usepackage[%
-  colorlinks = true,
-  citecolor  = RoyalBlue,
-  linkcolor  = RoyalBlue,
-  urlcolor   = RoyalBlue,
-  unicode,
-  ]{hyperref}
-
-\usepackage{bookmark}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% I like the bibliography to be extremely plain, showing only a numeric
-%% identifier and citing everything in simple brackets. The first names,
-%% if present, will be initialized. DOIs and URLs will be preserved.
-%
-%\usepackage[%
-%  autocite     = plain,
-%  backend      = biber,
-%  doi          = true,
-%  url          = true,
-%  giveninits   = true,
-%  hyperref     = true,
-%  maxbibnames  = 99,
-%  maxcitenames = 99,
-%  sortcites    = true,
-%  style        = numeric,
-%  ]{biblatex}
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Fonts
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\ifxetexorluatex
-  \usepackage{unicode-math}
-  \setmainfont{EB Garamond}
-  \setmathfont{Garamond Math}
-
-  % Load some missing symbols from another font.
-  \setmathfont{STIX Two Math}[%
-    range = {
-      \sharp,
-      \natural,
-      \flat,
-      \clubsuit,
-      \spadesuit,
-      \checkmark
-    }
-  ]
-  \setmonofont[Scale=MatchLowercase]{Source Code Pro}
-\else
-  \usepackage[lf]{ebgaramond}
-  \usepackage[oldstyle,scale=0.7]{sourcecodepro}
-  \singlespacing
-\fi
-
-\newglossaryentry{LaTeX}{%
-  name        = {\LaTeX},
-  description = {A document preparation system},
-  sort        = {LaTeX},
-}
-
-\newglossaryentry{Real numbers}{%
-  name        = {$\real$},
-  description = {The set of real numbers},
-  sort        = {Real numbers},
-}
-
-\makeindex
-\makeglossaries
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Ordinals
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\makeatletter
-\@ifundefined{st}{%
-  \newcommand{\st}{\textsuperscript{\textup{st}}\xspace}
-}{}
-\@ifundefined{rd}{%
-  \newcommand{\rd}{\textsuperscript{\textup{rd}}\xspace}
-}{}
-\@ifundefined{nd}{%
-  \newcommand{\nd}{\textsuperscript{\textup{nd}}\xspace}
-}{}
-\makeatother
-
-\renewcommand{\th}{\textsuperscript{\textup{th}}\xspace}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Incipit
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{{CPU-Scheduling}}
-\subtitle{What a waiting in line at the supermarket and your computer has in common}
-\author{Mark Krutzler}
-
-\begin{document}
-
-\frontmatter
-  \include{Sources/Title}
-  \include{Sources/Abstract}
-
-  \tableofcontents
-
-\mainmatter
-
-  \part[A good part]{%
-    A good part\\
-    %
-    \vspace{1cm}
-    %
-    \begin{minipage}[l]{\textwidth}
-    %
-    \textnormal{%
-      \normalsize
-      %
-      \begin{singlespace*}
-        \onehalfspacing
-        %
-        You can also use parts in order to partition your great work
-        into larger `chunks'. This involves some manual adjustments in
-        terms of the layout, though.
-      \end{singlespace*}
-    }
-    \end{minipage}
-  }
-
-  \include{Sources/Introduction}
-
-% This ensures that the subsequent sections are being included as root
-% items in the bookmark structure of your PDF reader.
-\bookmarksetup{startatroot}
-\backmatter
-
-  \begingroup
-    \let\clearpage\relax
-    \glsaddall
-   \printglossary
-  \endgroup
-
-  \printindex
-  % \printbibliography
-
-\end{document}
+here is some text

### 2024-11-13 LaTeX/Thesis.tex
further tweaking of the template

@@
 \title{{CPU-Scheduling}}
-\subtitle{What a waiting in line at the supermarket and your computer has in common}
+\subtitle{What waiting in line at the supermarket and your computer have in common}
 \author{Mark Krutzler}
@@
 
-  \part[A good part]{%
-    A good part\\
+  \part[The ABC's of CPU scheduling]{%
+    The ABC's of CPU scheduling\\
     %
@@
         %
-        You can also use parts in order to partition your great work
-        into larger `chunks'. This involves some manual adjustments in
-        terms of the layout, though.
+        Introduciton to Basics
       \end{singlespace*}
@@
 
-  \include{Sources/Introduction}
+  \include{Sources/Basics}
 
@@
 
-  \begingroup
-    \let\clearpage\relax
-    \glsaddall
-   \printglossary
-  \endgroup
-
-  \printindex
-  % \printbibliography
+\chapter{Sources}
+\begin{itemize}
+    \item source 1
+    \item source 2 
+\end{itemize}
 

### 2024-11-15 LaTeX/Sources/Basics.tex
added glossaries
added "minted" for code blocks

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Example}
+\chapter{Initial Problem}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-here is some text
+The main problem we face can be simplified down into a really easy to formulate but hard to answer question.
+How can we order the queued tasks, so that they run in the most optimal way?
+What optimal really means is a whole discussion itself not to mention search for the perfect algorithm to achieve that "best" solution.
+In this chapter I will give you a quick introduction to the main terms.
+In addition to that I will also try to make it as easy as possible to understand by using a situation that most of us already suffered through.
+
+Have you ever wondered why so many people buy already prebottled water? 
+How they could eat those snack that you despise?
+Why they have a whole plate full of milk in their shopping cart?
+Maybe these question seem rather sudden they have two things in common:
+One, you never get the answer to them and two, which is the relevant part, you ask them while standing in line at a supermarket.
+You ask yourself them while staring at the family of five with two shopping carts and while questioning, whether or not you should have queued at another line.
+Also wouldn't it be much better if you, with your two items, went before them?
+What is the best order to queue up these people?
+As you can already see this is the original, simplified question just reformulated.
+We are seaching for the best algorithm that arranges people at the supermarket.
+Matter of fact let's just call it by its real name: policy.
+In CPU-Scheduling we call the "algorithms" policies or disciplines.
+A single customer / shopper is called a job or a process.
+In the real world a process can be anything from your drivers to the web browser showing cute cat images.
+A queue is a line of people waiting to get their items scanned and the cashier is the "CPU".
+
+\begin{figure}[h]
+\begin{minted}[mathescape,
+    linenos,
+    numbersep=5pt,
+    gobble=2,
+    frame=lines,
+    framesep=2mm,
+    ]{python}
+  # test
+  print("Hello World!")
+  variable = 1
+  print(variable)
+\end{minted}
+\caption{Python: First in, First out}
+\label{code:fifo}
+\end{figure}
+
+Here is a reference to \hyperref[code:fifo]{code block}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{How do we compare policies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Evolving Supermarket}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-11-24 LaTeX/clean.sh
added clean.sh

@@ -1,0 +1,9 @@
+rm Sources/*.aux
+rm *.acn
+rm *.aux
+rm *.glo
+rm *.idx
+rm *.ist
+rm *.log
+rm *.toc
+rm *.synctex.gz

### 2024-11-25 LaTeX/Sources/Basics.tex
continued first chapter

@@
 
-The main problem we face can be simplified down into a really easy to formulate but hard to answer question.
+\section{Welcome to Costco}
+
+The main problem we face can be broken down into a really easy to formulate but hard to answer question.
 How can we order the queued tasks, so that they run in the most optimal way?
@@
 In this chapter I will give you a quick introduction to the main terms.
-In addition to that I will also try to make it as easy as possible to understand by using a situation that most of us already suffered through.
+In addition, I will also try to make it as easy as possible to understand by using a situation that most of us already suffered through.
 
-Have you ever wondered why so many people buy already prebottled water? 
-How they could eat those snack that you despise?
+Have you ever wondered why so many people buy prebottled water? 
+How they could eat those nasty snack that you despise?
 Why they have a whole plate full of milk in their shopping cart?
@@
 
+\section{How we compare policies}
+
+Before we dive deeper into theory it is important to note that there is a huge difference between fairness and performance.
+Fairness usually makes sure that everyone recieves the same amount of CPU time.
+This kind of policy turn out to be cyclic, which usually leads to a more responsive system, because each job gets a bit of activity every so often.
+On the other hand if we want to optimize for performance we should look at the so called "average turnaround time".
+Turnaround time is nothing more than the time that a person has to stand in line.
+Therefore if the take the average we'll have average waiting time until you exit the supermarket.
+
+\section{First lines of code}
+
+During the next Section we'll asume that one knows the amount of time that a process will take to finish.
+In a real world this is almost impossible, except if you can time travel.
+With this and setting aside the optimization part, one of the most straightforward policy is called "first comes, first served".
+This is what we usually suffer through in the queue to pay.
+The policy disregards how many item one has in the cart.
+The only important part is the time you arrive.
+The sooner the better.
+In the world of computer science the policy is better refered to as "first in, first out" and abbriviated as "FIFO".
+Here is a \hyperref[code:fifo]{simple python implementation}.
+
+We recieve the queue of people as a "list".
+If any other job joins, it will just get appended to the end.
+As for the scheduling it self:
+We just loop through the list until everyone is finished.
+
 \begin{figure}[h]
@@
     ]{python}
-  # test
-  print("Hello World!")
-  variable = 1
-  print(variable)
+  # FIFO implementation in python
+  queue = [] # initialize empty queue
+
+  # Adding a new process
+  def add_process(process):
+    queue.append(process)
+
+  # Schedule the processes
+  while queue != []:
+    if new_process = true: # check if there is a new process
+        add_process(process)
+
+    next = queue.pop(0) # picks next process
+    use_resource(next) # uses resources until finished
 \end{minted}
@@
 \end{figure}
-
-Here is a reference to \hyperref[code:fifo]{code block}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{How do we compare policies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2024-11-26 LaTeX/Sources/Basics.tex
Finished first chapter

@@
 The main problem we face can be broken down into a really easy to formulate but hard to answer question.
-How can we order the queued tasks, so that they run in the most optimal way?
-What optimal really means is a whole discussion itself not to mention search for the perfect algorithm to achieve that "best" solution.
+How can we order the queued tasks, so that they run in the most optimal order?
+What optimal really means is a whole discussion itself not to mention search for the perfect algorithm to achieve that desired best solution.
 In this chapter I will give you a quick introduction to the main terms.
-In addition, I will also try to make it as easy as possible to understand by using a situation that most of us already suffered through.
+In addition, I will also try to make it as easy as possible to understand by using a situation that most of us have already suffered through.
 
 Have you ever wondered why so many people buy prebottled water? 
-How they could eat those nasty snack that you despise?
-Why they have a whole plate full of milk in their shopping cart?
+How they can eat those nasty snack that you despise?
+Why they have the whole cart full of milk?
 Maybe these question seem rather sudden they have two things in common:
-One, you never get the answer to them and two, which is the relevant part, you ask them while standing in line at a supermarket.
+One, you never get the answer to them, and two, which is the relevant part, you ask them while standing in line at a supermarket.
 You ask yourself them while staring at the family of five with two shopping carts and while questioning, whether or not you should have queued at another line.
@@
 Matter of fact let's just call it by its real name: policy.
-In CPU-Scheduling we call the "algorithms" policies or disciplines.
+In CPU-Scheduling we call the algorithms policies or disciplines.
 A single customer / shopper is called a job or a process.
 In the real world a process can be anything from your drivers to the web browser showing cute cat images.
-A queue is a line of people waiting to get their items scanned and the cashier is the "CPU".
+A queue is a line of people waiting to get their items scanned and the cashier is the CPU.
 
@@
 
-Before we dive deeper into theory it is important to note that there is a huge difference between fairness and performance.
+Before we dive deeper into theory, it is important to note that there is a huge difference between fairness and performance.
 Fairness usually makes sure that everyone recieves the same amount of CPU time.
-This kind of policy turn out to be cyclic, which usually leads to a more responsive system, because each job gets a bit of activity every so often.
-On the other hand if we want to optimize for performance we should look at the so called "average turnaround time".
+This kind of policy turns out to be cyclic, which usually leads to a more responsive system, because each job gets a bit of activity every so often.
+On the other hand if we want to optimize for performance we should look at the so called average turnaround time.
 Turnaround time is nothing more than the time that a person has to stand in line.
-Therefore if the take the average we'll have average waiting time until you exit the supermarket.
+Therefore the average of it is the average waiting time until you exit the supermarket.
 
@@
 
-During the next Section we'll asume that one knows the amount of time that a process will take to finish.
-In a real world this is almost impossible, except if you can time travel.
-With this and setting aside the optimization part, one of the most straightforward policy is called "first comes, first served".
+During the next few chapters and sections we'll assume that one knows the amount of time that a process will take to finish, also known as burst time.
+Like bursting out of joy, when you finally finish the weekly shopping.
+In a real world this is almost impossible\footnote{Here I mean knowning the burst time and not escaping the supermarket.}, except if you can time travel.
+With this and setting aside the optimization part, one of the most straightforward policy is called first come, first served.
 This is what we usually suffer through in the queue to pay.
@@
 The sooner the better.
-In the world of computer science the policy is better refered to as "first in, first out" and abbriviated as "FIFO".
+In the world of computer science the policy is better refered to as First In, First Out (FIFO).
 Here is a \hyperref[code:fifo]{simple python implementation}.
 
-We recieve the queue of people as a "list".
+We save the queue of people as a python list.
 If any other job joins, it will just get appended to the end.
@@
   while queue != []:
-    if new_process = true: # check if there is a new process
+    if new_process == True: # check if there is a new process
         add_process(process)
@@
 \end{minted}
+\label{code:fifo}
 \caption{Python: First in, First out}
-\label{code:fifo}
 \end{figure}
+In this example we just assume that the use\_resource function on line 14 is already written.
+Also it is important to mention that the new\_process variable is just a bool, which gets update if a new process is waiting to join the queue.
 

### 2024-11-26 LaTeX/clean.sh

@@ -1,4 +1,5 @@
 rm Sources/*.aux
+rm Sources/*.log
 rm *.acn
 rm *.aux
 rm *.glo

### 2024-11-26 LaTeX/compile.sh

@@ -1,0 +1,2 @@
+#!/usr/bin/env sh
+pdflatex -synctex=1 -shell-escape -interaction=nonstopmode Thesis.tex

### 2024-11-26 LaTeX/clean.sh
modified clean.sh

@@ -8,3 +8,4 @@ rm *.ist
 rm *.log
 rm *.toc
 rm *.synctex.gz
+rm -rf _minted-Thesis

### 2024-12-08 LaTeX/Sources/Basics.tex
sun 8. dec: almost finished basics

@@
 
-\section{First lines of code}
+\section{First lines of code} \label{sec:first-lines-of-code}
 
@@
 Also it is important to mention that the new\_process variable is just a bool, which gets update if a new process is waiting to join the queue.
+In addition to that the empty queue initialization is purely symbolic, because if the queue is already empty, there will be no scheduling happening. 
+In general python is not really meant for low level programming anyways, so take it with a grain of salt.
 
@@
 
+Now that the scene is set, we return the original question. 
+How can we reduce our waiting time?
+To give an answer we will first take a step back.
+While the previous chapter talked about the initial situation and proper terms, this one will first loosen up some of the requirements in order to ensure that the policies are as easy to understand as possible.
+
+\section{Shorterst Job First}
+
+As the name already says it, the shortest job will be run first.
+In order to achieve this we assume that the cashier knows, who has the least amount of items.
+This buyer will be than handeled before anybody else. As you can already tell, this is not really a fair way to treat your customers.
+Still if it comes down to pure turnaround time it is far better than the basic First In, First Out.
+
+\begin{figure}[h]
+    \begin{minted}[mathescape,
+        linenos,
+        numbersep=5pt,
+        gobble=2,
+        frame=lines,
+        framesep=2mm,
+        ]{python}
+      # Shortest Job First implementation in python
+      class Process:
+        def __init__(self, burst):
+            self.burst = burst
+
+      # Adding a new process
+      def add_process(process):
+        queue.append(process)
+
+      # Schedule the processes
+      while queue != []:
+        if new_process == True: # check if there is a new process
+            add_process(process)
+
+        queue.sort(key=lambda a: a.burst) # sort according to burst time
+        next = queue.pop(0) # picks next process
+        use_resource(next) # uses resources until finished
+    \end{minted}
+    \label{code:sjf}
+    \caption{Python: First in, First out}
+\end{figure}
+
+As already mentionned in the section \ref{sec:first-lines-of-code}, python is not really meant for this type of task and therefore it should only be looked at as a simple way of expressing logic.
+
+\section{Upgrading the Cashier}
+
+In this section our cashier recieves a significant upgrade: it can save the state of one buyer and switch to another one.
+In technical terms a policy that can do that is is called a preemtive policy. 
+The switching itself is called a content switch.
+While previously if another customer came with just a single item, it would not get the priviledge to cut infront of the line. 
+He would only recieve the priority boost, if the current person is finished.
+However now with the preemptive capabilities the cashier can just save and put aside the people.
+This feature is not only useful if there is someone that has a shorter time to completion.
+Imagen that someone forgot to put something into their basket.
+Now they desparately send back their kid to get that time. While he returns the cashier can save the state and do a context switch.
+While processes wait for I/O others can run at their place.
+This waiting state is called being blocked and the technique of running another process is overlapping.
+At this stage it does not really make sense to present some code, because it is just too out of place, however in theory one would to the sorting of the processes everytime a job finished, gets blocked or another process joins.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-12-08 LaTeX/Thesis.tex
added glossaries
added "minted" for code blocks
Finished first chapter
sun 8. dec: almost finished basics

@@
 
-\newglossaryentry{LaTeX}{%
-  name        = {\LaTeX},
-  description = {A document preparation system},
-  sort        = {LaTeX},
-}
-
-\newglossaryentry{Real numbers}{%
-  name        = {$\real$},
-  description = {The set of real numbers},
-  sort        = {Real numbers},
-}
-
-\makeindex
-\makeglossaries
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Code Blocks
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\usepackage{minted}
+
+\usepackage{hyperref}
+
 \begin{document}
@@
         %
-        Introduciton to Basics
+        In this chapter I will give a quick introduction to the basics of CPU scheduling.
+We will look at the proper terms and common approaches to solving the problem of how one should schedule the processes waiting to be run.
       \end{singlespace*}
@@
 
-\chapter{Sources}
-\begin{itemize}
-    \item source 1
-    \item source 2 
-\end{itemize}
+\include{Sources/Sources}
 

### 2024-12-09 LaTeX/Sources/Basics.tex
Finished Basics

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Conclusion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{What about fairness?}
+
+All of the policies we looked at so far disregarded the emotions of humans.
+Most of us wanted to be treated at least as well as others.
+It does not matter whether in a supermarket or at home before the computer, we want fairness and responsiveness.
+A fair scheduler like the Round Robin does exactly just that. 
+It is the most simple to implement for a dynamic system, because all it does is goes around and gives every one contestant a bit of CPU time.
+The so called quantum or quanta in plural determines how long a single process can run, before it gets replace by another one.
+In supermarket terms: How many items get scanned per customer per cycle.
+The logical next step would be to make these quanta as small as possible making the system as fair as possible.
+This however has a heavy drawback.
+The context switching itself takes up resources, meaning that after a while the performance degrades, because of too many context switches.
+
+ \section{Conclusion}
+
+During this chapter we learned about the basic methodology of scheduling.
+The next step is to move to more sophisticated approches, which act as a whole system rather than just as a solution to a specific problem.
+We will also need to figure out the burst time.
+To be more precise we need to figure out how to avoid using it.
+Most of the above mentionned policies, except the Round Robin, rely on the fact that we know, when one process finishes.
+Take Shortest Job First for example: How can we order it according to the time to finish if we don't know when it finishes?

### 2024-12-10 LaTeX/Sources/Advanced.tex
started Advanced policies and added SVG template for the graphics

\chapter{Predicting the Future based on the Past}

\begin{figure}[h]
    \centering
    \includesvg[width=0.8\textwidth]{Assets/Template.drawio.svg}
    \caption{template svg file}
\end{figure}

### 2024-12-10 LaTeX/Thesis.tex
started Advanced policies and added SVG template for the graphics

@@
 
+% SVG support
+\usepackage{svg}
+
 \begin{document}
@@
 
+  \part[Advanced Policies]{%
+    Advanced Policies\\
+    %
+    \vspace{1cm}
+    %
+    \begin{minipage}[l]{\textwidth}
+    %
+    \textnormal{%
+      \normalsize
+      %
+      \begin{singlespace*}
+        \onehalfspacing
+        %
+        After looking at the Basics it is time to dive deeper.
+Here we will learn about policies that do not rely on the burst time to function.
+This means that technically they can be used in a real world application.
+In reality however they are usually first modified to fit the environment and the needs better.
+      \end{singlespace*}
+    }
+    \end{minipage}
+  }
+
+  \include{Sources/Advanced}
+
 % This ensures that the subsequent sections are being included as root

### 2024-12-11 LaTeX/Sources/Advanced.tex
added basic idea of mlfq

@@
 \chapter{Predicting the Future based on the Past}
+
+\section{Idea}
+
+The first policy we look at is formerly called Multi-Level Feedback Queue or MLFQ short. The create Fernando J. Corbató recieved a Turning Awaard for it in 1990. 
+As the title already says, this policy tries to predict the future behaviour of the processes based on the past.
+A job can be generally act in two ways.
+Either it is a resources intensive crunching problem (think about exporting a video or compiling code) or it is a program, which needs quick response time (think about your text editor).
+In reality most jobs jump between these two states.
+We usually want to give the response based processes priority, because that is what the user interacts with and it is here that he primarily feels a delay.
+
+The policy is based on multiple queues, which have each different priorities.
+Each process gets assigned to a queue. There are however not set in stone
+Based on the reasons above we want to assume that a new process is responsive, because in the worst case scenario, we need to just demote the non-responsive ones.
+If however a process turns out to be interactive, than the user does not feel any lag. 
+Each process can run a certain amount of time (also called allotment time) before it is deemed as unworthy of the current priority.
+If the allotment is used up the process get demoted into a queue below.
 
@@
     \centering
-    \includesvg[width=0.8\textwidth]{Assets/Template.drawio.svg}
-    \caption{template svg file}
+    \includegraphics[width=0.8\textwidth]{Assets/MLFQ-Example-1.pdf}
+    \caption{Simple working of MLFQ}
+    \label{fig:mlfq-example-1}
 \end{figure}
+
+As you can see in figure \ref{fig:mlfq-example-1} the process one gets demoted after a while.
+Also the lower the queue is, the higher the allotment time. 
+This is because we hope that all of the responsive focused finish before demoting.
+Once these are filtered out we have only resource heavy tasks left.
+These require more time anyways, so the allotment time is stretched out.

### 2024-12-11 LaTeX/Sources/Advanced.tex
added multiple processes mlfq

@@
-\chapter{Predicting the Future based on the Past}
+\chapter{Predicting the Future}
 
@@
 
-The first policy we look at is formerly called Multi-Level Feedback Queue or MLFQ short. The create Fernando J. Corbató recieved a Turning Awaard for it in 1990. 
+The first policy we look at is formerly called Multi-Level Feedback Queue or MLFQ short. The create Fernando J. Corbató recieved a Turning Award for it in 1990. 
 As the title already says, this policy tries to predict the future behaviour of the processes based on the past.
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{Assets/MLFQ-Example-1.pdf}
+    \includegraphics[width=0.7\textwidth]{Assets/MLFQ-Example-1.pdf}
     \caption{Simple working of MLFQ}
@@
 These require more time anyways, so the allotment time is stretched out.
+
+\section{Multiple Processes}
+
+What happens if we introduce another process?
+Well, it depends on the priorities. 
+Higher priority recieves the CPU time.
+If they are on the same queue, than they run using Round Robin, see section \ref{sec:rr}.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.7\textwidth]{Assets/MLFQ-Example-2.pdf}
+    \caption{Running multiple processes in MLFQ}
+    \label{fig:mlfq-example-2}
+\end{figure}
+
+As you can see in figure \ref{fig:mlfq-example-2} once process two is introduced process one is temporarily starved. The problem is solved once they land on the same queue.
+There they run alternately.
+Still the more tasks we introduce the more prevalent the starving issue gets.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.7\textwidth]{Assets/MLFQ-Example-3.pdf}
+    \caption{Starving Processes in MLFQ}
+    \label{fig:mlfq-example-3}
+\end{figure}

### 2024-12-11 LaTeX/Sources/Advanced.tex
finished mlfq

@@
 \end{figure}
+
+\section{Solving the Starving Problem}
+
+Solving starvation is actually easy.
+All we have to make sure that once in a while everybody gets their deserved CPU time.
+To do that we introduce a priority boost. All it does is just put every process into priority one every so often.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.7\textwidth]{Assets/MLFQ-Example-4.pdf}
+    \caption{Solving Starvation in MLFQ}
+    \label{fig:mlfq-example-4}
+\end{figure}
+
+\section{Wrapping Up}
+
+This chapter showed us, how one could approach a scheduling policy that works without needing to know the burst time.
+In the end we still end up with many variable like the allotment time per queue, the amount of queues, the quanta for the Round Robin and others.
+If used in a real system the best approach would be to set some sensible defaults and let the administrator to adjust the parameters when needed.
+
+\chapter{Using Lottery to Schedule Processes}

### 2024-12-11 LaTeX/Sources/Advanced.tex
explaining basic tickets

@@
 
-\chapter{Using Lottery to Schedule Processes}
+\chapter{Scheduling Based on Tickets}
+
+\section{How do tickets work?}
+
+Tickets are a form of currency for the computer.
+These can be handed out by the scheduler and give a process the right to use a resource.
+Usually this is done proportianally to the amount of tickets.
+Meaning a higher share of tickets will lead to a higher amount of CPU time.
+Logically the response time is reverse proportianal.
+These tickets enable easy comparison between processes and their resources.
+One important difference between a currency in real life and tickets is that tickets are consumed when used.
+Practically this means that if you buy something you won't lose your bill.
+This property leads to the tickets representing the share of the CPU that the job has.
+
+\section{Operations of Tickets}
+
+You can do two notable things with tickets. First you can transfer them. 
+This does exactly what it means. You decrease your amount of tickets and increase someone else's. 
+With this you could temporarily boost other processes.
+Say, for example, you have a server and a client.
+If the client needs something from the server, than it could temporarily boost the servers resources by transfering some of his own tickets.
+If course this would require that the server is trusted or else it could just scam you out of your tickets.
+The second way to operate with tickets is ticket inflation and deflation. This can only be done by the owner of the currency and it is only something that we will look at the next section.

### 2024-12-11 LaTeX/Sources/Advanced.tex
Finished ticket currencies

@@
 
-\section{Operations of Tickets}
+\section{Operating with Tickets}
 
@@
 The second way to operate with tickets is ticket inflation and deflation. This can only be done by the owner of the currency and it is only something that we will look at the next section.
+
+\section{Ticket Currencies}
+
+Ticket currencies are a way for parent processes to manage their child processes.
+Yes, that's right, not all processes are built the same.
+A child process is a process that is created by another process.
+A parent process is a process, which has atleast one child process.
+A job can be both at the same time.
+
+The idea behind ticket currencies is that a parent process creates its own unique currency. Therefore he himself can distribute as many bills as he wants. 
+However more custom bills does not equal to a higher share of CPU time in the whole system.
+Take figure \ref{fig:ticket-currencies} for example. Even though process A had created 100 tickets and gave 40 to process A$_1$ in the end it is all about ratios.
+Therefore the overall runtime that A$_1$ has is:
+
+$$\frac{20\text{T}_\text{G}}{50\text{T}_\text{G}} * \frac{40\text{T}_\text{A}}{100\text{T}_\text{A}} = 0.16 \Rightarrow 16\%$$
+
+There is another hidden benefit. If the parent process recieves more global tickets, than if it has a custom currency, there are no further things to do.
+If however it does not have a custom currency than the process has to transfer these new tickets to the child processes. Overall it can get very messy all too quickly.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{Assets/Ticket-Currency.pdf}
+    \caption{Example of a Custom Ticket Currency}
+    \label{fig:ticket-currencies}
+\end{figure}
+
+Another benefit comes with the ticket inflation or deflation operation, which can only be used by the owner of the currency.
+This is the case, because it involves creating or destroying tickets. 
+Still unlike the ticket transfer, ticket inflation does not need a sender and reciever.
+The selected process just get more tickets assigned.
+With this the parent process changes the ration of ownership and therefore boosts the selected job.

### 2024-12-14 LaTeX/Sources/Advanced.tex
started lottery sched

@@
 Meaning a higher share of tickets will lead to a higher amount of CPU time.
-Logically the response time is reverse proportianal.
+Logically the response time is reverse proportional.
 These tickets enable easy comparison between processes and their resources.
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{Assets/Ticket-Currency.pdf}
+    \includegraphics[width=\textwidth]{Assets/Ticket-Currency.pdf}
     \caption{Example of a Custom Ticket Currency}
@@
 
-Another benefit comes with the ticket inflation or deflation operation, which can only be used by the owner of the currency.
-This is the case, because it involves creating or destroying tickets. 
+Another benefit comes with custom currencies: the ticket inflation or deflation operation.
+One needs a custom currency, because it involves creating or destroying tickets. 
 Still unlike the ticket transfer, ticket inflation does not need a sender and reciever.
@@
 With this the parent process changes the ration of ownership and therefore boosts the selected job.
+
+\section{Lottery Scheduling}
+
+Like other things in computer science the name already says a lot. 
+Here randomization is exploited to achieve the proportianal share in runtime.
+The concept that if even over a small scale there will be unavoidable errors, the longer the scheduler runs the more accurate the share becomes.
+Think of the tickets as lottery tickets.
+If the randomization is fair, than the more tickets you have the more you will win.
+There is a downside though.
+The randomizer algorithm can get pretty expensive.
+There is a way to reuse one random seed though, but we will only quickly touch upon it later.
+
+For an example implementation in C take a look at figure \ref{code:lottery-sched}
+This is taken from phd-mit-tr677.pdf (CITATION)
+
+\begin{figure}[h]
+    \begin{minted}[mathescape,
+        linenos,
+        numbersep=5pt,
+        gobble=2,
+        frame=lines,
+        framesep=2mm,
+        ]{cpp}
+  /* Implementation of List-Based Lottery Scheduling Algorithm (Written by: pdh-mit-tr667.pdf CITATION) */
+
+  /*per-client state */
+  typedef struct {
+    ...
+    int tickets;
+  } *client_t;
+
+  /* current resource owner */
+  client_t current;
+
+  /* list of clients competing for resource */
+  list_t list;
+
+  /*global tick et sum */
+  int global_tickets= 0;
+
+  /* initialize client withspecified allocation */
+  void client_init(client_t c, int tickets)
+  {
+    /*initialize client state, updateglobal sum */
+    c->tickets = tickets;
+    global_tickets += tickets;
+    /*join competition for resource */
+    list_insert(list, c);
+  }
+
+  void allocate()
+  {
+    int winner, sum;
+    client_t c;
+    /* randomly select winning tick et */
+    winner = fast_random() % global_tickets;
+    /* searchlist to find client with winning tick et */
+    sum = 0;
+    for (c = list_first(list);
+         c != NULL;
+         c = list_next(list, c))
+    {
+        /*update running sum, stop at winner */
+        sum += c->tickets;
+        if (sum > winner)
+            break;
+    }
+    /*grant resource to winner forquantum */
+    current = c;
+    use_resource(current);
+  }
+    \end{minted}
+    \label{code:lottery-sched}
+    \caption{C: Lottery Scheduling}
+\end{figure}
+
+One could also use Binary trees...

### 2024-12-14 LaTeX/Sources/Advanced.tex
finished binary search

--- 2024-12-11 LaTeX/Sources/Advanced.tex
@@
 With this the parent process changes the ration of ownership and therefore boosts the selected job.
+
+\section{Lottery Scheduling}
+
+Like other things in computer science the name already says a lot. 
+Here randomization is exploited to achieve the proportianal share in runtime.
+The concept that if even over a small scale there will be unavoidable errors, the longer the scheduler runs the more accurate the share becomes.
+Think of the tickets as lottery tickets.
+If the randomization is fair, than the more tickets you have the more you will win.
+There is a downside though.
+The randomizer algorithm can get pretty expensive.
+There is a way to reuse one random seed though, but we will only quickly touch upon it later.
+For an example implementation in C take a look at phd-mit-tr677.pdf (CITATION).
+
+\subsection{Binary Search}
+
+Basing the datastructure on a list is quite inefficient, because of the long search time.
+Everytime you want to reschedule, you have to create a random number and loop through the list.
+However you could use a balanced binary tree to make it much faster.
+This is because a list search time scales with n. 
+Therefore if you add one more element you will have to search through one more element.
+In contrast a binary search scales not with n.
+To be more precise it takes on $O(log(n))$ to find a specific value.
+On average, with each decision we can eliminate half of the remaining candidates.
+Therefore overall it is much faster, especially the more values the list contains.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{Assets/Binary-Search.pdf}
+    \caption{Binary Search Example}
+    \label{fig:binary-search}
+\end{figure}
+
+To understand it more, let's go through an example. Let's say there are 48 tickets and our winning one is 25.
+To understand the decision better, feel free to look at figure \ref{fig:binary-search}.
+For our first decision we need to choose if 25 is above 10 or not.
+It is above ten, so we choose the right path.
+At this point we have only 38 tickets left and we have to readjust our winning ticket.
+We took away 10 tickets from the front so our new winning tickets is 15.
+15 is smaller than 20, so we take a left turn and there is no readjusting needed.
+In the next step we compare 12 and 15.
+15 is larger, which means we take the right path and arrive at our winner process.

### 2024-12-14 LaTeX/Sources/Advanced.tex
added implementation of lott. sched.

@@
 There is a way to reuse one random seed though, but we will only quickly touch upon it later.
-For an example implementation in C take a look at phd-mit-tr677.pdf (CITATION).
+For an example implementation in C take a look at figure \ref{code:lottery-sched}, which is written by (CITATION).
+
+\newpage
+
+\begin{figure}[H]
+    \begin{minted}[mathescape,
+        linenos,
+        numbersep=5pt,
+        gobble=2,
+        frame=lines,
+        framesep=2mm,
+        ]{cpp}
+  typedef struct { /*per-client state */
+    ...
+    int tickets;
+  } *client_t;
+
+  client_t current; /* current resource owner */
+  list_t list; /* list of clients competing for resource */
+  int global_tickets = 0; /* global ticket sum */
+
+  /* initialize client withspecified allocation */
+  void client_init(client_t c, int tickets) {
+    c->tickets = tickets; /* initialize client state */ 
+    global_tickets += tickets; /* update global sum */
+    list_insert(list, c); /* join competition for resource */
+  }
+
+  void allocate() {
+    int winner, sum;
+    client_t c;
+    winner = fast_random() % global_tickets; /* randomly select winning ticket */
+    /* search list to find client with winning ticket */
+    sum = 0;
+    for (c = list_first(list);
+         c != NULL;
+         c = list_next(list, c))
+    {
+        /* update running sum, stop at winner */
+        sum += c->tickets;
+        if (sum > winner)
+            break;
+    }
+    current = c;
+    use_resource(current); /* grant resource to winner for quantum */
+  }
+    \end{minted}
+    \caption{Implementation of List-Based Lottery Scheduling Algorithm\\Written by: pdh-mit-tr667.pdf CITATION}
+    \label{code:lottery-sched}
+\end{figure}
+
+In general implementing dynamic operations is not hard, because the only part that changes is the scheduling part, like choosing a winning ticket.
+The clients structure does not need to be modified.
+In a List-Based system adding and removing clients can be as easy as adding them to the list and adjusting the global max tickets.
 
@@
 
-Basing the datastructure on a list is quite inefficient, because of the long search time.
+Even though it is easy to implement, basing the datastructure on a list is quite inefficient, because of the long search time.
 Everytime you want to reschedule, you have to create a random number and loop through the list.
@@
 To understand it more, let's go through an example. Let's say there are 48 tickets and our winning one is 25.
-To understand the decision better, feel free to look at figure \ref{fig:binary-search}.
+To understand the though process better, feel free to look at figure \ref{fig:binary-search}.
 For our first decision we need to choose if 25 is above 10 or not.
@@
 15 is larger, which means we take the right path and arrive at our winner process.
+
+\subsection{Multiwinner Lottery}

### 2024-12-16 LaTeX/Sources/Advanced.tex
started stride scheduling & finished lottery

@@
 
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
 \subsection{Multiwinner Lottery}
+
+Multiwinner Lottery Scheduling is a semi randomized semi deterministic scheduling method.
+It is a method to try to minimize the cost of the randomization algorithm.
+The basic idea is to reuse the winner.
+After you select your winner you take a specific offset to determine the other winners.
+One of these lotteries is called a superquantum.
+
+\section{Stride Scheduling}
+
+Another approach to implement a proportional-share policy is the so called stride scheduling.
+The tickets are used to assign so called strides to the processes.
+This stride value will be than used to increment a counter called pass.
+The process with the lowest pass value than gets run.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{Assets/Stride-Scheduling.pdf}
+    \caption{Stride Scheduling Example}
+    \label{fig:stride-scheduling}
+\end{figure}
+
+The benefit of stride scheduling is that one doesn't need a randomizer and the policy itself is much simpler to implement.
+With such good benefits there comes some disadvantages as well: The processes have global values, meaning that if a new job wants to join, than it has to be adjusted to the current global circumstances.
+In contrast all you have to do in the lottery is add the assigned tickets to the global maximum tickets.

### 2024-12-27 LaTeX/Sources/Advanced.tex
first corrections

@@
 
-\section{Idea}
+\section{Basic Idea}
 
-The first policy we look at is formerly called Multi-Level Feedback Queue or MLFQ short. The create Fernando J. Corbató recieved a Turning Award for it in 1990. 
+The first policy we look at is formerly called Multi-Level Feedback Queue, or MLFQ short. The creator Fernando J. Corbató recieved a Turning Award for it in 1990. 
 As the title already says, this policy tries to predict the future behaviour of the processes based on the past.
-A job can be generally act in two ways.
+A job can generally act in two ways.
 Either it is a resources intensive crunching problem (think about exporting a video or compiling code) or it is a program, which needs quick response time (think about your text editor).
 In reality most jobs jump between these two states.
-We usually want to give the response based processes priority, because that is what the user interacts with and it is here that he primarily feels a delay.
+We usually want to give the response-focused processes priority, because that is what the user interacts with and it is here that he primarily feels a delay.
 
-The policy is based on multiple queues, which have each different priorities.
+The policy introduces multiple queues, which each have different priorities.
 Each process gets assigned to a queue. There are however not set in stone
-Based on the reasons above we want to assume that a new process is responsive, because in the worst case scenario, we need to just demote the non-responsive ones.
-If however a process turns out to be interactive, than the user does not feel any lag. 
+Based on the reasons above we want to assume that a new process is responsive, because in the worst case scenario, we just need to demote the non-responsive ones.
+If, however a process turns out to be interactive, then the user does not feel any lag. 
 Each process can run a certain amount of time (also called allotment time) before it is deemed as unworthy of the current priority.
-If the allotment is used up the process get demoted into a queue below.
+If the allotment is used up, the process gets demoted into a lower queue.
 
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{Assets/MLFQ-Example-1.pdf}
+    \includegraphics[width=0.8\textwidth]{Assets/MLFQ-Example-1.pdf}
     \caption{Simple working of MLFQ}
@@
 
-As you can see in figure \ref{fig:mlfq-example-1} the process one gets demoted after a while.
-Also the lower the queue is, the higher the allotment time. 
-This is because we hope that all of the responsive focused finish before demoting.
-Once these are filtered out we have only resource heavy tasks left.
+As you can see in figure \ref{fig:mlfq-example-1} process one gets demoted after a while.
+The lower the queue is, the higher the allotment time. 
+This is because we hope that all of the responsive-oriented jobs finish before demoting.
+Once these are filtered out we will only have resource heavy tasks left.
 These require more time anyways, so the allotment time is stretched out.
+After a while the process just ends up at the bottom, at which point it runs until it is finished.
+Keep in mind that if the process gets blocked, meaning that it gives up its CPU before the allotment is used up, then they will stay in the same queue and can use up the rest of the allotment before they get demoted.
+Therefore not the actual time spent in a priority matters but the time spent using the resource.
 
@@
 Higher priority recieves the CPU time.
-If they are on the same queue, than they run using Round Robin, see section \ref{sec:rr}.
+If they are on the same queue, then they run using Round Robin (see section \ref{sec:rr}).
 
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{Assets/MLFQ-Example-2.pdf}
+    \includegraphics[width=0.8\textwidth]{Assets/MLFQ-Example-2.pdf}
     \caption{Running multiple processes in MLFQ}
@@
 
-As you can see in figure \ref{fig:mlfq-example-2} once process two is introduced process one is temporarily starved. The problem is solved once they land on the same queue.
+As you can see in figure \ref{fig:mlfq-example-2}, once process two is introduced process one is temporarily starved. The problem is solved once they land on the same queue.
 There they run alternately.
 Still the more tasks we introduce the more prevalent the starving issue gets.
+Take a look at figure \ref{fig:mlfq-example-3} for example. 
+Here we have four processes instead of two, which results in process 1 being completely left out.
+It could only start running again if the processes 2 to 4 either finish or get demoted into priority 5.
+That takes quite a lot of time and requires that we don't introduce new tasks during the time waiting.
 
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{Assets/MLFQ-Example-3.pdf}
+    \includegraphics[width=0.8\textwidth]{Assets/MLFQ-Example-3.pdf}
     \caption{Starving Processes in MLFQ}
@@
 
-\section{Solving the Starving Problem}
+\newpage
 
-Solving starvation is actually easy.
-All we have to make sure that once in a while everybody gets their deserved CPU time.
+\section{Solving Starvation}
+
+Solving starvation is actually pretty easy.
+All we have to do is to make sure that once in a while everybody gets their deserved CPU time.
 To do that we introduce a priority boost. All it does is just put every process into priority one every so often.
+Even though this introduces another unknown variable that has to be configured, it is just such a crucial element that we have to live with that.
 
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{Assets/MLFQ-Example-4.pdf}
+    \includegraphics[width=0.8\textwidth]{Assets/MLFQ-Example-4.pdf}
     \caption{Solving Starvation in MLFQ}
@@
 This chapter showed us, how one could approach a scheduling policy that works without needing to know the burst time.
-In the end we still end up with many variable like the allotment time per queue, the amount of queues, the quanta for the Round Robin and others.
+In the end we still end up with many variables like the allotment time per queue, the amount of queues, the quanta for the Round Robin and others.
 If used in a real system the best approach would be to set some sensible defaults and let the administrator to adjust the parameters when needed.
+To summarize the Multi-Level Feedback Queue I would like to quote the rules of it from the book OStep \cite{ostep}:
+\begin{quote}
+\emph{
+\begin{enumerate}
+        \item If Priority (A) $>$ Priority (B) $\Rightarrow$ A runs and B doesn't
+        \item If Priority (A) $=$ Priority (B) $\Rightarrow$ A and B run in Round Robin
+        \item When a job enters the system, it si placed at the highest priority
+        \item Once a job uses up its time allotment at a given level, its priority is reduced
+        \item After some period S, move all the jobs in the system to the topmost queue
+\end{enumerate}
+}
+\end{quote}
 
@@
 
-\section{How do tickets work?}
+\section{How Do Tickets Work?}
 
@@
 Logically the response time is reverse proportional.
-These tickets enable easy comparison between processes and their resources.
-One important difference between a currency in real life and tickets is that tickets are consumed when used.
+These tickets enable easy comparison between processes.
+One important difference between a currency in real life and tickets is that tickets are not consumed when used.
 Practically this means that if you buy something you won't lose your bill.
@@
 If the client needs something from the server, than it could temporarily boost the servers resources by transfering some of his own tickets.
-If course this would require that the server is trusted or else it could just scam you out of your tickets.
+If course this requires that the server is trusted or else it could just scam you out of your tickets.
 The second way to operate with tickets is ticket inflation and deflation. This can only be done by the owner of the currency and it is only something that we will look at the next section.
@@
 A parent process is a process, which has atleast one child process.
-A job can be both at the same time.
+A job can be both a parent and a child at the same time.
 
@@
 
-There is another hidden benefit. If the parent process recieves more global tickets, than if it has a custom currency, there are no further things to do.
+There is another hidden benefit. If the parent process recieves more global tickets, then if it has a custom currency, there are no further things to do.
 If however it does not have a custom currency than the process has to transfer these new tickets to the child processes. Overall it can get very messy all too quickly.
@@
 One needs a custom currency, because it involves creating or destroying tickets. 
-Still unlike the ticket transfer, ticket inflation does not need a sender and reciever.
-The selected process just get more tickets assigned.
+Unlike the ticket transfer, ticket inflation does not need a sender and reciever.
+A selected process just gets more tickets assigned.
 With this the parent process changes the ration of ownership and therefore boosts the selected job.
@@
 Here randomization is exploited to achieve the proportianal share in runtime.
-The concept that if even over a small scale there will be unavoidable errors, the longer the scheduler runs the more accurate the share becomes.
+The concept is, that even if over a small scale there will be unavoidable errors, the longer the scheduler runs the more accurate it becomes.
 Think of the tickets as lottery tickets.
@@
 There is a way to reuse one random seed though, but we will only quickly touch upon it later.
-For an example implementation in C take a look at figure \ref{code:lottery-sched}, which is written by (CITATION).
+For an example implementation in C take a look at figure \ref{code:lottery-sched}, which is written by Carl A. Waldspurger \cite{waldspurger95}.
 
@@
     \end{minted}
-    \caption{Implementation of List-Based Lottery Scheduling Algorithm\\Written by: pdh-mit-tr667.pdf CITATION}
+    \caption{Implementation of List-Based Lottery Scheduling Algorithm\\Written by: Carl A. Waldspurger \cite{waldspurger95}}
     \label{code:lottery-sched}
@@
 The clients structure does not need to be modified.
+This is because there are no global states.
+The processes are kept completely isolated from each other and just by alooking at the next scheduling you couldn't know how long the scheduler has been running.
 In a List-Based system adding and removing clients can be as easy as adding them to the list and adjusting the global max tickets.
@@
 In contrast a binary search scales not with n.
-To be more precise it takes on $O(log(n))$ to find a specific value.
+To be more precise it takes $O(log(n))$ to find a specific value.
 On average, with each decision we can eliminate half of the remaining candidates.
@@
 
-To understand it more, let's go through an example. Let's say there are 48 tickets and our winning one is 25.
+To understand it more, let's go through an example. Let's say there are 48 tickets and our winning one is 20.
 To understand the though process better, feel free to look at figure \ref{fig:binary-search}.
-For our first decision we need to choose if 25 is above 10 or not.
+For our first decision we need to choose if 20 is above 10 or not.
 It is above ten, so we choose the right path.
 At this point we have only 38 tickets left and we have to readjust our winning ticket.
-We took away 10 tickets from the front so our new winning tickets is 15.
-15 is smaller than 20, so we take a left turn and there is no readjusting needed.
-In the next step we compare 12 and 15.
-15 is larger, which means we take the right path and arrive at our winner process.
+We took away 10 tickets from the front so our new winning tickets is 10.
+10 is smaller than 20, so we take a left turn and there is no readjusting needed.
+In the next step we compare 12 and 10.
+12 is larger, which means we take the right path and arrive at our winner process, because 10 is contained in 12.
 
@@
 
-Multiwinner Lottery Scheduling is a semi randomized semi deterministic scheduling method.
+Multiwinner Lottery Scheduling is a semi-randomized, semi-deterministic scheduling method.
 It is a method to try to minimize the cost of the randomization algorithm.
@@
 The tickets are used to assign so called strides to the processes.
-This stride value will be than used to increment a counter called pass.
+This stride value will then be used to increment a counter called pass.
 The process with the lowest pass value than gets run.
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{Assets/Stride-Scheduling.pdf}
+    \includegraphics[width=\textwidth]{Assets/Stride-Scheduling.pdf}
     \caption{Stride Scheduling Example}
@@
 
+To understand the policy better, let's loot at the example in figure \ref{fig:stride-scheduling}.
+We have three processes.
+Process A has a stride of 160, Process B 320 and Process C has 30.
+At $t_0$ process A runs.
+Afterwards its B's turn.
+C has a much smaller stride, that's why it will have more turns before it catches up to A.
+After a while it gets to cyclic pattern.
+
 The benefit of stride scheduling is that one doesn't need a randomizer and the policy itself is much simpler to implement.
@@
 In contrast all you have to do in the lottery is add the assigned tickets to the global maximum tickets.
+
+To solve this, one needs to introduce a constant global pass, because otherwise the joining processes will monopolize the CPU. 
+Look at figure \ref{fig:stride-scheduling} for example.
+If a process D joins at the last time interval, than it has to catch up to atleast C. During that time it will have a unreasonable priority over everybody else.
+A promised global pass value would eliminate the problem, because one can adjust the entry pass value according to the current situation.
+If a task goes into a blocked state or leaves the competition temporarily and than rejoins than its pass value will be global stride plus the stride of the process.
+The same goes for a brand new job that joins.

### 2024-12-27 LaTeX/Sources/Basics.tex
added basic idea of mlfq
finished binary search
first corrections

@@
 How can we order the queued tasks, so that they run in the most optimal order?
-What optimal really means is a whole discussion itself not to mention search for the perfect algorithm to achieve that desired best solution.
+What optimal really means is a whole discussion itself (not to mention the search for the perfect algorithm to achieve that desired best solution).
 In this chapter I will give you a quick introduction to the main terms.
@@
 Why they have the whole cart full of milk?
-Maybe these question seem rather sudden they have two things in common:
-One, you never get the answer to them, and two, which is the relevant part, you ask them while standing in line at a supermarket.
-You ask yourself them while staring at the family of five with two shopping carts and while questioning, whether or not you should have queued at another line.
+Maybe these question seem rather sudden, but they have two things in common:
+One, you never get them answered, and two, which is the relevant part, you ask them while standing in line at a supermarket.
+You ask them while staring at the family of five with two shopping carts and while questioning, whether or not you should have queued at another line.
 Also wouldn't it be much better if you, with your two items, went before them?
 What is the best order to queue up these people?
-As you can already see this is the original, simplified question just reformulated.
-We are seaching for the best algorithm that arranges people at the supermarket.
-Matter of fact let's just call it by its real name: policy.
+As you can already see this is the original, simplified question, just reformulated.
+We are searching for the best algorithm that arranges people at the supermarket.
+Matter of fact, let's just call it by its real name: policy.
 In CPU-Scheduling we call the algorithms policies or disciplines.
 A single customer / shopper is called a job or a process.
-In the real world a process can be anything from your drivers to the web browser showing cute cat images.
+In the real world a process can be anything from your drivers to the web browser showing the cute cat images.
 A queue is a line of people waiting to get their items scanned and the cashier is the CPU.
 
-\section{How we compare policies}
+\section{How We Compare Policies}
 
-Before we dive deeper into theory, it is important to note that there is a huge difference between fairness and performance.
+Before we dive deeper into the theory, it is important to note that there is a huge difference between fairness and performance.
 Fairness usually makes sure that everyone recieves the same amount of CPU time.
@@
 Therefore the average of it is the average waiting time until you exit the supermarket.
+Note that when we compare the different policies, we assume that they have the same clients queuing at the same time. 
+This is common sense, because if you have more jobs to process then naturally you will take more time.
 
-\section{First lines of code} \label{sec:first-lines-of-code}
+\section{First lines of Code} \label{sec:first-lines-of-code}
 
@@
 In a real world this is almost impossible\footnote{Here I mean knowning the burst time and not escaping the supermarket.}, except if you can time travel.
-With this and setting aside the optimization part, one of the most straightforward policy is called first come, first served.
+One of the most straightforward policy is called first come, first served.
 This is what we usually suffer through in the queue to pay.
 The policy disregards how many item one has in the cart.
-The only important part is the time you arrive.
+The only important attribute is the time you arrive.
 The sooner the better.
 In the world of computer science the policy is better refered to as First In, First Out (FIFO).
-Here is a \hyperref[code:fifo]{simple python implementation}.
 
@@
   # Schedule the processes
-  while queue != []:
+  while True:
     if new_process == True: # check if there is a new process
@@
 \end{minted}
+\caption{Python: First in, First out}
 \label{code:fifo}
-\caption{Python: First in, First out}
 \end{figure}
-In this example we just assume that the use\_resource function on line 14 is already written.
-Also it is important to mention that the new\_process variable is just a bool, which gets update if a new process is waiting to join the queue.
-In addition to that the empty queue initialization is purely symbolic, because if the queue is already empty, there will be no scheduling happening. 
+In the example above we assume that the use\_resource function on line 14 is already written.
+Also it is important to mention that the new\_process variable is just a bool, which gets updated if a new process is waiting to join the queue.
 In general python is not really meant for low level programming anyways, so take it with a grain of salt.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Evolving Supermarket}
+\chapter{Evolving the Supermarket}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 To give an answer we will first take a step back.
-While the previous chapter talked about the initial situation and proper terms, this one will first loosen up some of the requirements in order to ensure that the policies are as easy to understand as possible.
+While the previous chapter talked about the initial situation and proper terms, this one will first loosen up some of the rules in order to ensure that the policies are as easy to understand as possible.
 
@@
     \end{minted}
+    \caption{Python: First in, First out}
     \label{code:sjf}
-    \caption{Python: First in, First out}
 \end{figure}
@@
 As already mentionned in the section \ref{sec:first-lines-of-code}, python is not really meant for this type of task and therefore it should only be looked at as a simple way of expressing logic.
+Even though there were some major improvements, the fact remains that if a client is too late than it will have to wait at least  until the previous shortest customer finishes.
+This can take a long time and it would be in our best interest to give these newcomers a fair chance.
+It would be great if a rescheduling happens, every time someone joins the queue.
 
@@
 This feature is not only useful if there is someone that has a shorter time to completion.
-Imagen that someone forgot to put something into their basket.
-Now they desparately send back their kid to get that time. While he returns the cashier can save the state and do a context switch.
+Imagine that someone forgot to put something into their basket.
+Now they desparately send back their kid to get that item. While he returns the cashier can save the state and do a context switch.
 While processes wait for I/O others can run at their place.
@@
 
-\section{What about fairness?}
+\section{What about Fairness?} \label{sec:rr}
 
 All of the policies we looked at so far disregarded the emotions of humans.
-Most of us wanted to be treated at least as well as others.
+Most of us want to be treated at least as well as others.
 It does not matter whether in a supermarket or at home before the computer, we want fairness and responsiveness.
@@
 It is the most simple to implement for a dynamic system, because all it does is goes around and gives every one contestant a bit of CPU time.
-The so called quantum or quanta in plural determines how long a single process can run, before it gets replace by another one.
+The so called quantum or quanta in plural determines how long a single process can run before it gets replaced by another one.
 In supermarket terms: How many items get scanned per customer per cycle.
@@
 The context switching itself takes up resources, meaning that after a while the performance degrades, because of too many context switches.
+Therefore the goal is find a quantum, which balances responsiveness and fairness with the performance loss.
 
@@
 
-During this chapter we learned about the basic methodology of scheduling.
-The next step is to move to more sophisticated approches, which act as a whole system rather than just as a solution to a specific problem.
+During this chapter we learned the basic methodology of scheduling.
+The next step is to move to more sophisticated policies, which act as a whole system rather than just as a solution to a specific problem.
 We will also need to figure out the burst time.
 To be more precise we need to figure out how to avoid using it.
-Most of the above mentionned policies, except the Round Robin, rely on the fact that we know, when one process finishes.
-Take Shortest Job First for example: How can we order it according to the time to finish if we don't know when it finishes?
+Most of the above mentioned policies, except the Round Robin, rely on the fact that we know, when one process finishes.
+Take Shortest Job First for example: How can we order the processes according to the time to finish if we don't even know when they finish?

### 2024-12-27 LaTeX/Thesis.tex
first corrections

@@
 \usepackage{bookmark}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Bibliography
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% I like the bibliography to be extremely plain, showing only a numeric
-%% identifier and citing everything in simple brackets. The first names,
-%% if present, will be initialized. DOIs and URLs will be preserved.
-%
-%\usepackage[%
-%  autocite     = plain,
-%  backend      = biber,
-%  doi          = true,
-%  url          = true,
-%  giveninits   = true,
-%  hyperref     = true,
-%  maxbibnames  = 99,
-%  maxcitenames = 99,
-%  sortcites    = true,
-%  style        = numeric,
-%  ]{biblatex}
-%
-%
+% I like the bibliography to be extremely plain, showing only a numeric
+% identifier and citing everything in simple brackets. The first names,
+% if present, will be initialized. DOIs and URLs will be preserved.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subtitle{What waiting in line at the supermarket and your computer have in common}
-\author{Mark Krutzler}
+\author{Mark Krutzler 4a}
 
@@
   \include{Sources/Title}
-  \include{Sources/Abstract}
+  %\include{Sources/Abstract}
 
@@
         %
-        After looking at the Basics it is time to dive deeper.
-Here we will learn about policies that do not rely on the burst time to function.
+        After looking at the basics, it is time to dive deeper.
+Next, we will learn about policies that do not rely on the burst time to function.
 This means that technically they can be used in a real world application.
@@
 
-\include{Sources/Sources}
+\begin{thebibliography}{99}
+    \bibitem{ostep}
+    Remzi Hussein Arpaci-Dusseau \& Andrea Carol Arpaci-Dusseau (2018) \emph{Operating Systems: Three Easy Pieces}, CreateSpace Independent Publishing Platform.
+    \bibitem{coffman68}
+    Edward G. Coffman \& Leonard Kleinrock (1968) \emph{Computer scheduling methods and their countermeasures}, Association for Computing Machinery.
+    \bibitem{waldspurger95}
+    Carl A. Waldspurger (1995) \emph{Lottery and Stride Scheduling: Flexible Proportional-Share Resource Management}, Massachusetts Institute of Technology.
+    \bibitem{waldspurger94}
+    Carl A. Waldspurger \& William E. Weihl (1994) \emph{Lottery scheduling: Flexible proportional-share resource management}, Massachusetts Institute of Technology.
+    \bibitem{solaris}
+    Andrea Carol Arpaci-Dusseau (2000) \emph{Multilevel Feedback Queue Scheduling in Solaris}, available at: \url{https://pages.cs.wisc.edu/~remzi/OSTEP/Citations/notes-solaris.pdf}
+\end{thebibliography}
+
+\section*{Further Sources}
+\begin{itemize}
+    \item \url{https://github.com/Pseudomanifold/latex-mimosis}
+    \item \url{https://en.wikipedia.org/wiki/Binary_search}
+    \item \url{https://www.quora.com/Why-is-look-up-faster-for-a-Binary-Tree-than-a-Linked-List}
+    \item \url{https://ceunican.github.io/aos/09.Scheduling_Proportional_Share.pdf}
+\end{itemize}
 

### 2024-12-27 Visual/compare.py
added basic visualization

@@ -1,0 +1,19 @@
+h1 = 0
+h2 = 0
+h3 = 0
+order = []
+for i in range(20):
+    if i == 0:
+        order.append(0)
+    else:
+        if h1 <= h2 and h1 <= h3:
+            h1 += 1.5
+            print(h1)
+            order.append(1)
+        elif h2 <= h1 and h2 <= h3:
+            h2 += 1
+            order.append(2)
+        else:
+            h3 += 0.5
+            order.append(3)
+print(order)

### 2024-12-27 Visual/testing-project/scene.py

@@ -1,0 +1,105 @@
+from manim import *
+
+class Scheduler(Scene):
+    def stride(self):
+        p1 = []
+        p2 = []
+        p3 = []
+        h1 = 0
+        h2 = 0
+        h3 = 0
+        order = []
+        for i in range(16):
+            if i == 0:
+                order.append(0)
+            else:
+                if h1 <= h2 and h1 <= h3:
+                    h1 += 1.5
+                    print(h1)
+                    order.append(1)
+                elif h2 <= h1 and h2 <= h3:
+                    h2 += 1
+                    order.append(2)
+                else:
+                    h3 += 0.5
+                    order.append(3)
+        label = []
+        h1 = 0
+        h2 = 0
+        h3 = 0
+        for i in range(16):
+            if order[i] == 1:
+                h1 += 1.5
+                p1.append(Square(color=BLUE, stroke_opacity=0, fill_opacity=0.7))
+                p1[i].stretch_to_fit_width(0.5)
+                p1[i].stretch_to_fit_height(h1)
+                p1[i].align_to([0, -2.5, 0], DOWN)
+                p1[i].align_to([-1, -2.5, 0], LEFT)
+            elif i == 0:
+                p1.append(Square(color=BLUE, stroke_opacity=0, fill_opacity=0.7))
+                p1[i].stretch_to_fit_width(0.5)
+                p1[i].stretch_to_fit_height(0)
+                p1[i].align_to([0, -2.5, 0], DOWN)
+                p1[i].align_to([-1, -2.5, 0], LEFT)
+            else:
+                p1.append(p1[i-1])
+
+            if order[i] == 2:
+                h2 += 1
+                p2.append(Square(color=ORANGE, stroke_opacity=0, fill_opacity=0.7))
+                p2[i].stretch_to_fit_width(0.5)
+                p2[i].stretch_to_fit_height(h2)
+                p2[i].next_to(p1[i], RIGHT, buff=0)
+                p2[i].align_to(p1[i], DOWN)
+            elif i == 0:
+                p2.append(Square(color=ORANGE, stroke_opacity=0, fill_opacity=0.7))
+                p2[i].stretch_to_fit_width(0.5)
+                p2[i].stretch_to_fit_height(0)
+                p2[i].next_to(p1[i], RIGHT, buff=0)
+                p2[i].align_to([0, -2.5, 0], DOWN)
+            else:
+                p2.append(p2[i-1])
+
+
[…]

### 2024-12-30 LaTeX/Sources/Advanced.tex
further correction

@@
-\chapter{Predicting the Future}
+\chapter{Advanced Policies}
+After looking at the basics, it is time to dive deeper.
+Next, we will learn about policies that do not rely on the burst time to function.
+This means that technically they can be used in a real world application.
+In reality however, they are usually first modified to fit the environment and the needs better.
 
-\section{Basic Idea}
+\section{Predicting the Future}
 
-The first policy we look at is formerly called Multi-Level Feedback Queue, or MLFQ short. The creator Fernando J. Corbató recieved a Turning Award for it in 1990. 
-As the title already says, this policy tries to predict the future behaviour of the processes based on the past.
-A job can generally act in two ways.
-Either it is a resources intensive crunching problem (think about exporting a video or compiling code) or it is a program, which needs quick response time (think about your text editor).
+The first policy we look at is officially called Multi-Level Feedback Queue, or MLFQ in short. The creator Fernando J. Corbató received a Turning Award for it in 1990. 
+As the title already says, this policy tries to predict the future behavior of the processes based on the past.
+A job can generally act in two ways:
+Either it is a resource-intensive crunching problem (think about exporting a video or compiling code) or it is a program, which needs quick response time (think about your text editor).
 In reality most jobs jump between these two states.
@@
 
+\subsection{Basic Idea}
+
 The policy introduces multiple queues, which each have different priorities.
-Each process gets assigned to a queue. There are however not set in stone
+Each process is assigned to a queue. There are, however, not set in stone.
 Based on the reasons above we want to assume that a new process is responsive, because in the worst case scenario, we just need to demote the non-responsive ones.
-If, however a process turns out to be interactive, then the user does not feel any lag. 
+If however, a process turns out to be interactive, then the user does not feel any lag. 
 Each process can run a certain amount of time (also called allotment time) before it is deemed as unworthy of the current priority.
@@
 As you can see in figure \ref{fig:mlfq-example-1} process one gets demoted after a while.
-The lower the queue is, the higher the allotment time. 
+The lower the queue is, the longer the allotment time. 
 This is because we hope that all of the responsive-oriented jobs finish before demoting.
@@
 
-\section{Multiple Processes}
+\subsection{Multiple Processes}
 
@@
 Well, it depends on the priorities. 
-Higher priority recieves the CPU time.
-If they are on the same queue, then they run using Round Robin (see section \ref{sec:rr}).
+Higher priority receives the CPU time.
+If they are on the same queue, then they run using \emph{Round Robin} (see section \ref{sec:rr}).
 
@@
 
-As you can see in figure \ref{fig:mlfq-example-2}, once process two is introduced process one is temporarily starved. The problem is solved once they land on the same queue.
+As you can see in figure \ref{fig:mlfq-example-2}, once process two is introduced, process one is temporarily starved. The problem is solved once they land on the same queue.
 There they run alternately.
@@
 Here we have four processes instead of two, which results in process 1 being completely left out.
-It could only start running again if the processes 2 to 4 either finish or get demoted into priority 5.
+It could only start running again if processes 2 to 4 either finish or get demoted into priority 5.
 That takes quite a lot of time and requires that we don't introduce new tasks during the time waiting.
@@
 
-\newpage
-
-\section{Solving Starvation}
+\subsection{Solving Starvation}
 
@@
 To do that we introduce a priority boost. All it does is just put every process into priority one every so often.
-Even though this introduces another unknown variable that has to be configured, it is just such a crucial element that we have to live with that.
+Even though this introduces another unknown variable that has to be configured, it is just such a crucial element that we have to live with it.
 
@@
 
-\section{Wrapping Up}
+\newpage
+
+\subsection{Wrapping Up}
 
 This chapter showed us, how one could approach a scheduling policy that works without needing to know the burst time.
-In the end we still end up with many variables like the allotment time per queue, the amount of queues, the quanta for the Round Robin and others.
-If used in a real system the best approach would be to set some sensible defaults and let the administrator to adjust the parameters when needed.
-To summarize the Multi-Level Feedback Queue I would like to quote the rules of it from the book OStep \cite{ostep}:
+In the end we still end up with many parameters like the allotment time per queue, the number of queues, the quanta for the Round Robin and others.
+If used in a real system, the best approach would be to set some sensible defaults and let the administrator adjust the parameters when needed.
+To summarize the Multi-Level Feedback Queue, I would like to quote its rules of it from the book OStep \cite{ostep}:
 \begin{quote}
@@
         \item If Priority (A) $=$ Priority (B) $\Rightarrow$ A and B run in Round Robin
-        \item When a job enters the system, it si placed at the highest priority
+        \item When a job enters the system, it is placed at the highest priority
         \item Once a job uses up its time allotment at a given level, its priority is reduced
@@
 
-\chapter{Scheduling Based on Tickets}
+\newpage
 
-\section{How Do Tickets Work?}
+\section{Scheduling Based on Tickets}
 
-Tickets are a form of currency for the computer.
+Tickets are a form of currency for the computer. Think of it like money.
 These can be handed out by the scheduler and give a process the right to use a resource.
-Usually this is done proportianally to the amount of tickets.
+Usually, the rigth to a resource is reflected proportianally by the amount of tickets.
 Meaning a higher share of tickets will lead to a higher amount of CPU time.
-Logically the response time is reverse proportional.
+Logically, the response time is inversely proportional. The more tickets you have the less time you have to wait until you run.
 These tickets enable easy comparison between processes.
-One important difference between a currency in real life and tickets is that tickets are not consumed when used.
+One important difference between money in real life and tickets is that tickets are not consumed when used.
 Practically this means that if you buy something you won't lose your bill.
@@
 
-\section{Operating with Tickets}
+One could imagen these tickets like reservation places in a restaurant.
+For now it does not matter how the restaurant assigns the places, like it does not matter how the scheduler policy implements the transition from tickets to CPU time.
+The only thing that is important is that in the end you will have your places and can eat at the restaurant.
+
+\subsection{Operating with Tickets}
 
@@
 Say, for example, you have a server and a client.
-If the client needs something from the server, than it could temporarily boost the servers resources by transfering some of his own tickets.
-If course this requires that the server is trusted or else it could just scam you out of your tickets.
+If the client needs something from the server, then it could temporarily boost the servers resources by transfering some of his own tickets.
+Of course this requires that the server is trusted or else it could just scam you out of your tickets.
 The second way to operate with tickets is ticket inflation and deflation. This can only be done by the owner of the currency and it is only something that we will look at the next section.
 
-\section{Ticket Currencies}
+\subsection{Ticket Currencies}
 
@@
 A child process is a process that is created by another process.
-A parent process is a process, which has atleast one child process.
+A parent process is a process, which has at least one child process.
 A job can be both a parent and a child at the same time.
@@
 
-There is another hidden benefit. If the parent process recieves more global tickets, then if it has a custom currency, there are no further things to do.
-If however it does not have a custom currency than the process has to transfer these new tickets to the child processes. Overall it can get very messy all too quickly.
+There is another hidden benefit. If the parent process receives more global tickets, then if it has a custom currency, there are no further things to do.
+If, however, it does not have a custom currency than the process has to transfer these new tickets to the child processes. Overall it can get very messy all too quickly.
 
@@
 One needs a custom currency, because it involves creating or destroying tickets. 
-Unlike the ticket transfer, ticket inflation does not need a sender and reciever.
+Unlike the ticket transfer, ticket inflation does not need a sender and receiver.
 A selected process just gets more tickets assigned.
@@
 
-\section{Lottery Scheduling}
+\subsection{Lottery Scheduling}
 
@@
 There is a way to reuse one random seed though, but we will only quickly touch upon it later.
-For an example implementation in C take a look at figure \ref{code:lottery-sched}, which is written by Carl A. Waldspurger \cite{waldspurger95}.
+For an example implementation in C take a look at listing \ref{code:lottery-sched}, which is written by Carl A. Waldspurger \cite{waldspurger95}.
 
@@
 This is because there are no global states.
-The processes are kept completely isolated from each other and just by alooking at the next scheduling you couldn't know how long the scheduler has been running.
+The processes are kept completely isolated from each other and just by looking at the next scheduling you couldn't know how long the scheduler has been running.
 In a List-Based system adding and removing clients can be as easy as adding them to the list and adjusting the global max tickets.
 
-\subsection{Binary Search}
+\subsubsection{Binary Search}
 
@@
 
-\subsection{Multiwinner Lottery}
+\subsubsection{Multiwinner Lottery}
 
@@
 
-\section{Stride Scheduling}
+\subsection{Stride Scheduling}
 
@@
 
-To solve this, one needs to introduce a constant global pass, because otherwise the joining processes will monopolize the CPU. 
+To solve this, one needs to introduce a constant global pass value, because otherwise the joining processes will monopolize the CPU. 
 Look at figure \ref{fig:stride-scheduling} for example.

### 2024-12-30 LaTeX/Sources/Basics.tex
further correction

@@
+\chapter{The ABCs of CPU scheduling}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Initial Problem}
+\section{Initial Problem}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Welcome to Costco}
+In this chapter, I will give a quick introduction to the basics of CPU scheduling.
+We will look at the proper terms and common approaches to solving the problem of how one should schedule the processes waiting to be run.
+
+\subsection{Welcome to Costco}
 
 The main problem we face can be broken down into a really easy to formulate but hard to answer question.
-How can we order the queued tasks, so that they run in the most optimal order?
+How can we order the queued tasks so that they run in the most optimal order?
 What optimal really means is a whole discussion itself (not to mention the search for the perfect algorithm to achieve that desired best solution).
 In this chapter I will give you a quick introduction to the main terms.
-In addition, I will also try to make it as easy as possible to understand by using a situation that most of us have already suffered through.
+In addition, I will also try to make the entry into the world of CPU scheduling as easy as possible to understand by using a situation that most of us have already suffered through.
 
-Have you ever wondered why so many people buy prebottled water? 
-How they can eat those nasty snack that you despise?
+Have you ever wondered why so many people buy bottled water? 
+How they can eat those nasty snacks that you despise?
 Why they have the whole cart full of milk?
-Maybe these question seem rather sudden, but they have two things in common:
+Maybe these questions seem rather sudden, but they have two things in common:
 One, you never get them answered, and two, which is the relevant part, you ask them while standing in line at a supermarket.
 You ask them while staring at the family of five with two shopping carts and while questioning, whether or not you should have queued at another line.
-Also wouldn't it be much better if you, with your two items, went before them?
+Also, wouldn't it be much better if you, with your two items, went before them?
 What is the best order to queue up these people?
-As you can already see this is the original, simplified question, just reformulated.
+As you can already see, this is the original, simplified question, just reformulated.
 We are searching for the best algorithm that arranges people at the supermarket.
-Matter of fact, let's just call it by its real name: policy.
+As a matter of fact, let's just call it by its real name: policy.
 In CPU-Scheduling we call the algorithms policies or disciplines.
 A single customer / shopper is called a job or a process.
-In the real world a process can be anything from your drivers to the web browser showing the cute cat images.
+In the real world, a process can be anything from your drivers to the web browser showing the cute cat images.
 A queue is a line of people waiting to get their items scanned and the cashier is the CPU.
 
-\section{How We Compare Policies}
+\subsection{How We Compare Policies}
 
 Before we dive deeper into the theory, it is important to note that there is a huge difference between fairness and performance.
-Fairness usually makes sure that everyone recieves the same amount of CPU time.
-This kind of policy turns out to be cyclic, which usually leads to a more responsive system, because each job gets a bit of activity every so often.
-On the other hand if we want to optimize for performance we should look at the so called average turnaround time.
-Turnaround time is nothing more than the time that a person has to stand in line.
-Therefore the average of it is the average waiting time until you exit the supermarket.
-Note that when we compare the different policies, we assume that they have the same clients queuing at the same time. 
-This is common sense, because if you have more jobs to process then naturally you will take more time.
+Fairness usually ensures that everyone receives the same amount of CPU time.
+This kind of policy usually leads to a more responsive system, because each job gets a bit of activity every so often.
+The response time is defined as:
+$$T_{\text{Response}} = T_{\text{First Run}} - T_{\text{Arrival}}$$
+As already mentioned the other measurable aspect is "performance".
+Performance is usually measured using the turnaround time.
+$$T_{\text{Turnaround}} = T_{\text{Completion}} - T_{\text{Arrival}}$$
+There is always a tradeoff between performance and responsiveness, because to achieve the best average turnaround time, the tasks can't be interrupted, or else the time of completion is dragged out.
+Think of the average turnaround time like the predicted time that you have to wait in line until you finish with the payment of your groceries.
+If you constantly get interrupted then this time will go up.
+To achieve the best fairness the tasks need to be interrupted, because else new processes have to wait until the previous finishes.
 
-\section{First lines of Code} \label{sec:first-lines-of-code}
+The two formulas above are for a single process.
+To get a feeling of the overall system, we will look at the average.
+This average highly depends on the system, because the length of the jobs scales with the average turnaround time proportionally.
+If on average there are more items for the cashier to scan, then the average time until he is finished will go up.
+Therefore we’ll have to look at the measurements for a predetermined set of jobs.
+If we do not have the same system, we can not compare the policies.
+
+\subsection{First lines of Code} \label{sec:first-lines-of-code}
 
 During the next few chapters and sections we'll assume that one knows the amount of time that a process will take to finish, also known as burst time.
-Like bursting out of joy, when you finally finish the weekly shopping.
-In a real world this is almost impossible\footnote{Here I mean knowning the burst time and not escaping the supermarket.}, except if you can time travel.
+Burst time is the time that a process has to run until it is finished.
+Like bursting out of joy, when you finally finish your weekly shopping.
+In a real world this is almost impossible\footnote{Here I mean knowning the burst time and not escaping the supermarket.}, unless you can time travel.
 One of the most straightforward policy is called first come, first served.
@@
 The sooner the better.
-In the world of computer science the policy is better refered to as First In, First Out (FIFO).
-
-We save the queue of people as a python list.
-If any other job joins, it will just get appended to the end.
-As for the scheduling it self:
-We just loop through the list until everyone is finished.
+In the world of computer science, the policy is better refered to as \emph{First In, First Out (FIFO)}.
+Take a look at listing \ref{code:fifo} for a basic python implementation.
 
@@
 \end{figure}
+
+We save the queue of people as a python list.
+If any other job joins, it will just get appended to the end.
+As for the scheduling itself:
+We just loop through the list until everyone is finished.
+
 In the example above we assume that the use\_resource function on line 14 is already written.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Evolving the Supermarket}
+\section{Evolving the Supermarket}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 How can we reduce our waiting time?
-To give an answer we will first take a step back.
+To give an answer, we will first take a step back.
 While the previous chapter talked about the initial situation and proper terms, this one will first loosen up some of the rules in order to ensure that the policies are as easy to understand as possible.
 
-\section{Shorterst Job First}
+\subsection{Shorterst Job First}
 
-As the name already says it, the shortest job will be run first.
-In order to achieve this we assume that the cashier knows, who has the least amount of items.
-This buyer will be than handeled before anybody else. As you can already tell, this is not really a fair way to treat your customers.
-Still if it comes down to pure turnaround time it is far better than the basic First In, First Out.
+As the name already says, the shortest job will be run first.
+In order to achieve this, we assume that the cashier knows, who has the least amount of items.
+This buyer will be then handeled before anybody else. As you can already tell, this is not really a fair way to treat your customers.
+Still if it comes down to pure turnaround time it is in most cases far better\footnote{With the assumption that the tasks arrive at the same time.} than the basic \emph{First In, First Out}.
+Waiting for shortest jobs to finish first is far less severe for longer jobs, because they already have a long burst time.
+However making the short processes wait out the long ones increases the turnaround time by a lot.
+For example let's take process A with a burst time of 10 second and process B with 100 seconds. In case of letting A run first we will have $T_A = 10$ and $T_B = 110$.
+Therefore, the average turnaround time is:
+$$\frac{10 + 110}{2} = \frac{120}{2} = 60 \text{ seconds}$$
+In contrast to that if we let process B run first, we will have the turnaround time of 100 and 110. Therefore, the average is:
+$$\frac{110 + 100}{2} = \frac{210}{2} = 105 \text{ seconds}$$
 
@@
 
-As already mentionned in the section \ref{sec:first-lines-of-code}, python is not really meant for this type of task and therefore it should only be looked at as a simple way of expressing logic.
-Even though there were some major improvements, the fact remains that if a client is too late than it will have to wait at least  until the previous shortest customer finishes.
+As already mentionned in the section \ref{sec:first-lines-of-code}, python is not really meant for writing low level code and therefore it should only be looked at as a simple way of expressing logic.
+Even though there were some major improvements, the fact remains that if a client is too late then it will have to wait at least  until the previous shortest customer finishes.
 This can take a long time and it would be in our best interest to give these newcomers a fair chance.
@@
 
-\section{Upgrading the Cashier}
+\subsection{Upgrading the Cashier}
 
-In this section our cashier recieves a significant upgrade: it can save the state of one buyer and switch to another one.
-In technical terms a policy that can do that is is called a preemtive policy. 
+In this section our cashier receives a significant upgrade: it can save the state of one buyer and switch to another one.
+In technical terms a policy that can do that is called a preemtive policy. 
 The switching itself is called a content switch.
-While previously if another customer came with just a single item, it would not get the priviledge to cut infront of the line. 
-He would only recieve the priority boost, if the current person is finished.
-However now with the preemptive capabilities the cashier can just save and put aside the people.
+While previously, if another customer came with just a single item, it would not get the privilege of cutting infront of the line. 
+He would only receive the priority boost, if the current person is finished.
+However, now with the preemptive capabilities the cashier can just save and put aside the people.
 This feature is not only useful if there is someone that has a shorter time to completion.
@@
 Now they desparately send back their kid to get that item. While he returns the cashier can save the state and do a context switch.
-While processes wait for I/O others can run at their place.
+While processes wait for I/O, others can run in their place.
 This waiting state is called being blocked and the technique of running another process is overlapping.
-At this stage it does not really make sense to present some code, because it is just too out of place, however in theory one would to the sorting of the processes everytime a job finished, gets blocked or another process joins.
+At this stage it does not really make sense to present some code, because it is just too out of place, however, in theory, one would do the sorting of the processes everytime a job finishes, gets blocked or another process joins.
 
-\section{What about Fairness?} \label{sec:rr}
+\subsection{What about Fairness?} \label{sec:rr}
 
@@
 It does not matter whether in a supermarket or at home before the computer, we want fairness and responsiveness.
-A fair scheduler like the Round Robin does exactly just that. 
+A fair scheduler like the Round Robin does exactly that. 
 It is the most simple to implement for a dynamic system, because all it does is goes around and gives every one contestant a bit of CPU time.
@@
 The logical next step would be to make these quanta as small as possible making the system as fair as possible.
-This however has a heavy drawback.
+This, however, has a heavy drawback.
 The context switching itself takes up resources, meaning that after a while the performance degrades, because of too many context switches.
@@
 
- \section{Conclusion}
+ \subsection{Conclusion}
 
@@
 We will also need to figure out the burst time.
-To be more precise we need to figure out how to avoid using it.
+To be more precise, we need to figure out how to avoid using it.
 Most of the above mentioned policies, except the Round Robin, rely on the fact that we know, when one process finishes.
-Take Shortest Job First for example: How can we order the processes according to the time to finish if we don't even know when they finish?
+Take \emph{Shortest Job First} for example: How can we order the processes according to the time to finish if we don't even know when they will finish?

### 2024-12-30 LaTeX/Thesis.tex
further correction

@@
 
-  \part[The ABC's of CPU scheduling]{%
-    The ABC's of CPU scheduling\\
-    %
-    \vspace{1cm}
-    %
-    \begin{minipage}[l]{\textwidth}
-    %
-    \textnormal{%
-      \normalsize
-      %
-      \begin{singlespace*}
-        \onehalfspacing
-        %
-        In this chapter I will give a quick introduction to the basics of CPU scheduling.
-We will look at the proper terms and common approaches to solving the problem of how one should schedule the processes waiting to be run.
-      \end{singlespace*}
-    }
-    \end{minipage}
-  }
-
           \include{Sources/Basics}
-
-  \part[Advanced Policies]{%
-    Advanced Policies\\
-    %
-    \vspace{1cm}
-    %
-    \begin{minipage}[l]{\textwidth}
-    %
-    \textnormal{%
-      \normalsize
-      %
-      \begin{singlespace*}
-        \onehalfspacing
-        %
-        After looking at the basics, it is time to dive deeper.
-Next, we will learn about policies that do not rely on the burst time to function.
-This means that technically they can be used in a real world application.
-In reality however they are usually first modified to fit the environment and the needs better.
-      \end{singlespace*}
-    }
-    \end{minipage}
-  }
 
@@
             \item \url{https://ceunican.github.io/aos/09.Scheduling_Proportional_Share.pdf}
+            \item \url{https://courses.cs.washington.edu/courses/cse451/12au/l11.pdf}
           \end{itemize}

### 2024-12-30 LaTeX/Sources/Advanced.tex
added a bit more detail

@@
 In reality however, they are usually first modified to fit the environment and the needs better.
+We will get to know two subclasses of schedulers: real-time and proportional-share.
 
-\section{Predicting the Future}
+\section{Real-Time Scheduling}
 
@@
 We usually want to give the response-focused processes priority, because that is what the user interacts with and it is here that he primarily feels a delay.
+The \emph{MLFQ} is used in the Solaris 2.6 Time-Sharing scheduler.
+It is part of the time sharing subclass, meaning that it schedules based on an event-driven model, which is usually periodic.
 
@@
 
-\section{Scheduling Based on Tickets}
+\section{Proportional-Share Scheduling}
 
+In this chapter we will look at proportional-share scheduling. This type of policy works with weights and assigns them to the processes. For us these weights will take the form of tickets.
 Tickets are a form of currency for the computer. Think of it like money.
@@
 This property leads to the tickets representing the share of the CPU that the job has.
+Higher priority processes are weighted more heavily, meaning that they get more tickets
 

### 2024-12-30 LaTeX/compile.sh

@@ -1,2 +1,3 @@
 #!/usr/bin/env sh
 pdflatex -synctex=1 -shell-escape -interaction=nonstopmode Thesis.tex
+open Thesis.pdf

### 2024-12-30 LaTeX/Sources/Advanced.tex
finished correction

@@
 These can be handed out by the scheduler and give a process the right to use a resource.
-Usually, the rigth to a resource is reflected proportianally by the amount of tickets.
+Usually, the right to a resource is reflected proportionally by the amount of tickets.
 Meaning a higher share of tickets will lead to a higher amount of CPU time.
@@
 This property leads to the tickets representing the share of the CPU that the job has.
-Higher priority processes are weighted more heavily, meaning that they get more tickets
+Higher priority processes are weighted more heavily, meaning that they get more tickets.
 
-One could imagen these tickets like reservation places in a restaurant.
+One could imagine these tickets like reservation places in a restaurant.
 For now it does not matter how the restaurant assigns the places, like it does not matter how the scheduler policy implements the transition from tickets to CPU time.
@@
 Say, for example, you have a server and a client.
-If the client needs something from the server, then it could temporarily boost the servers resources by transfering some of his own tickets.
+If the client needs something from the server, then it could temporarily boost the servers resources by transferring some of his own tickets.
 Of course this requires that the server is trusted or else it could just scam you out of your tickets.
@@
 Like other things in computer science the name already says a lot. 
-Here randomization is exploited to achieve the proportianal share in runtime.
+Here randomization is exploited to achieve the proportional share in runtime.
 The concept is, that even if over a small scale there will be unavoidable errors, the longer the scheduler runs the more accurate it becomes.
@@
 
-  /* initialize client withspecified allocation */
+  /* initialize client with specified allocation */
   void client_init(client_t c, int tickets) {
@@
 
-Even though it is easy to implement, basing the datastructure on a list is quite inefficient, because of the long search time.
-Everytime you want to reschedule, you have to create a random number and loop through the list.
+Even though it is easy to implement, basing the datas tructure on a list is quite inefficient, because of the long search time.
+Every time you want to reschedule, you have to create a random number and loop through the list.
 However you could use a balanced binary tree to make it much faster.
@@
 
-\subsubsection{Multiwinner Lottery}
+\subsubsection{Multi-Winner Lottery}
 
-Multiwinner Lottery Scheduling is a semi-randomized, semi-deterministic scheduling method.
+Multi-winner Lottery Scheduling is a semi-randomized, semi-deterministic scheduling method.
 It is a method to try to minimize the cost of the randomization algorithm.
@@
 After you select your winner you take a specific offset to determine the other winners.
-One of these lotteries is called a superquantum.
+One of these lotteries is called a super-quantum.
 
@@
 At $t_0$ process A runs.
-Afterwards its B's turn.
+Afterwards it's B's turn.
 C has a much smaller stride, that's why it will have more turns before it catches up to A.
@@
 Look at figure \ref{fig:stride-scheduling} for example.
-If a process D joins at the last time interval, than it has to catch up to atleast C. During that time it will have a unreasonable priority over everybody else.
+If a process D joins at the last time interval, then it has to catch up to at least C. During that time it will have a unreasonable priority over everybody else.
 A promised global pass value would eliminate the problem, because one can adjust the entry pass value according to the current situation.

### 2024-12-30 LaTeX/Sources/Basics.tex
finished correction

@@
 Therefore we’ll have to look at the measurements for a predetermined set of jobs.
-If we do not have the same system, we can not compare the policies.
+If we do not have the same system, we cannot compare the policies.
 
@@
 Like bursting out of joy, when you finally finish your weekly shopping.
-In a real world this is almost impossible\footnote{Here I mean knowning the burst time and not escaping the supermarket.}, unless you can time travel.
+In a real world this is almost impossible\footnote{Here I mean knowing the burst time and not escaping the supermarket.}, unless you can time travel.
 One of the most straightforward policy is called first come, first served.
@@
 The sooner the better.
-In the world of computer science, the policy is better refered to as \emph{First In, First Out (FIFO)}.
+In the world of computer science, the policy is better referred to as \emph{First In, First Out (FIFO)}.
 Take a look at listing \ref{code:fifo} for a basic python implementation.
@@
 
-\subsection{Shorterst Job First}
+\subsection{Shortest Job First}
 
@@
 In order to achieve this, we assume that the cashier knows, who has the least amount of items.
-This buyer will be then handeled before anybody else. As you can already tell, this is not really a fair way to treat your customers.
+This buyer will be then handled before anybody else. As you can already tell, this is not really a fair way to treat your customers.
 Still if it comes down to pure turnaround time it is in most cases far better\footnote{With the assumption that the tasks arrive at the same time.} than the basic \emph{First In, First Out}.
@@
 
-As already mentionned in the section \ref{sec:first-lines-of-code}, python is not really meant for writing low level code and therefore it should only be looked at as a simple way of expressing logic.
+As already mentioned in the section \ref{sec:first-lines-of-code}, python is not really meant for writing low level code and therefore it should only be looked at as a simple way of expressing logic.
 Even though there were some major improvements, the fact remains that if a client is too late then it will have to wait at least  until the previous shortest customer finishes.
@@
 In this section our cashier receives a significant upgrade: it can save the state of one buyer and switch to another one.
-In technical terms a policy that can do that is called a preemtive policy. 
+In technical terms a policy that can do that is called a preemptive policy. 
 The switching itself is called a content switch.
-While previously, if another customer came with just a single item, it would not get the privilege of cutting infront of the line. 
+While previously, if another customer came with just a single item, it would not get the privilege of cutting in front of the line. 
 He would only receive the priority boost, if the current person is finished.
@@
 Imagine that someone forgot to put something into their basket.
-Now they desparately send back their kid to get that item. While he returns the cashier can save the state and do a context switch.
+Now they desperately send back their kid to get that item. While he returns the cashier can save the state and do a context switch.
 While processes wait for I/O, others can run in their place.
 This waiting state is called being blocked and the technique of running another process is overlapping.
-At this stage it does not really make sense to present some code, because it is just too out of place, however, in theory, one would do the sorting of the processes everytime a job finishes, gets blocked or another process joins.
+At this stage it does not really make sense to present some code, because it is just too out of place, however, in theory, one would do the sorting of the processes every time a job finishes, gets blocked or another process joins.
 

### 2024-12-31 LaTeX/Sources/Advanced.tex
added cfs and conclusion

@@
 The same goes for a brand new job that joins.
+
+\newpage
+
+\section{THe Linux 2.6 Completely Fair Scheduler}
+
+As of 2023, this scheduler was replaced in the Linux kernel 6.6 by another scheduler based on the \emph{EEVDF}.
+The CFS is therefore technically retired, however it was used for a very long time, more than 15 years in fact.
+The goal of the policy was to be fair. To do this it was based on a proportional-share algorithm.
+
+\subsection{How it works}
+
+The aim is to divide the CPU time completely fairly.
+This means that each process gets a $\frac{100\%}{N}$ share of runtime, where $N$ is the total number of jobs competing.
+The policy turns out to be periodic and in each cycle everyone runs the same amount of time. 
+This is the ideal outcome of course.
+To do this we measure a so called vruntime.
+This is pretty similar to the pass value in stride scheduling, with the main difference being how this value is incremented.
+While in stride scheduling we use strides, in the CFS we increment the vruntime by the time that the process ran.
+If we have to make a scheduling decision, we just choose the process with the smallest vruntime.
+New jobs just receive the lowest vruntime value, so that they get a "priority" in the beginning.
+
+\subsection{Red-Black Binary Trees}
+
+Here I will quickly mention, why a red black binary tree is used, but not how it is implemented.
+That concept is out of the scope of this paper, because the goal is to explain the scheduling policies and not data structures.
+The CFS uses a special Red-Black Binary tree as its data structure.
+The Red-Black Binary tree is a Balanced Binary Search tree.
+This means that it is sorted from lowest vruntime to highest vruntime.
+In addition to that the smallest branch is no smaller than half the size of the longest branch.
+The lowest vruntime is on the left of the tree.
+This is useful, because it is really easy to find the next job to run: It is on the outer most left leaf.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{Assets/Red-Black-Tree.png}
+    \caption{Red-Black Binary Tree\\Source: \url{https://en.wikipedia.org/wiki/Red\%E2\%80\%93black_tree}}
+    \label{fig:stride-scheduling}
+\end{figure}
+
+After the process finished running, we just need to put it back into the right place.
+To do so, we first adjust it's vruntime value by adding the time it ran and than use the insertion algorithm.
+Due to the properties of red-black binary trees, all operations are O(log(n)) expensive.
+
+\subsection{Priorities}
+
+There isn't a special priority queue. The priority is adjusted by weighing the vruntime.
+$$ \text{vruntime}_k = \text{vruntime}_{k-1} + t_{\text{ran}}*\text{weight} $$
+If a process is more important, we adjust the weight, so that the time that it ran is counted as less. 
+This means that even though it ran for the same amount of time as the other, it didn't receive the same "penalty".
+Therefore it runs sooner than any other task.
+In linux the weight is called "niceness" and the $t_{\text{ran}}$ is multiplied by $\frac{1}{\text{niceness}}$
+
+\section*{Conclusion}
+\addcontentsline{toc}{chapter}{Conclusion}
+
+In this paper we went from taking the first steps to trying to understand the real use application. 
+I hoped to make it easy to get started by using relatable a analogy.
+However, there is a point, where one has to abandon old beliefs to grasp newer ones.
+As time went on the algorithms became more and more complex.

### 2024-12-31 LaTeX/Thesis.tex
added a bit more detail
added cfs and conclusion

@@
             \item \url{https://courses.cs.washington.edu/courses/cse451/12au/l11.pdf}
+            \item \url{https://en.wikipedia.org/wiki/Earliest_eligible_virtual_deadline_first_scheduling}
+            \item \url{https://en.wikipedia.org/wiki/Proportional_share_scheduling}
+            \item \url{https://www.youtube.com/watch?v=MkJfuI5_hjc&t=29s}
+            \item \url{https://www.youtube.com/watch?v=qvZGUFHWChY&list=PL9xmBV_5YoZNqDI8qfOZgzbqahCUmUEin&index=1&pp=iAQB}
+            \item \url{https://www.youtube.com/watch?v=95s3ndZRGbk&list=PL9xmBV_5YoZNqDI8qfOZgzbqahCUmUEin&index=2&pp=iAQB}
+            \item \url{https://en.wikipedia.org/wiki/Red%E2%80%93black_tree} 
           \end{itemize}

### 2025-01-01 LaTeX/Sources/Advanced.tex
finished paper

--- 2024-12-30 LaTeX/Sources/Advanced.tex
@@
 After looking at the basics, it is time to dive deeper.
-Next, we will learn about policies that do not rely on the burst time to function.
+Next, we will learn about policies that do not rely on burst time to function.
 This means that technically they can be used in a real world application.
 In reality however, they are usually first modified to fit the environment and the needs better.
-We will get to know two subclasses of schedulers: real-time and proportional-share.
+Finally, we will look at the \emph{Linux 2.6 Completely Fair Scheduler}, so that we know how a real world application could look like.
 
-\section{Real-Time Scheduling}
+\section{Predicting the Future}
 
-The first policy we look at is officially called Multi-Level Feedback Queue, or MLFQ in short. The creator Fernando J. Corbató received a Turning Award for it in 1990. 
+The first policy we look at is officially called \emph{Multi-Level Feedback Queue (MLFQ)}. The creator Fernando J. Corbató received a Turning Award for it in 1990.
 As the title already says, this policy tries to predict the future behavior of the processes based on the past.
@@
 In reality most jobs jump between these two states.
-We usually want to give the response-focused processes priority, because that is what the user interacts with and it is here that he primarily feels a delay.
-The \emph{MLFQ} is used in the Solaris 2.6 Time-Sharing scheduler.
-It is part of the time sharing subclass, meaning that it schedules based on an event-driven model, which is usually periodic.
+We usually want to give the response-focused processes priority, because that is what the user interacts with and it is here that he feels a delay.
+The \emph{MLFQ} is used in the Solaris 2.6 Time-Sharing Scheduler.
 
@@
 
-The policy introduces multiple queues, which each have different priorities.
-Each process is assigned to a queue. There are, however, not set in stone.
-Based on the reasons above we want to assume that a new process is responsive, because in the worst case scenario, we just need to demote the non-responsive ones.
+The policy introduces multiple queues, each with different priorities.
+Each process is assigned to a queue. These are, however, not set in stone.
+Based on the reasons above we want to assume that a new process is responsive.
+If it is not, we just need to demote it.
 If however, a process turns out to be interactive, then the user does not feel any lag. 
@@
 
-As you can see in figure \ref{fig:mlfq-example-1} process one gets demoted after a while.
+As you can see in figure~\ref{fig:mlfq-example-1}, process one gets demoted after a while.
 The lower the queue is, the longer the allotment time. 
@@
 Once these are filtered out we will only have resource heavy tasks left.
-These require more time anyways, so the allotment time is stretched out.
+These do not need a special priority, because the user does not interact them directly.
+They require more time anyways, so the allotment time is stretched out.
 After a while the process just ends up at the bottom, at which point it runs until it is finished.
@@
 Well, it depends on the priorities. 
-Higher priority receives the CPU time.
-If they are on the same queue, then they run using \emph{Round Robin} (see section \ref{sec:rr}).
+Higher priority receives the CPU.
+If they are on the same queue, then they run using \emph{Round Robin} (see Section~\ref{sec:rr}).
 
@@
 
-As you can see in figure \ref{fig:mlfq-example-2}, once process two is introduced, process one is temporarily starved. The problem is solved once they land on the same queue.
+As you can see in figure~\ref{fig:mlfq-example-2}, once process two is introduced, process one is temporarily starved. The problem is solved once they land on the same queue.
 There they run alternately.
-Still the more tasks we introduce the more prevalent the starving issue gets.
-Take a look at figure \ref{fig:mlfq-example-3} for example. 
+Still, the more tasks we introduce the more prevalent the starving issue gets.
+Take a look at figure~\ref{fig:mlfq-example-3} for example.
 Here we have four processes instead of two, which results in process 1 being completely left out.
 It could only start running again if processes 2 to 4 either finish or get demoted into priority 5.
-That takes quite a lot of time and requires that we don't introduce new tasks during the time waiting.
+That takes quite a lot of time and requires that we do not introduce new tasks during the time waiting.
 
@@
 
-Solving starvation is actually pretty easy.
+Solving starvation is pretty easy.
 All we have to do is to make sure that once in a while everybody gets their deserved CPU time.
-To do that we introduce a priority boost. All it does is just put every process into priority one every so often.
-Even though this introduces another unknown variable that has to be configured, it is just such a crucial element that we have to live with it.
+To do that we introduce a priority boost. Priority boost puts every process into priority one every so often.
+Even though this introduces another unknown variable that has to be configured, namely when to priority boost, it is such a crucial element that we will have to live with it.
 
@@
 If used in a real system, the best approach would be to set some sensible defaults and let the administrator adjust the parameters when needed.
-To summarize the Multi-Level Feedback Queue, I would like to quote its rules of it from the book OStep \cite{ostep}:
+To summarize the \emph{Multi-Level Feedback Queue}, I would like to quote its rules from the book OStep~\cite{ostep}:
 \begin{quote}
-\emph{
+\emph{``
 \begin{enumerate}
-        \item If Priority (A) $>$ Priority (B) $\Rightarrow$ A runs and B doesn't
+        \item If Priority (A) $>$ Priority (B) $\Rightarrow$ A runs and B does not
         \item If Priority (A) $=$ Priority (B) $\Rightarrow$ A and B run in Round Robin
@@
 \end{enumerate}
-}
+''}
 \end{quote}
@@
 One important difference between money in real life and tickets is that tickets are not consumed when used.
-Practically this means that if you buy something you won't lose your bill.
+Practically, this means that if you buy something you will not lose your bill.
 This property leads to the tickets representing the share of the CPU that the job has.
@@
 
-One could imagine these tickets like reservation places in a restaurant.
-For now it does not matter how the restaurant assigns the places, like it does not matter how the scheduler policy implements the transition from tickets to CPU time.
-The only thing that is important is that in the end you will have your places and can eat at the restaurant.
+One could imagine these tickets like a reservation in a restaurant. For now it does not matter how the restaurant assigns the seats.
+The same way it does not matter how the scheduler policy implements the transition from tickets to CPU time.
+The only thing that is important is that in the end you will have your place and can eat at the restaurant.
 
@@
 
-You can do two notable things with tickets. First you can transfer them. 
+You can do two notable things with tickets. First, you can transfer them.
 This does exactly what it means. You decrease your amount of tickets and increase someone else's. 
-With this you could temporarily boost other processes.
+With the transfer you could temporarily boost other processes.
 Say, for example, you have a server and a client.
-If the client needs something from the server, then it could temporarily boost the servers resources by transferring some of his own tickets.
+If the client needs something from the server, then it could temporarily boost the servers resources by transferring some of its own tickets.
 Of course this requires that the server is trusted or else it could just scam you out of your tickets.
-The second way to operate with tickets is ticket inflation and deflation. This can only be done by the owner of the currency and it is only something that we will look at the next section.
+The second way to operate with tickets is ticket inflation and deflation. This can only be done by the owner of the currency and it is something that we will look at in next section.
 
@@
 Ticket currencies are a way for parent processes to manage their child processes.
-Yes, that's right, not all processes are built the same.
+Yes, that is right, not all processes are built the same.
 A child process is a process that is created by another process.
@@
 
-The idea behind ticket currencies is that a parent process creates its own unique currency. Therefore he himself can distribute as many bills as he wants. 
-However more custom bills does not equal to a higher share of CPU time in the whole system.
-Take figure \ref{fig:ticket-currencies} for example. Even though process A had created 100 tickets and gave 40 to process A$_1$ in the end it is all about ratios.
+The idea behind ticket currencies is that a parent can create its own unique currency. Therefore he himself can distribute as many bills as he wants.
+However more custom bills do not equal to a higher share of CPU time.
+Take figure~\ref{fig:ticket-currencies} for example. Even though process A has created 100 tickets and gave 40 to process A$_1$ in the end it is all about ratios.
 Therefore the overall runtime that A$_1$ has is:
@@
 There is another hidden benefit. If the parent process receives more global tickets, then if it has a custom currency, there are no further things to do.
-If, however, it does not have a custom currency than the process has to transfer these new tickets to the child processes. Overall it can get very messy all too quickly.
+If, however, it does not have a custom currency than the process has to transfer these new tickets to the child processes. Overall it gets very messy all too quickly.
 
@@
 
-Another benefit comes with custom currencies: the ticket inflation or deflation operation.
-One needs a custom currency, because it involves creating or destroying tickets. 
-Unlike the ticket transfer, ticket inflation does not need a sender and receiver.
+Another benefit comes with custom currencies: the ticket inflation and deflation operations.
+One needs a custom currency, because it involves creating and destroying tickets.
+Unlike the ticket transfer, ticket inflation does not need a sender.
 A selected process just gets more tickets assigned.
-With this the parent process changes the ration of ownership and therefore boosts the selected job.
+With this the parent process changes the ratio of ownership and therefore boosts the selected job.
 
@@
 
-Like other things in computer science the name already says a lot. 
+Like with other things in computer science the name already says a lot.
 Here randomization is exploited to achieve the proportional share in runtime.
-The concept is, that even if over a small scale there will be unavoidable errors, the longer the scheduler runs the more accurate it becomes.
 Think of the tickets as lottery tickets.
+The policy works by having a global tickets counter, which keeps track of all the jobs that compete for the CPU.
+At every scheduling stop a lottery is held, which chooses a winning ticket. The process with the winning ticket will use the resource for a specified quantum.
 If the randomization is fair, than the more tickets you have the more you will win.
+Say for example that process A has three times as many tickets as B. This means that over a time period process A will run three times as often as B.
+Later, while looking at binary search trees, we look at how such a winner is optimally found.
+The fact is, that even if over a small scale there will be unavoidable errors, the longer the scheduler runs the more accurate it becomes.
+Just like how in probability theory the distribution of winners in a larger dataset will more accurately represent the theoretical calculations.
 There is a downside though.
@@
 There is a way to reuse one random seed though, but we will only quickly touch upon it later.
-For an example implementation in C take a look at listing \ref{code:lottery-sched}, which is written by Carl A. Waldspurger \cite{waldspurger95}.
+For an example implementation in C take a look at listing~\ref{code:lottery-sched}, which is written by Carl A. Waldspurger~\cite{waldspurger95}.
 
@@
 
-\begin{figure}[H]
+\begin{listing}[H]
     \begin{minted}[mathescape,
@@
     \end{minted}
-    \caption{Implementation of List-Based Lottery Scheduling Algorithm\\Written by: Carl A. Waldspurger \cite{waldspurger95}}
+    \caption{Implementation of List-Based Lottery Scheduling Algorithm\\Written by: Carl A. Waldspurger~\cite{waldspurger95}}
     \label{code:lottery-sched}
-\end{figure}
+\end{listing}
 
-In general implementing dynamic operations is not hard, because the only part that changes is the scheduling part, like choosing a winning ticket.
+Implementing dynamic operations is not hard, because the only part that changes is the scheduling part, like choosing a winning ticket.
 The clients structure does not need to be modified.
 This is because there are no global states.
-The processes are kept completely isolated from each other and just by looking at the next scheduling you couldn't know how long the scheduler has been running.
-In a List-Based system adding and removing clients can be as easy as adding them to the list and adjusting the global max tickets.
+The processes are kept completely isolated from each other and just by looking at the next scheduling stop you could not know how long the scheduler has been running.
+In a List-Based system adding and removing clients can be as easy as adding them to the list of competition and adjusting the global max tickets.
 
-\subsubsection{Binary Search}
+\subsubsection{Binary Search} \label{sec:bin-sea}
 
-Even though it is easy to implement, basing the datas tructure on a list is quite inefficient, because of the long search time.
+Even though it is easy to implement, basing the data structure on a list is quite inefficient, because of a long search time.
 Every time you want to reschedule, you have to create a random number and loop through the list.
@@
 Therefore overall it is much faster, especially the more values the list contains.
+In Section~\ref{sec:rb-tree} we will look at how the red black tree works. The red black tree is a special binary search tree, which is balanced. More about that later on.
 
@@
 
-To understand it more, let's go through an example. Let's say there are 48 tickets and our winning one is 20.
-To understand the though process better, feel free to look at figure \ref{fig:binary-search}.
+To understand it more, let us go through an example. Say there are 48 tickets and our winning one is 20.
+To understand the though process better, feel free to look at figure~\ref{fig:binary-search}.
 For our first decision we need to choose if 20 is above 10 or not.
-It is above ten, so we choose the right path.
-At this point we have only 38 tickets left and we have to readjust our winning ticket.
+It is above 10, so we choose the right path.
+At this point we have only 38 tickets left and we have to readjust our winning ticket, because we choose the right path.
 We took away 10 tickets from the front so our new winning tickets is 10.
@@
 In the next step we compare 12 and 10.
-12 is larger, which means we take the right path and arrive at our winner process, because 10 is contained in 12.
+12 is larger, which means we take the left path and arrive at our winner process.
 
@@
 
-Multi-winner Lottery Scheduling is a semi-randomized, semi-deterministic scheduling method.
-It is a method to try to minimize the cost of the randomization algorithm.
-The basic idea is to reuse the winner.
+Multi-Winner Lottery Scheduling is a semi-randomized, semi-deterministic scheduling method.
+It is a method that tries to minimize the cost of the randomization algorithm.
+The basic idea is to reuse the previous random winner.
 After you select your winner you take a specific offset to determine the other winners.
@@
 This stride value will then be used to increment a counter called pass.
-The process with the lowest pass value than gets run.
+The process with the lowest pass value gets to run.
+To understand the policy better, let us look at the example in figure 2.7.
 
@@
     \includegraphics[width=\textwidth]{Assets/Stride-Scheduling.pdf}
-    \caption{Stride Scheduling Example}
+    \caption{Stride Scheduling Example}\label{fig:stride-scheduling}
+\end{figure}
+
+We have three processes.
+Process A has a stride of 160, Process B 320 and Process C has 30.
+At $t_0$ process A runs.
+Afterwards it is B's turn.
+C has a much smaller stride, that is why it will have more turns before it catches up to A.
+After a while it gets to periodic pattern.
+
+The benefit of stride scheduling is that one does not need a randomizer and the policy itself is much simpler to implement.
+With such good benefits there comes some disadvantages as well: The processes have global values, meaning that if a new job wants to join, than it has to be adjusted to the current global circumstances.
+In contrast all you have to do in lottery scheduling is add the assigned tickets to the global maximum tickets.
+
+To solve this, one needs to introduce a constant global pass value, because otherwise the joining processes will monopolize the CPU. 
+Look at figure 2.7 for example.
+If a process D joins at the last time interval, then it has to catch up to at least C. During that time it will have a unreasonable priority over everybody else.
+A promised global pass value would eliminate the problem, because one can adjust the entry pass value according to the current situation.
+If a task goes into a blocked state or leaves the competition temporarily and than rejoins than its pass value will be global pass plus the stride of the process.
+The same goes for a brand new job that joins.
+
+\newpage
+
+\section{The Linux 2.6 Completely Fair Scheduler}
+
+In 2023, this scheduler was replaced in the Linux kernel 6.6 by another scheduler based on the \emph{EEVDF}.
+The \emph{CFS} is therefore technically retired, however it was used for a very long time. More than 15 years in fact.
+The goal of the policy was to be fair. To do this it was based on a proportional-share algorithm.
+
+\subsection{How it works}
+
+The aim is to divide the CPU time completely fairly.
+This means that each process gets a $\frac{100\%}{N}$ share of runtime, where $N$ is the total number of jobs competing.
+The policy turns out to be periodic and in each cycle everyone runs the same amount of time. 
+This is the ideal outcome of course.
+To do this we measure a so called vruntime.
+This is pretty similar to the pass value in stride scheduling, with the main difference being how this value is incremented.
+While in stride scheduling we use strides, in the CFS we increment the vruntime by the time that the process ran.
+If we have to make a scheduling decision, we just choose the process with the smallest vruntime.
+New jobs receive the lowest current vruntime value, so that they get a ``priority'' in the beginning.
+
+\subsection{Red-Black Binary Trees} \label{sec:rb-tree}
+
+In this section I will quickly mention, why a red black binary tree is used, but not how it is implemented.
+That concept is out of the scope of this paper, because the goal is to explain the scheduling policies and not data structures.
+The \emph{CFS} uses a special Red-Black Binary tree as its data structure.
+The Red-Black Binary tree is a Balanced Binary Search tree.
+The search property means that it is sorted from the lowest to the highest vruntime.
+The lowest vruntime is on the left of the tree.
+This is useful, because it is really easy to find the next job to run: It is on the outer most left leaf.
+In addition to that, a tree must have the following qualities to qualify as a legit red-black tree:
+\begin{enumerate}
+  \item Every node is either red or black
+  \item The root and leaves (NIL) are black
+  \item If a node is red, then the children must be black
+  \item All paths from a node to its NIL descendants contain the same amount of black nodes
+\end{enumerate}
+Following these rules, the longest branch can not be larger than twice the size of the smallest branch.
+You have only black nodes as the shortest branch and a red / black alternation as the longest branch.
+The beginning and the end are black, which means that there are a maximal of $n-1$ red nodes, where $n$ denotes the black ones.
+In the end for the shortest path we have $n$ nodes and for the longest we have $n + (n-1) = 2n-1$ nodes.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{Assets/Red-Black-Tree.png}
+    \caption{Red-Black Binary Tree, \\Source: \url{https://en.wikipedia.org/wiki/Red\%E2\%80\%93black_tree}}
     \label{fig:stride-scheduling}
@@
 
-To understand the policy better, let's loot at the example in figure \ref{fig:stride-scheduling}.
-We have three processes.
-Process A has a stride of 160, Process B 320 and Process C has 30.
-At $t_0$ process A runs.
-Afterwards it's B's turn.
-C has a much smaller stride, that's why it will have more turns before it catches up to A.
-After a while it gets to cyclic pattern.
+While the process runs, we take out the choosen node.
+After the process finished running, we need to put the node back into the right place.
+To do so, we first adjust its vruntime value by adding the time it ran and then use the insertion algorithm.
+We need a special algorithm to put the node back, else the above mentionned demands are not fulfilled and the red black tree become illegible.
+The special algorithm just rebalances the tree after insertion.
+Due to the properties of red-black binary trees, all operations are O(log(n)) expensive.
 
-The benefit of stride scheduling is that one doesn't need a randomizer and the policy itself is much simpler to implement.
-With such good benefits there comes some disadvantages as well: The processes have global values, meaning that if a new job wants to join, than it has to be adjusted to the current global circumstances.
-In contrast all you have to do in the lottery is add the assigned tickets to the global maximum tickets.
+\subsection{Priorities}
 
-To solve this, one needs to introduce a constant global pass value, because otherwise the joining processes will monopolize the CPU. 
-Look at figure \ref{fig:stride-scheduling} for example.
-If a process D joins at the last time interval, then it has to catch up to at least C. During that time it will have a unreasonable priority over everybody else.
-A promised global pass value would eliminate the problem, because one can adjust the entry pass value according to the current situation.
-If a task goes into a blocked state or leaves the competition temporarily and than rejoins than its pass value will be global stride plus the stride of the process.
-The same goes for a brand new job that joins.
+There is not a special priority queue. The priority is adjusted by weighing the vruntime.
+$$ \text{vruntime}_k = \text{vruntime}_{k-1} + t_{\text{ran}}*\text{weight} $$
+If a process is more important, we adjust the weight, so that the time that it ran is counted as less. 
+This means that even though it ran for the same amount of time as the other, it did not receive the same ``penalty''.
+Therefore it runs sooner and gets overall more time than any other task.
+In linux the weight is called ``niceness'' and the $t_{\text{ran}}$ is multiplied by $\frac{1}{\text{niceness}}$
+
+\section*{Conclusion}
+\addcontentsline{toc}{chapter}{Conclusion}
+
+In this paper we went from taking the first steps to trying to understand a real world application.
+I hoped to make it easy to get started by using a relatable analogy.
+However, there is a point, where one has to abandon old beliefs in order to grasp newer ones.
+As time went on the algorithms became more and more complex.
+In the end there is no perfect fit for all. Each policy has some advantages and disadvantages and fortunately we, as regular consumers, do not need to make a choice.

### 2025-01-01 LaTeX/Sources/Basics.tex
finished paper

@@
 
-The main problem we face can be broken down into a really easy to formulate but hard to answer question.
+The main problem we face can be broken down into a really easy to formulate but hard to answer question:
 How can we order the queued tasks so that they run in the most optimal order?
@@
 In this chapter I will give you a quick introduction to the main terms.
-In addition, I will also try to make the entry into the world of CPU scheduling as easy as possible to understand by using a situation that most of us have already suffered through.
+I will also try to make the entry into the world of CPU scheduling as easy as possible to understand by using a situation that most of us have already suffered through.
 
@@
 Maybe these questions seem rather sudden, but they have two things in common:
-One, you never get them answered, and two, which is the relevant part, you ask them while standing in line at a supermarket.
+One, you never get them answered, and two, which is the relevant part, you ask them while standing in line at the supermarket.
 You ask them while staring at the family of five with two shopping carts and while questioning, whether or not you should have queued at another line.
-Also, wouldn't it be much better if you, with your two items, went before them?
+Also, would not it be much better if you, with your two items, went before them?
 What is the best order to queue up these people?
-As you can already see, this is the original, simplified question, just reformulated.
+As you can see, this is the original, simplified question, just reformulated.
 We are searching for the best algorithm that arranges people at the supermarket.
-As a matter of fact, let's just call it by its real name: policy.
-In CPU-Scheduling we call the algorithms policies or disciplines.
-A single customer / shopper is called a job or a process.
-In the real world, a process can be anything from your drivers to the web browser showing the cute cat images.
-A queue is a line of people waiting to get their items scanned and the cashier is the CPU.
+As a matter of fact, let us just call the algorithm by its real name: policy.
+In the field of CPU-Scheduling we call these algorithms policies or disciplines.
+A single customer / shopper is called a job, process or task.
+In the real world, a process can be anything from your drivers to the web browser showing cute cat images.
+If the queue is a line of people waiting to get their items scanned, then the cashier is the CPU.
 
@@
 $$T_{\text{Response}} = T_{\text{First Run}} - T_{\text{Arrival}}$$
-As already mentioned the other measurable aspect is "performance".
+$T_{\text{Response}}$ measures the time it takes until a process is run for the first time.
+The average response time is a metric, which tends to be lower in fair systems.
+This comes from the fact that every task is treated equally and therefore get to run in a reasonable time.
+As already mentioned the other measurable aspect is ``performance''.
 Performance is usually measured using the turnaround time.
+The turnaround time is the time it takes until a task is finished. This also includes the time that the process has to wait while others run.
 $$T_{\text{Turnaround}} = T_{\text{Completion}} - T_{\text{Arrival}}$$
-There is always a tradeoff between performance and responsiveness, because to achieve the best average turnaround time, the tasks can't be interrupted, or else the time of completion is dragged out.
-Think of the average turnaround time like the predicted time that you have to wait in line until you finish with the payment of your groceries.
-If you constantly get interrupted then this time will go up.
-To achieve the best fairness the tasks need to be interrupted, because else new processes have to wait until the previous finishes.
-
 The two formulas above are for a single process.
 To get a feeling of the overall system, we will look at the average.
-This average highly depends on the system, because the length of the jobs scales with the average turnaround time proportionally.
+This highly depends on the system, because the average turnaround time scales proportionally with the length of the jobs.
 If on average there are more items for the cashier to scan, then the average time until he is finished will go up.
@@
 
-\subsection{First lines of Code} \label{sec:first-lines-of-code}
+There is always a tradeoff between ``performance'' and fairness.
+To achieve the best average turnaround time, the tasks can not be interrupted, or else the time of completion is dragged out.
+Think of the average turnaround time like the predicted time that you have to wait in line until you finish with the payment of your groceries.
+If you constantly get interrupted then this average will go up.
+To achieve the best fairness the tasks need to be interrupted, because else new processes have to wait until the previous finishes.
+This is clearly a contradiction.
 
-During the next few chapters and sections we'll assume that one knows the amount of time that a process will take to finish, also known as burst time.
-Burst time is the time that a process has to run until it is finished.
-Like bursting out of joy, when you finally finish your weekly shopping.
-In a real world this is almost impossible\footnote{Here I mean knowing the burst time and not escaping the supermarket.}, unless you can time travel.
-One of the most straightforward policy is called first come, first served.
-This is what we usually suffer through in the queue to pay.
-The policy disregards how many item one has in the cart.
-The only important attribute is the time you arrive.
-The sooner the better.
-In the world of computer science, the policy is better referred to as \emph{First In, First Out (FIFO)}.
-Take a look at listing \ref{code:fifo} for a basic python implementation.
+\subsection{First Lines of Code} \label{sec:first-lines-of-code}
 
-\begin{figure}[h]
+\begin{listing}[h]
 \begin{minted}[mathescape,
@@
 \label{code:fifo}
-\end{figure}
+\end{listing}
 
-We save the queue of people as a python list.
+During the next few chapters and sections we will assume that one knows the amount of time that a process will take to finish.
+This time is known as burst time.
+Burst time is the time that a process has to run until it is finished.
+In contrast to the turnaround time, the burst time only measures the time that the process runs and does not include waiting.
+Like bursting out of joy, when you finally finish your weekly shopping.
+In a real world this is almost impossible\footnote{Here I mean knowing the burst time and not escaping the supermarket.}, unless you can time travel.
+One of the most straightforward policies is called first come, first served.
+This is what we usually suffer through in the queue to pay.
+The policy disregards how many items one has in the cart.
+The only important attribute is the time you arrive.
+The sooner the better.
+In the world of computer science, the policy is better referred to as \emph{First In, First Out (FIFO)}.
+Take a look at listing~\ref{code:fifo} for a basic python implementation.
+
+We will save the queue of people as a python list.
 If any other job joins, it will just get appended to the end.
@@
 In the example above we assume that the use\_resource function on line 14 is already written.
-Also it is important to mention that the new\_process variable is just a bool, which gets updated if a new process is waiting to join the queue.
-In general python is not really meant for low level programming anyways, so take it with a grain of salt.
+Also, it is important to mention that the new\_process variable is just a bool, which gets updated if a new process is waiting to join the queue.
+In general python is not really meant for low level programming anyways, so take the code with a grain of salt.
 
@@
 
-Now that the scene is set, we return the original question. 
-How can we reduce our waiting time?
+Now that the scene is set, we return to the original question.
+How can we reduce our turnaround time?
 To give an answer, we will first take a step back.
-While the previous chapter talked about the initial situation and proper terms, this one will first loosen up some of the rules in order to ensure that the policies are as easy to understand as possible.
+While the previous chapter talked about the initial situation and proper terms, this one will propose new policies, which improve our metrics in one way or another.
 
@@
 As the name already says, the shortest job will be run first.
-In order to achieve this, we assume that the cashier knows, who has the least amount of items.
-This buyer will be then handled before anybody else. As you can already tell, this is not really a fair way to treat your customers.
+In order to achieve this, we assume that the cashier knows, who the person is with the least amount of items.
+This customer will be then handled before anybody else. As you can already tell, this is not really a fair way to treat your customers.
 Still if it comes down to pure turnaround time it is in most cases far better\footnote{With the assumption that the tasks arrive at the same time.} than the basic \emph{First In, First Out}.
@@
 However making the short processes wait out the long ones increases the turnaround time by a lot.
-For example let's take process A with a burst time of 10 second and process B with 100 seconds. In case of letting A run first we will have $T_A = 10$ and $T_B = 110$.
+For example let us take process A with a burst time of 10 second and process B with 100 seconds. In case of letting A run first, we will have $T_A = 10$ and $T_B = 110$.
 Therefore, the average turnaround time is:
 $$\frac{10 + 110}{2} = \frac{120}{2} = 60 \text{ seconds}$$
-In contrast to that if we let process B run first, we will have the turnaround time of 100 and 110. Therefore, the average is:
+In contrast to that if we let process B run first, we will have the turnaround time of $T_A = 110$ and $T_B = 100$. Therefore, the average is:
 $$\frac{110 + 100}{2} = \frac{210}{2} = 105 \text{ seconds}$$
 
-\begin{figure}[h]
+\begin{listing}[h]
     \begin{minted}[mathescape,
@@
       # Schedule the processes
-      while queue != []:
+      while True:
         if new_process == True: # check if there is a new process
@@
     \label{code:sjf}
-\end{figure}
+\end{listing}
 
-As already mentioned in the section \ref{sec:first-lines-of-code}, python is not really meant for writing low level code and therefore it should only be looked at as a simple way of expressing logic.
-Even though there were some major improvements, the fact remains that if a client is too late then it will have to wait at least  until the previous shortest customer finishes.
-This can take a long time and it would be in our best interest to give these newcomers a fair chance.
-It would be great if a rescheduling happens, every time someone joins the queue.
+As already mentioned in the Section~\ref{sec:first-lines-of-code}, python is not really meant for writing low level code and therefore listing~\ref{code:sjf} should only be looked at as a simple way of expressing logic.
+Even though there were some major improvements, the fact remains that if a client is too late then it will have to wait at least  until the previous customer finishes.
+This can take a long time and it would be in our best interest to give these newcomers a chance to compete.
+It would be great, if a rescheduling happens every time someone joins the queue.
 
@@
 In this section our cashier receives a significant upgrade: it can save the state of one buyer and switch to another one.
-In technical terms a policy that can do that is called a preemptive policy. 
+In technical terms a policy that can do this is called a preemptive policy.
 The switching itself is called a content switch.
-While previously, if another customer came with just a single item, it would not get the privilege of cutting in front of the line. 
+Previously, if another customer came with just a single item, it would not get the privilege of cutting in front of the line.
 He would only receive the priority boost, if the current person is finished.
-However, now with the preemptive capabilities the cashier can just save and put aside the people.
+However, now with the preemptive capabilities the cashier can just save and put aside people.
 This feature is not only useful if there is someone that has a shorter time to completion.
@@
 This waiting state is called being blocked and the technique of running another process is overlapping.
-At this stage it does not really make sense to present some code, because it is just too out of place, however, in theory, one would do the sorting of the processes every time a job finishes, gets blocked or another process joins.
+The evolved version of \emph{SJF} is called \emph{Preemptive Shortest Job First}, or \emph{PSJF}.
+If we are talking about implementation, then in theory, one would do the sorting of the processes every time a job finishes, gets blocked or another process joins.
 
@@
 It does not matter whether in a supermarket or at home before the computer, we want fairness and responsiveness.
-A fair scheduler like the Round Robin does exactly that. 
-It is the most simple to implement for a dynamic system, because all it does is goes around and gives every one contestant a bit of CPU time.
-The so called quantum or quanta in plural determines how long a single process can run before it gets replaced by another one.
+A fair scheduler like \emph{Round Robin} gives exactly that.
+It is the simplest to implement for a dynamic system, because all it does is that it goes around and gives every one of the contestant a bit of CPU time.
+The so called quantum (quanta in plural) determines how long a single process can run before it gets replaced by another one.
 In supermarket terms: How many items get scanned per customer per cycle.
@@
 This, however, has a heavy drawback.
-The context switching itself takes up resources, meaning that after a while the performance degrades, because of too many context switches.
-Therefore the goal is find a quantum, which balances responsiveness and fairness with the performance loss.
+The context switching itself takes up resources, meaning that after a while the performance degrades, due to too many context switches.
+Therefore the goal is finding a quantum, which balances responsiveness and fairness with the performance loss.
 
@@
 To be more precise, we need to figure out how to avoid using it.
-Most of the above mentioned policies, except the Round Robin, rely on the fact that we know, when one process finishes.
-Take \emph{Shortest Job First} for example: How can we order the processes according to the time to finish if we don't even know when they will finish?
+Most of the above mentioned policies, except the \emph{Round Robin}, rely on the fact that we know, when the processes finish.
+Take \emph{Shortest Job First} for example: How can we order the processes according to the time they finish if we do not even know when they will finish?
